%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 05-01-acknowledgments.tex -- 
%% Author          : Aaron A. Kagawa
%% Created On      : Fri Sep 25 14:35:00 1998
%% Last Modified By: Aaron Kagawa
%% Last Modified On: Sat Jul  2 22:11:37 2005
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{acknowledgments}  
I will always cherish the memories of my Graduate education. It has been
a long and difficult journey, but at the same time it has also been a lot
of fun and very educational. Any success and achievements that I have
accomplished are a direct result of the support of many people. To these
people, I say thank you from the bottom of my heart.

There is nothing more important to me than my family. I am eternally
grateful for the strength, support, and above all, the love that all of you
have given to me.

This Master's Thesis would have not been possible without the tutelage of
Professor Philip Johnson. Thanks for your encouragement and guidance. There
is no doubt that you have greatly influenced many of my skills in software
engineering, Java programming, software design, decision making, and team
leadership.

I would also like to send out many thanks to my fellow CSDL members. You
have provided an excellent environment for all of us to achieve our
educational goals with your brilliant minds and outstanding support. Most
of all, thanks for all the fun times we've had together. One day we'll
achieve our goal of being the only PGA tour members with Graduate degrees
in Computer Science! Until then, I think we should stick to hacking Java.

I would also like to thank my committee members, Professor Philip Johnson,
Professor Dan Suthers, and Professor Martha Crosby. Thank you for your
guidance, criticisms, and most of all, praise.

Last of all, I would like to thank the whole Department of Information and
Computer Sciences here at the University of Hawaii Manoa. My undergraduate
and graduate studies have been educational, memorable and believe it or
not, it has been fun. 
\end{acknowledgments}
