%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% budget.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Jul 27 13:31:23 1994
%% Last Modified By: Philip Johnson
%% Last Modified On: Thu Jan 12 10:53:17 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{proposalnsf}
\usepackage[final]{graphicx}

% NSF proposal generation template style file.
% based on latex stylefiles  written by Stefan Llewellyn Smith and
% Sarah Gille, with contributions from other collaborators.

% Fix things so that figures tend to stay away from the last page. 
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

% this handles hanging indents for publications
\def\rrr#1\\{\par
\medskip\hbox{\vbox{\parindent=2em\hsize=6.12in
\hangindent=4em\hangafter=1#1}}}

\def\baselinestretch{1}

\begin{document}

\section*{Budget Justification}

\pagenumbering{arabic}
\renewcommand{\thepage} {F--\arabic{page}}

\subsection*{Salaries}

The project budget provides salary support for the principal investigator
(two summer months) and two graduate research
assistants (11 months) for each of the three years.  The principal
investigator will perform or supervise all major system design enhancements
and empirical studies, and manage and train the graduate student
assistant.

\subsection*{Travel}

The project budget provides funds for two trips per year from Hawaii to the
mainland.  These funds will be used by the principal investigator or
graduate assistant to attend two relevant conferences (BECC, ICSE) per
year.

\subsection*{Materials and Supplies}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise server and several
workstations, along with a printer and networking peripherals.  The project
budget provides funds of \$10,000 per year for each year of the grant to
provide materials and supplies related to this research, including: (a)
computers and storage for research data; (b) software for data analysis;
and (c) energy meters and installation for selected dormitories.

\subsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
The budget provides for two months of summer salary (\$12,918) for the principal investigator for
each of the three years.  A 4\% increase is projected for each of the following two years. 

\paragraph*{Research Assistants.}  
The budget provides for two graduate research assistantships
for each of the three years.  The salary for each research assistant is
based upon an 11 month RA-Step 7 salary (\$19,680) for each year, with a 
4\% increase each year. 

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 3.5\% for the principal
investigator and 25\% for the graduate assistants.

\paragraph*{Overhead.}  
University of Hawaii overhead cost is calculated as 36.7\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).

\end{document}
