%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% project.sensing.tex -- 
%% Author          : Philip Johnson
%% Created On      : Fri Jan 13 07:58:21 2012
%% Last Modified By: Philip Johnson
%% Last Modified On: Wed Jan 18 10:42:05 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Sensing, monitoring, and data gathering}

We will consider a complex micro-grid with a high penetration of
distributed renewable energy generation.  Theoretical studies will be
considered along with extensive simulations in conjunction with working on
a real test-bed, the University of Hawaii at Manoa (UHM) micro-grid.  The
UHM micro-grid test-bed will provide validation on simulation and
theoretical studies.

Today's modern electrical grid system relies on a vast web of distribution
networks (D) and transmission corridors (T) to deliver power to customers.
Tremendous control, visibility and communication resources and standards
exist to support the ``T'' portion.  The ``D'' portion has received much
less attention and poses one of the weakest links to enabling wide adoption
of more distributed resources.  The ``D'' portion tends to be more disperse,
less automated and non-standard across the nation.  Making matters worse,
the utility operators have limited or no visibility to the distributed
resources emerging on their systems.  The influx of customer-sited PV
systems and emerging plug-in-hybrid vehicle technologies deployed at the
distribution level have accelerated the need to accurately model impacts of
these distributed generation (DG) resources on the electric system.  On the
UHM micro-grid test-bed this requires design and installation of
appropriate sensing equipment to monitor environmental resource data,
electrical grid data, and building energy usage.  This is closely tied to
other aspects of the project including modeling and analysis, optimization
and control, and the economic and social implications.
 
This research considers both theoretical and practical aspects of gathering
data from sensor and monitors.  Multiple data types will be gathered
requiring construction of sensor networks and data fusion from the multiple
sensor networks.  A key is to deploy enough monitors and sensors so that
important information can be gathered and processed.  Information that is
needed are events that can will impact the operation of the grid.  This
includes detecting ramps that occur with solar energy due to changing
weather conditions (from cloudy to sunny or vice versa) and faults that can
occur on the electrical grid.  With sufficient sensor networks deployed
intelligent control and optimization strategies can be made to reduce
energy usage in buildings subject to economic considerations and energy
requirements made on labs, classrooms, and offices in the building.  In
this subsection we discuss the data types, data collection, building sensor
networks, and gathering information from the data (sensor fusion,
compression, and data mining).
 
\paragraph{Data types.}
 
To accurately model and analyze the micro-grid along with designing
optimization and control strategies for the grid several sets of data need
to be monitored.  We also need to consider measuring data both spatially
and temporally.  The data that we consider in this proposal are
environmental resource data, electrical grid data, building energy usage,
and data gathered from users of energy on the micro-grid.  This data will
be gathered from our test-bed, the UHM micro-grid.
 
Environmental resource data is monitored so that we can better forecast the
energy generation of distributed renewable energy sources such as solar PV
panels.  Here we build and deploy sensor boxes that measure solar
irradiation, wind speed and direction, temperature, and humidity.  These
sensors and monitors will be placed on rooftops of UHM buildings with key
considerations being how fast we sample data, how many sensors do we
deploy, and where do we place sensors.
 
Electrical grid data will also be gathered from existing utility meters.
Additional meters will be placed at key locations on buses and electrical
lines.  Currents, voltages, complex power, and energy usage will be
measured.  Again key considerations are how fast we sample data, where to
deploy meters, and what sensors and monitors do we use that are already
deployed on the grid.
 
We would like to get better information about load data and this can be
done by deploying sensors and monitors in buildings.  Here we can monitor
energy usage by deploying sensors to measure energy usage in rooms from air
conditioning, electrical equipment, and lighting.  (Philip, perhaps you can
help here, WattDepot).
 
Data is gathered from different sources to understand how energy is
generated and consumed on the micro-grid.  This requires understanding user
behavior.  Recently, an energy dorm competition was held, the Kukui cup
where the goal was for student in the dorm to use knowledge about energy
consumption to reduce energy usage.  Numerous survey results were conducted
to understand student behavior as it related to energy consumption.  This
will continue with our proposal.  (Philip, perhaps you can add here too).
 
\paragraph{Data collection.}

There are several issues to be considered when getting data from sensors, monitors, and meters.
How much information is to be gathered, how do we collect this data, and how do we organize
this data.  The first issue is discussed here with the other two issues discussed in subsequent
subsections.

Where do we place the sensors?  There has been substantial research on the sensor placement
problem from researchers studying sensor networks with applications for environmental
monitoring, monitoring the electrical grid with PMUs, and biomedical monitoring.  Tools from
signal processing, statistics, and machine learning are used.  The problem is often
framed as an optimization problem of minimizing some error criterion or maximizing some
information criterion.    Once space has been discretized it can be formulated as placing
$m$ sensors among $n>m$ possible locations.  Optimal algorithms are computationally
infeasible for large $m$ and $n$, but greedy algorithms that run in polynomial time of $m$
and $n$ often give good approximations to optimal algorithms.   Some key aspects of the
optimization include properties of submodularity and monotonicity.    We study these algorithms
as it applies to our algorithm for both static and dynamic cases.    A state space model can
be used to model the dynamics of different parameters with a linear state space model given
by 
\begin{displaymath}
x(k+1) = A x(k) + B u(k)
\end{displaymath}
\begin{displaymath}
y(k) = C (x(k) + v(k))
\end{displaymath}
where $x(k)$ is the state of the system (e.g. on the electrical grid, current phasor at $n$ different
nodes).  A goal is to determine the best $C$ matrix which is a binary 0/1 matrix with at most one
1 in any column and exactly one 1 in every row.

We must also determine how often to sample data.    The sensor placement problem and sampling 
problem depend on the information we need to control and optimize the micro-grid.  This
depends on the modeling and analysis of the micro-grid and also what events we are trying
to detect.  Other factors that come into play are the economics associated with placing sensors,
monitors, and meters.  A goal is to place and sample sensors so that they give the most
useful information.    With limited resources we would consider placing sensors on the
grid and to monitor the environment along an electrical line where there is significant
renewable generation and hence intermittent generation.  This will give us a better understanding
of effects of renewable energy generation, how we can model this, and designing control
and optimization algorithms that can mitigate intermittency subject to available resources;
batteries and demand response.

 
\paragraph{Sensor network Implementation.}

We will create two sensor networks at the micro-grid level.  One sensor
network will consist of environmental resource data collected from student
built self powered energy boxes that measure solar irradiance, wind speed
and direction, temperature, and humidity.  There will also be other
environmental sensors such as solar resource locational monitors (e.g. LM2
monitors).  A goal is to monitor environmental resource data both spatially
and temporally on the UHM campus.  Most of these sensors will be placed on
top of UHM buildings.  This will aid in understanding energy produced by
distributed PV sources placed around the campus.  Wireless networks will be
used to gather this data to a central server and storage system.

A second sensor network will monitor the electrical grid.  This will
consist of sampling existing HECO utility meters and deploying demand side
load monitors (e.g.  AMI and PMI type devices) and PMI field device data
concentrators.  These sensors will be connected together via a wireless
network and again sent to a central server and storage system.

(Also need a discussion of building sensor networks: Philip)

\paragraph{Information gathering.}

We have a Smart Energy Laboratory in 493 Holmes to collect the data from different
sensor networks.  This will be stored in a data server and storage system.  Data will
be gathered from sensor networks monitoring environmental resource data, electrical grid
data, and sensor networks in buildings monitoring energy usage.  This data will be time
stamped.  In addition there will be data collected from students, staff, and faculty on campus
about their energy usage.  A goal in the data collection is to perform sensor fusion so that
information can be obtained to  assist in detecting key events that happen on the
micro-grid (e.g ramps in energy produced by distributed solar PV and faults that occur on
the electrical grid).  

Tools are needed to organize and fuse the raw data from sensor networks.  This includes
using  filtering, sensor fusion, source coding, compressed sensing,  and machine learning to
process the data to extract useful information from the data.  Once this is done models can
be created of the micro-grid to analyze and simulate the behavior of the grid.   Then control
and optimization algorithms can be tested on simulation networks to deal with events such
as ramping and faults.  Software packages such as Powerworld, matlab, and PScad will
be used along with visualization tools such as TREX designed by Referentia.




  


 



