\chapter{Appendix}
\section {Appendix A:  Growing list of Gamification thought leaders}

\textbf{Jane McGonigal}, Author of "Reality is Broken", works at the "Institute for the Future" and founder of "Gameful".  Twitter: @avantgame
\\\\
\textbf{Jesse Schell}, Author of "The Art of Game Design". works at Carnegie Mellon’s Entertainment Technology Center. Twitter: @jesseschell
\\\\
\textbf{B.J. Fogg}, teaches at Stanford, inventor of Captology. Twitter: @bjfogg
\\\\
\textbf{Dennis “Dens” Crowley},  co-founder of Foursquare and teaches (adjunct) at NYU. Twitter: @dens
\\\\
\textbf{Amy Jo Kim}, Researcher, CEO of Shufflebrain, author of "Google Talk: Putting the Fun in Functional". Twitter: @amyjokim
\\\\
\textbf{Natron Baxter}, Developer of "Evoke", blog ”Fun is not the enemy of work”. Twitter: @natronbaxter
\\\\
\textbf{Jen McCabe}, Founder of healthmonth.com, mobile health game imoveyou,  Twitter: @jensmccabe
\\\\
\textbf{Nicole Lazzaro}, President of XEODesign. Author of "Four keys to Fun". Twitter: @nicolelazzaro
\\\\
\textbf{Raph Koster}, Author of "A Theory of Fun for Game Design".  Twitter: @raphkoster
\\\\
\textbf{Mark Pincus}, Co-founder and CEO of Zynga. Twitter: @markpinc (not active)
\\\\
\textbf{Gabe Zichermann}, founder of Gamification Co.  Twitter: @gzicherm
\\\\
\textbf{Eric Zimmerman}, Co-author of "Rules of Play", blog "Being Playful". Twitter: @zimmermaneric (not active)
\\\\
\textbf{Keith Lee}, CEO and co-founder of Booyah (MyTown). Twitter: @keithlee0 (not active)
\\\\
\textbf{Byron Reeves}, Professor at Stanford, Co-founder of Seriosity, Inc, author of "Total Engagement" and talk "Work Sucks, Games are Great". Twitter: @Seriosity
\\\\
\textbf{Colleen Macklin}, Professor at the Parsons New School for Design and the director of the PETLab. Twitter: @colleenmacklin
\\\\
\textbf{Sebastian Deterding}, Researcher and UX designer, blog "gamification-research.org", web "codingconduct.cc". Twitter: @dingstweets
\\\\
\textbf{Keith Smith}, Co-Founder and CEO of BigDoor. Twitter: @ChiefDoorman
\\\\
\textbf{The Gamification Daily},  daily newsletter at  http://paper.li/busterbenson/gamification
