\documentclass[12pt,oneside]{report}

%%% Load some useful packages:
%% "New" LaTeX2e graphics support.
\usepackage{graphicx}
%%	using final option to force graphics to be included even in draft mode
%\usepackage[final]{graphicx}
%% Tell graphicx the default directory for all figures
\graphicspath{{figures/}}

% Enable subfigure support
\usepackage{subfigure}

%% Make subsubsections numbered and included in ToC
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}

%% Package to linebreak URLs in a sane manner.
\usepackage{url}

%% Define a new 'smallurl' style for the package that will use a smaller font.
\makeatletter
\def\url@smallurlstyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{smallurl}

%% Define 'tinyurl' style for even smaller URLs (such as in tables)
\makeatletter
\def\url@tinyurlstyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\scriptsize\ttfamily}}}
\makeatother

%% Provides additional functionality for tabular environments
\usepackage{array}

%% Puts space after macros, unless followed by punctuation
\usepackage{xspace}

%% Make margins less ridiculous
\usepackage{fullpage}

%% Allows insertion of fixme notes for future work
\usepackage[footnote, nomargin]{fixme}

%%%% Turned off for tech report, should be turned on for research portfolio
%% Turn on double spacing
\usepackage{setspace}
\usepackage{mdwlist}
\doublespacing

%% Make URLs clickable
%\usepackage[colorlinks, bookmarks=false]{hyperref}
\usepackage[colorlinks, bookmarks=true]{hyperref}

%% Since I'm using the LaTeX Makefile that uses dvips, I need this
%% package to make URLs break nicely
\usepackage{breakurl}

\usepackage{todonotes}
\usepackage{amsmath,amsfonts}
\numberwithin{equation}{subsection}
%%\usepackage{nonfloat}
\usepackage{bbm}
\usepackage{setspace}
\onehalfspacing
\usepackage{tabularx}

%%% End of preamble
\begin{document}

\input{title.tex}

%% Philip suggests it needs a ToC
\tableofcontents

\begin{abstract}
%Abstract goes here if needed.
Software process defines a structure imposed on a human activity resulting in a software. 
Many different kinds of software processes were designed up today, however they are not
proven to deliver consistently and fail with equal probability. 
This phenomena was widely observed and identified as a ``software crisis'' in 1968 and 
since then it was the subject of numerous research aiming the process control and improvement. 
Today formal pocess models, standards, guidlines and recomendations provide teams 
and individuals with easy-to-understand methodologies and allow a great flexibility of 
processes for project needs, but the rate of failing software projects remains about the same. 
Recently, as an alternative opinion - the ``software development process as a craft'' 
idea emerged - this approach emphasizes roles of highly motivated, creative and skilled
individuals in software creation. Software craftsmanship paradighm of software processes 
finding  some strong supporters in the industrial and research communities recently. 

However both views, while supported by excellent research work and industrial success stories, 
follow conventional ``top-down'' technique - at first someone has to invent a
process, design and implement its building blocks and empirically evaluate it after. 
The problem with this approach, as shown in \cite{citeulike:9758924} by van der Alast,
is that one assumes an idealized versions of real processes and tend to produce 
``paper lions'' - process models which are likely to be disruptive and unacceptable for end users.

In my work I explore an alternative approach for the software process analysis - 
through the discovery of recurrent behaviors (habits) from software process artifacts trails.
Previously it was shown that it is possible to automatically recognize \cite{citeulike:2703162} 
test-driven development patterns, in my work I expand this previous experience by intoducing and 
exploring the applicability of a new temporal data mining technique in oder to find 
unknown recurrent behaviors in the space of all possible permutations of existing processes. 
Finding and understanding these recurrent behaviors may shed light on programming 
habits and their role and interactions within a software process.
\end{abstract}

% my main text chapters
\input{introduction}
\input{software_process}
\input{process_mining}
\input{getting_data}
\input{software_process_discovery}
\input{case_studies}
\input{epilogue}

%%% Input file for bibliography
\bibliography{seninp}
%% Use this for an alphabetically organized bibliography
\bibliographystyle{plain}
%% Use this for a reference order organized bibliography
%\bibliographystyle{unsrt}
%% Try using this BibTeX style that hopefully will print annotations in
%% the bibliography. This will allow me to make notes on papers in the
%% BibTeX file and have them readable in the references section until
%% I turn them into a conceptual literature review 
%\bibliographystyle{annotation}

\end{document}
