\documentclass[12pt,oneside]{report}

%%% Load some useful packages:
%% "New" LaTeX2e graphics support.
\usepackage{graphicx}
%%	using final option to force graphics to be included even in draft mode
%\usepackage[final]{graphicx}
%% Tell graphicx the default directory for all figures
\graphicspath{{figures/}}

% Enable subfigure support
\usepackage{subfigure}

%% Make subsubsections numbered and included in ToC
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}

%% Package to linebreak URLs in a sane manner.
\usepackage{url}

%% Define a new 'smallurl' style for the package that will use a smaller font.
\makeatletter
\def\url@smallurlstyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{smallurl}

%% Define 'tinyurl' style for even smaller URLs (such as in tables)
\makeatletter
\def\url@tinyurlstyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\scriptsize\ttfamily}}}
\makeatother

%% Provides additional functionality for tabular environments
\usepackage{array}

%% Puts space after macros, unless followed by punctuation
\usepackage{xspace}

%% Make margins less ridiculous
\usepackage{fullpage}

%% Allows insertion of fixme notes for future work
\usepackage[footnote, nomargin]{fixme}

%%%% Turned off for tech report, should be turned on for research portfolio
%% Turn on double spacing
\usepackage{setspace}
\usepackage{mdwlist}
\doublespacing

%% Make URLs clickable
%\usepackage[colorlinks, bookmarks=false]{hyperref}
\usepackage[colorlinks, bookmarks=true]{hyperref}

%% Since I'm using the LaTeX Makefile that uses dvips, I need this
%% package to make URLs break nicely
\usepackage{breakurl}

\usepackage{amsmath,amsfonts}
\numberwithin{equation}{subsection}
%%\usepackage{nonfloat}
\usepackage{bbm}
\usepackage{setspace}
\onehalfspacing
\usepackage{tabularx}

%%% End of preamble
\begin{document}

\input{title.tex}

%% Philip suggests it needs a ToC
\tableofcontents

\begin{abstract}
%Abstract goes here if needed.
Software process defines a structure imposed on a human activity resulting in a software. Many different kinds of software processes were designed up today. However, they are not proven to deliver consistently and fail with equal probability. This phenomena was widely observed and defined as a ``software crisis'' in 1968. Since then it was studied by generations of researchers. Over these years the research in software process matured and yielded a number of standardized approaches for the process design mechanisms and general principles aimed a ``crisis'' resolution. However, the overall picture changed a little. As an alternative opinion - the ``software development process as a craft'' idea, which emphasize unique roles of developers in creating software, emerged and found some strong supporters in the industrial and research communities. While both views on the software process composition are supported by excellent research work and industrial success stories, in general, software processes remain designed ``top-down'' - i.e. first someone has to invent a process, design and implement its building blocks and try it after. In this work I explore an alternative approach for the software process analysis - through the discovery of recurrent behaviors from software process artifacts trails. These behaviors are thought to be the finest building blocks of larger processes and their understanding may potentially shed light on the larger scale blocks and their interactions within a software process and its formal models.
\end{abstract}

% my main text chapters
\input{introduction}
\input{vc-intro}
\input{eclipse-report}

%%% Input file for bibliography
\bibliography{seninp}
%% Use this for an alphabetically organized bibliography
\bibliographystyle{plain}
%% Use this for a reference order organized bibliography
%\bibliographystyle{unsrt}
%% Try using this BibTeX style that hopefully will print annotations in
%% the bibliography. This will allow me to make notes on papers in the
%% BibTeX file and have them readable in the references section until
%% I turn them into a conceptual literature review 
%\bibliographystyle{annotation}

\end{document}