%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 04-14.tex -- Thesis white paper - software inspections
%% Author          : Aaron A. Kagawa
%% Created On      : Mon Sep 23 11:52:28 2004
%% Last Modified By: Aaron Kagawa
%% Last Modified On: Sat Nov 20 13:21:38 2004
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2004 Aaron A. Kagawa
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\documentclass[11pt,twocolumn]{article} 
\input{/export/home/csdl/tex/psfig/psfig}
\usepackage{/export/home/csdl/tex/icse2003/latex8}
\usepackage{times}
%% A verbatim-like environment which allows font changes
%%\usepackage{alltt}
%% New LaTeX2e graphics support
\usepackage[final]{graphicx}
% uncomment the % away on next line to produce the final camera-ready version
% and uncomment the \thispagestyle{empty} following \maketitle
\pagestyle{empty}

\begin{document}

\title{Limited Resource Software Inspections}

\author{\protect\begin{tabular}{ccc}
Aaron A. Kagawa \\
\end{tabular}\\
\em  Collaborative Software Development Laboratory \\
\em  Department of Information and Computer Sciences \\
\em  University of Hawai'i \\
\em  Honolulu, HI 96822 \\
\em  kagawaa@hawaii.edu}
\maketitle
\thispagestyle{empty}

\begin{abstract}  % 200 words
Imagine that your project manager has budgeted 200 person-hours for the
next month to inspect newly created source code.  Unfortunately, in order
to inspect all of the documents adequately, you estimate that it will
take 400 person-hours.  However, your manager refuses to increase the
budgeted resources for the inspections.  How do you decide
which documents to inspect and which documents to skip?

The classic definition of Inspection, does not provide any advice on how to
handle this situation. For example, the notion of entry criteria used in
Software Inspection determines when documents are ready for inspection
rather than if inspection is needed at all \cite{Ebenau94}.

%% I could talk about previous approaches here. Sampling and Up-Stream documents

This proposed research will investigate how to prioritize inspection
resources and apply them to areas of the system that need them the most.
It is commonly assumed that defects are not uniformly distributed across
all documents in a system - that a relatively small subset of a system
accounts for a relatively large proportion of defects \cite{Boehm01}.  If
inspection resources are limited, then it will be most effective to
identify and inspect the defect-prone areas.

To accomplish this research, I will construct an evaluation framework based
upon automated process and product measures to distinguish documents that
are in ``most need of inspection'' from those in ``least need of
inspection''.  Some examples of the process and product measures that are
being considered include: reported defects, unit tests, test coverage,
active time, and number of changes. Based on this framework I hypothesize,
that code deemed in most need of inspection will generate more critical
issues than code deemed least need of inspection.

%% Each measure acts as an indedendent variable for determining the inspection
%% candidacy of a document and can be assigned an individual weight. 

Each measure affects the determination of ``most and least'' differently.
For example, suppose that test coverage should be weighted more than active
time. Therefore, weights of each measure will be calibrated based on my
initial guesses.  My research will employ a very simple evaluation
strategy, which includes selecting code to inspect, analyzing the results,
and adjusting the calibration of the measures.

There are three milestones that measure my progress in this reasearch.
Milestone 1: implementation of Hackystat Extension, January 2005.
Milestone 2: completed evaluation, March 2005. Milestone 3: thesis
submission and defense, April 2005.
\end{abstract}

\input{04-14-intro}

\input{04-14-relatedwork}

\input{04-14-system}

%\include{04-14-evaluation}
\input{04-14-evaluation}

%\input{04-14-initalresults}

\input{04-14-contribution}

\input{04-14-timeline}


\bibliographystyle{/export/home/csdl/tex/icse2003/latex8}
\bibliography{/export/home/csdl/techreports/04-14/04-14,/export/home/csdl/bib/csdl-trs,/export/home/csdl/bib/hackystat,/export/home/csdl/bib/ftr}
\end{document}










