%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% uhtest-body.tex -- 
%% Author          : Robert Brewer
%% Created On      : Fri Oct  2 16:30:37 1998
%% Last Modified By: Robert Brewer
%% Last Modified On: Mon Oct  5 16:01:29 1998
%% RCS: $Id: uhtest-body.tex,v 1.1 1998/10/06 02:07:14 rbrewer Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\chapter{Introduction}

It is widely recognized that the global climate is warming due to
anthropogenic sources [?]. There are an increasing number of people
interested in making personal changes to reduce their contribution to
climate change. We focus our efforts on these people who are actively
seeking to reduce their carbon footprint. These users have questions
about how best to direct their efforts, such as “how much additional
electricity does increasing the thermostat on the air conditioner by one
degree consume?” or “how much less carbon is released by carpooling with
someone who lives nearby rather than driving alone?” We need to provide
a system that allows users to perform informal experiments related to
their daily lives and provide rapid feedback on the results of those
experiments.

Another important question these users face is “what are the relative
contributions of different activities to my carbon footprint (driving,
air travel, heating/cooling home, entertainment, food, consumer
purchases)?” While tracking usage in individual areas (home electricity
usage, automobile gasoline consumption) is important, the comparative
contributions to the user’s carbon footprint must be determined for
rational decision–making. This approach allows users to prioritize among
the many possible ways they can reduce their environmental impact.


\section{Thesis Statement}

%% related work is in an external file so it can be used for my research
%% portfolio and for my ICS 699 final report
%% Using input instead of include because include cannot be nested. The input
%% file has the chapter heading so it will carry over to the report as well.
\input{proposal-related-work}

\chapter{Research Questions}

\section{Do COTS Sensors Cover A Sufficient Range Of Activities To Support
Estimation Of Carbon Footprint From Sensor Data?}

\section{Is Sensor Data Accurate Enough To Calculate Carbon Footprints?}

\section{How Closely Do Carbon Footprint Estimators Reflect People's Actual
Activities Inferred From Sensor Data?}

\section{Does PET's Aggregation Of Data Help Users To Reduce Carbon Footprint?}

\section{Does PET's "What If?" Functionality Allow Users To Make Better Choices
On How To Reduce Their Carbon Footprint}



\chapter{System Description}

\section{Overview}

\subsection{System Architecture}


\section{Server-Side Data Aggregation}

\subsection{REST API}


\section{Sensor Inputs}

\subsection{Sensors Targeted}

\subsection{Data Upload}


\section{Client User Interface}

\subsection{Web Application}

\subsection{Mobile Application}



\chapter{Experiment Design and Analysis}

\section{Sensor Accuracy}

\section{Carbon Footprint Estimators vs Calculations From Sensor Data}

\section{User Study}

Woodruff et al recruited through ``contacts in green movements, contacts made at green home tours, and advertising in community email lists.'' \cite{Woodruff2008-bright-green}.

\subsection{Questionnaire}



\chapter{Research Plan}



\chapter{Conclusion}
