\documentstyle [nftimes,12pt,
                /home/4/johnson/Tex/definemargins,
                /home/4/johnson/Tex/letteradj] {letter}  

%Default: \definemargins{1.2in}{1.2in}{1.2in}{1.2in}{0.3in}{0.3in}
%               {top}    {bottom} {left}  {right} {headup} {footdown}
\definemargins {0.75in} {1.0in}  {1.0in} {1.0in} {0.25in} {0.25in}

% CHECK THIS!
%\makelabels

\begin {document}
\name {Philip Johnson}
\signature {Philip Johnson, Ph.D.\\
            Associate Professor}

\address{548 Kaimake Loop\\
         Kailua, HI 96734}
\telephone{(808) 956-3489}  %%department phone
%\telephone{(808) 262-3845}  %%home phone
\fax{(808) 956-3548}
\netaddress{johnson@hawaii.edu}
\indentclosing

\letterhead{coins}
\begin {letter}
  {
  Janice Morita, Fiscal Officer
  College of Natural Sciences
  Bilger Hall
  }


\opening {{\bf Justification for Late Travel Request Submission}}

The purpose of this letter is to justify the late arrival of my recent
travel request documentation to the Dean's Office.  The current policy is
that travel requests where no advance is required should arrive at
Disbursing at least five days prior to travel.  At least two days should be
added for processing by the ICS Department and the Dean's Office.
These policies are necessary in order to ensure, for example, that 
health insurance remains in effect during the absence. 

In the current situation, I submitted the travel documentation only two
(working) days before travel began.  I can offer only the following
explanation (it does not really constitute a justification) for this
occurrence. Historically, I have been quite good about filing travel
requests on time.  However, this travel was different from prior trips in
two ways: (1) my wife left for a two week business trip prior to my travel,
and (2) my parents arrived from the Mainland to take care of my children in
my wife's and my absence.  With the responsibilities to help my wife
prepare for her travel, be the sole parent after she left, prepare for my
parent's arrival, and prepare for my own trip, I simply failed to 
carry out all of these responsibilities (such as the travel request) in a
timely fashion.  

I do recognize that not filing the travel request on time has severe
consequences for me personally, such as lack of health coverage. I also
realize that late requests put unnecessary strain on administrators who
then feel obligated to rush the form through in order to keep health
coverage in effect for the travelor.  I sincerely apologize, and do not
intend to make this error again.

\closing{Sincerely,}

\end{letter}
\end{document}
