\begin{thebibliography}{99}

%1
\bibitem {outliers_survey}
Gupta, M., Gao, J., Aggarwal, C.C., Han, J.,
{\em Outlier Detection for Temporal Data: A Survey},
IEEE Trans. on Knowledge and Data Engineering, 25, 1 (2013)

%2
\bibitem{chan_anomaly} 
Chandola, V., Cheboli, D., and Kumar, V.,
{\em Detecting Anomalies in a Time Series Database},
CS TR 09--004 (2009)

%3
\bibitem{hawkins} 
Hawkins, D. M.,
{\em Identification of Outliers},
Chapman and Hall, 1980

%4
\bibitem {hot_sax}
Keogh, E., Lin, J., Fu, A.,
{\em HOT SAX: Efficiently Finding the Most Unusual Time Series Subsequence},
In Proc. ICDM'05

%6
\bibitem {kolmogorov}
Kolmogorov, A.N.,
{\em Three approaches to the quantitative definition of information},
Problems Inform. Transms., 1965

%7
\bibitem {solomonoff}
Solomonoff, R.J.,
{\em A formal theory of inductive inference. Part I},
Rockford Research Institute, Inc., 1962

%8
\bibitem {li_vitanyi}
Li, M. and Vit{\'a}nyi, P.M.B.,
{\em An Introduction to Kolmogorov Complexity and Its Applications},
Springer-Verlag, 1993

%9
\bibitem{mdl}
Gr{\"u}nwald, P. D.,
{\em The Minimum Description Length Principle},
MIT Press, 2007

%10
\bibitem{ncd}
Cilibrasi, R., Vit{\'a}nyi, P.M.B.,
{\em Clustering by compression}, 
IEEE Trans. Inform. Theory, 2005

%11
\bibitem {param_free}
Keogh, E., Lonardi, S., Ratanamahatana, C.A.,
{\em Towards parameter-free data mining},
In Proc. KDD, 2004

%12
\bibitem {grigorieff}
Ferbus-Zanda, M., Grigorieff, S.,
{\em Is Randomness ``Native'' to Computer Science?},
arXiv, 2008

%13
\bibitem {per_lof}
Martin-L\"{o}f, Per,
{\em On the definition of random sequences},
Information and Control, MIT, 9:602-61 (1966)

%14
\bibitem{compression}
Nevill-Manning, C. and Witten, I.,
{\em Linear-Time, Incremental Hierarchy Inference for Compression},
In Proc. of IEEE Conference on Data Compression, 1997

%5
\bibitem {sax}
Patel, P., Keogh, E., Lin, J., Lonardi, S.,
{\em Mining Motifs in Massive Time Series Databases},
In Proc. ICDM, 2002

%15
\bibitem {grammarviz}
Li, Y., Lin, J., and Oates, T.,
{\em Visualizing variable-length time series motifs},
In Proc. of SDM, 2012

%16
\bibitem {sequitur}
Nevill-Manning, C. and Witten, I.,
{\em Identifying Hierarchical Structure in Sequences: A linear-time algorithm},
Journal of Artificial Intelligence Research, 7, 67-84 (1997)

%17
\bibitem {lin_motifs}
Lin, J., Keogh, E., Patel, P., and Lonardi, S.,
{\em Finding Motifs in Time Series},
The 2nd Workshop on Temporal Data Mining, the 8th ACM Int'l Conference on KDD, 2002

%18
\bibitem {grammarviz2}
Senin, P., Lin, J., Wang, X., Oates, T., Gandhi, S., Boedihardjo, A.P., Chen, C., Frankenstein, S., Lerner, M.,
{\em GrammarViz 2.0: a tool for grammar-based pattern discovery in time series},
In Proc. ECML/PKDD, 2014

\bibitem {dutchpd}
Van Wijk, J.J. and Van Selow, E.R., 
{\em Cluster and calendar based visualization of time series data},
In Proc. IEEE Symposium on Information Visualization, 1999

%16
\bibitem {physionet}
Goldberger, A.L. et al., 
{\em PhysioBank, PhysioToolkit, and PhysioNet: components of a new research resource for complex physiologic signals},
Circulation, 101(23) (2000)

\bibitem {hilbert}
Hilbert, D.,
\"{U}eber stetige Abbildung einer Linie auf ein Fl\"{a}chenst\"{u}ck. 
Mathematische Annalen, 38:459--460, 1891

\bibitem {msequitur}
Oates, T., Boedihardjo, A., Lin, J., Chen, C., Frankenstein, S., Gandhi, S.,
{\em Motif discovery in spatial trajectories using grammar inference},
In Proc. of ACM CIKM, 2013

%19
\bibitem {hashing} 
Wei, L., Keogh, E., Xi, X.,
{\em SAXually explicit images: Finding unusual shapes},
In Proc. ICDM, 2006

%20
\bibitem{haar_1} 
Fu, A., Leung, O., Keogh, E., Lin, J.,
{\em Finding Time Series Discords based on Haar Transform},
In Proc. of Intl. Conf. on Adv. Data Mining and Applications, 2006

%10
\bibitem{haar_2} 
Bu, Y., Leung, O., Fu, A., Keogh, E., Pei, J., Meshkin, S.,
{\em WAT: Finding Top-K Discords in Time Series Database},
In Proc. of SIAM Intl. Conf. on Data Mining, 2007

%11
\bibitem{disk}
Yankov, D., Keogh, E., Rebbapragada, U.,
{\em Disk aware discord discovery: finding unusual time series in terabyte sized data sets},
Knowledge and Information Systems, 241-262 (2008)

%12
\bibitem{viztree}
Lin, J., Keogh, E., Lonardi, S., Lankford, J.P., Nystrom, D. M.,
{\em Visually mining and monitoring massive time series},
In Proc. ACM SIGKDD Intn`l Conf. on KDD, 2004

%13
\bibitem{ano_pattern} 
Chen, X., Zhan, Y.,
{\em Multi-scale Anomaly Detection Algorithm based on Infrequent Pattern of Time Series},
J. of Computational and Applied Mathematics, 2008

%14
\bibitem{bitmaps} 
Wei, L., Kumar, N., Lolla, V., Keogh, E., Lonardi, S., Ratanamahatana, C.,
{\em Assumption-free Anomaly Detection in Time Series},
In Proc. SSDBM, 2005


%17
\bibitem{grammarviz_src}
Paper authors. Supporting webpage:
\url{http://github.com/GrammarViz2}

\end{thebibliography}
