%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Plain-Tex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex -- 
%% Author          : Rosemary Andrada
%% Created On      : Tue Jul  4 17:16:14 1995
%% Last Modified By: Rosemary Andrada
%% Last Modified On: Fri Jul  7 11:16:51 1995
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1995 Rosemary Andrada
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\acknowledgements{

Many people contributed to this thesis, each with very different roles.
Without the support of every single person, I could not have survived as I
have.  In return, I humbly give my thanks to these people.

To Philip Johnson, who always provided the right words in the face of
self-doubt, I thank you for your encouragement and for never letting me do any
less than I was capable of.  I can see so clearly how you have guided your
students well, having learned from your mentors.  I hope to give back in the
same way someday.

Martha Crosby has been my mentor, advisor, boss, committee member, alternate
mother; you name it and she has done it.  I cannot begin to count the ways she
has helped to positively shape my future.  Your selfless deeds are appreciated
more than you know.

Kazuo Sugihara, whom I hold in the highest regard, thank you for your advice
and help throughout the years.  I realize you had a million other things to do
besides serving on my committee and yet you found the time and made the effort.

I could not have kept my sanity nor meet the deadline without Cam Moore there
with me every step of the way.  Working on our theses together, we could have
taken the long road or the easy way out, but we stuck together in the hard
times and made it through.

I give thanks to the rest of my research group, the Collaborative Software
Development Laboratory (CSDL).  You were the first community with which I felt
importance and belonging.

I am grateful to all of my friends at school, especially Lucy Weldon, Sally
Dunan, Stan Yamada, and Steve Buczynski.  Thank you all for your support,
encouragement, stress-relieving capabilities and most of all, for letting me
talk your ears off.

I want to thank my dear friends, Donna Noriye and Amber Knight, who supported
me across the miles.  I give thanks to my wonderful sister, Julie Andrada.
Your faith in me speaks a thousand words.  I want to thank my parents and my
brothers, Arthur and Arnold, none of whom really understood the painful process
of doing a thesis, but who put up with me graciously anyway.

Finally, I am forever indebted to Jose Sumajit, Jr.  You gave me everything you
had in you, and then some.  Having you by my side was motivation enough to get
me through the rough spots.  I could never give you back the time you lost when
you spent it with me, as if you had done a thesis too.  I could never give you
back all the energy it took to keep me level-headed long enough to finish my
work.  You did not have to do any of this, but I am so glad you did.  For this
reason, it is to you that I dedicate this thesis.

}



