%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% nsf93-budget.tex -- 
%% RCS:            : $Id: nsf93-budget.tex,v 1.2 93/09/01 17:26:30 johnson Exp $
%% Author          : Philip Johnson
%% Created On      : Thu Aug 12 11:07:02 1993
%% Last Modified By: Philip Johnson
%% Last Modified On: Wed Sep  1 17:26:10 1993
%% Status          : Unknown
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1993 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%% History
%% 12-Aug-1993		Philip Johnson	
%%    

\section{Budget}

\subsection{Budget justification}

\paragraph{Salaries.}  The project budget provides salary support for 
the principal investigator and four graduate student assistants.  The
principal investigator will be responsible for performing major design
activities on the CSRS system, development of experimental designs for the
empirical studies, management and training of the graduate student
assistants, and coordination with external sites during evaluation and
adoption of CSRS.  The effort necessary to perform the four major
activities associated with this research project is considerable and will
require four graduate students to assist the principal investigator with
these tasks. This project is fortunate to have four talented and
experienced M.S. and Ph.D. level students already associated with it.

\paragraph{Travel.}  Funds for two trips from Hawaii to the mainland 
each year are included in the budget for all four years.  These funds will
be used by the principal investigator to attend relevent conferences where
he may report on the results of his own work and obtain first hand
information on related work in automated support for software review.
Funds for trips to external sites to supervise installation of CSRS will be
provided by the sponsoring organization.

\paragraph{Equipment.}  
The Department of Information and Computer Sciences has provided the
Principal Investigator with one SPARCstation 10 workstation, and the
Department in conjunction with the PI's previous NSF RIA award has provided
him with an additional SPARCstation 10 workstation.  Equipment funding is
budgeted to obtain a third SUN Sparcstation10 workstation in the first year
of the grant.  Three workstations are needed to support the intensive
software development activities of the five personnel (four graduate
students and the principal investigator) on this research project.
Department-supplied, public access X Window terminals and other systems
will be used by the two remaining personnel to access the three Unix
workstations during periods of high demand.

\paragraph{Supplies.}
\$2,000 is budgeted for miscellaneous small office supplies, software, and
hardware in each of the four years.  These purchases are needed to 
support equipment upgrades and miscellaneous clerical work involved in 
this research.

\subsection{Cost breakdown}
\label{cost-breakdown}

The costs associated with this research project can be broken down by year
and category, as illustrated in NSF Form 1030.  The next section documents
how the dollar amounts for each cost category were calculated.  The
following section documents two potential contract options if funding for
the full four-year award is not available.

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
The first cost category is Summer Salary for the Principal Investigator.
Summer salary is calculated as two additional months of the Principal
Investigator's annual nine month salary.  The annual nine month salary for
each of the next three years is estimated by taking the current annual nine
month salary (\$55,452 at Instructional Rank 3, nine month, Step 13) and
estimating a 5\% yearly increase, starting in the first year.

\paragraph{Research Assistants.}  
This cost category provides support for four graduate research
assistants.  The salary for each research assistant is derived by
taking their current annual salary (\$15,558 for Graduate Assistant
Stipend, 11 month, Step 5) and estimating a 5\% yearly increase, 
starting in the first year.

\paragraph{Fringe benefits.} 
Fringe benefits are calculated as the sum of 2\% of the summer salary
plus 15\% of the research assistant salaries.

\paragraph{Overhead.}  
University of Hawaii overhead cost is calculated as 45\% of
salaries, travel, and supplies (equipment is excluded).


\subsection{Contract Options}

This project has been designed to accomodate two contract options at
reduced funding levels, although such funding reductions entail reductions
in the scope of the research.  These contract options also support awarding
of this grant on a continuing basis, with an initial two year funding
period to be followed by two one year extensions, where continuing funding
will be based upon demonstration of adequate performance during the
preceding period.

\begin{itemize}
\item {\em Continuing development and empirical studies only (two years).}
  Under this contract option, funding would be provided only for the first
  two research activities in this project.  This corresponds to funding 
  only Years 1 and 2 of the research, with a total budget of approximately
  \$275,000.  
  
\item {\em Continuing development, empirical studies, and public release
  (three years).} Under this contract option, the project would not attempt
  the final activity involving the creation of an on-line database of
  review data.  This corresponds to funding only Years 1-3 of the research,
  with a total budget of approximately \$425,000. 
 
\end{itemize}








