%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% project-background.tex -- 
%% Author          : Philip Johnson
%% Created On      : Tue Mar 31 11:44:58 2009
%% Last Modified By: Philip Johnson
%% Last Modified On: Wed Dec 16 15:29:18 2009
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2009 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{Results from prior NSF support}

The PIs for this project have had the following NSF-supported projects during the past five years:

\begin{enumerate}
\item P. Johnson, {\em Human centered information integration for the Smart Grid}, NSF
  Grant IIS-1017126, 8/15/10 - 7/31/14, \$413,467. The objective of this
  research is to design information technology and associated experimental
  methods to help understand what information, provided in what ways and at
  what times, enables consumers to make positive, sustained changes to
  their energy consumption behaviors. Selected publications include:
  \cite{csdl2-10-05,csdl2-10-07,csdl2-10-08,csdl2-11-02,csdl2-11-03,csdl2-12-06,csdl2-11-07, csdl2-12-12,csdl2-13-10,csdl2-13-05,csdl2-13-03}.
  
\vspace{-.1in}

\item A. Kuh, {\em Incremental and Distributed Learning in Nonstationary
    Environments with Applications to Wind Forecasting}, NSF Grant ECCS-098344,
  9/01/09 - 8/31/13, \$150,251.  The objective of this research is to
  design novel nonlinear kernel online and distributed learning algorithms
  for applications including wind forecasting.   Research was also conducted
  to model the microgrid using a factor-graph framework. 
  Selected publications for
  this project include \cite{kuh-etal10,kowahl-kuh,hu-etal10,hu-kuh-yang-kavcic,hu-kuh-kavcic-nakafuji,ji-wei-kuh,uddin12,kuh-isess,carland,navid}.
  
%\vspace{-.1in}

\item A. Kuh, {\em US-Japan Joint Seminar Information Theory}, NSF Grant 0508025
  \$35,750.  Funds used to support graduate students for conferences and
  for visit to Japan.

\end{enumerate}

%\vspace{-.1in}

Profs. Johnson and Kuh also had other NSF support dating back more than 5 years.


