%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% bio.kuh.tex -- 2011 Sustainable Energy Pathways Proposal
%% Author          : Philip Johnson
%% Created On      : Tue Nov  4 10:26:48 1997
%% Last Modified By: Philip Johnson
%% Last Modified On: Thu Jan 12 10:41:48 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{center}
Anthony Kuh \\
Electrical Engineering \hfill (808) 956-7527

University of Hawaii              \hfill fax: (808) 956-3426

1680 East-West Road               \hfill kuh@hawaii.edu

Honolulu, HI~~~96822              \hfill http://www.ee.hawaii.edu/~kuh/info.html

\end{center}

\begin{Professional Preparation}
\item Ph.D. in Electrical Engineering, Princeton University. 1987
\item M.A.  in Electrical Engineering, Princeton University. 1984
\item M.S.  in Electrical Engineering, Stanford University. 1980
\item B.S.  in Electrical Engineering and Computer Science, University of California, Berkeley. 1979
\end{Professional Preparation}

\begin{Appointments}
\item Director, Renewable Energy and Island Sustainability (REIS) Group, Univ. of Hawaii.  2009--present
\item Chair of Electrical Engineering, University of Hawaii. 2007-- present
\item Interim Director, Hawaii Center for Advanced Communications, University of Hawaii. 2000--2001
\item Chair of Electrical Engineering, University of Hawaii. 1999--2002
\item Graduate Chair of Electrical Engineering, University of Hawaii. 1998--1999
\item Professor of Electrical Engineering, University of Hawaii. 1997--present
\item Associate Professor of Electrical Engineering, University of Hawaii. 1991--1997
\item Consultant, Siemens Corporate Research, Princeton, NJ. 1988--1993
\item Assistant Professor of Electrical Engineering, University of Hawaii. 1986--1991
\item Teaching and Research Assistant in Electrical Engineering, Princeton University. 1982--1986
\item Member of Technical Staff, Network Performance Planning Center, Bell Labs, Holmdel, N.J.. 1979--1982

\end{Appointments}


\begin{Publications: Closely Related}

\item  D. Bakken, A. Bose, K. Mani Chandy, P. Khargonekar, A. Kuh, S. Low, S. von Meier, 
K. Poolla, P.P. Varaiya, F. Wu.``GRIP: Grid with Intelligent Periphery: A Control Architecture
for Grid 2050'',  {\it IEEE Smart Grid Comm}, Brussels, Belgium,  Oct. 2011.
\item Y. Hu, A. Kuh, T. Yang, A. Kavcic.
``A Belief Propagation Based Power Distribution System State Estimator'', {\it IEEE Computational Intelligence Magazine}, Vol. 6,3, 36--46,  Aug. 2011.
\item
\newblock D. Mandic, S. Javidi, S. L. Goh, A. Kuh and K. Aihara.
\newblock Complex Valued Prediction of Wind Profile Using Augmented
Complex Statistics.
\newblock \emph{Renewable Energy}, 34(1), 196--201, 2009.
\item
\newblock Y. Hu, A. Kuh, A. Kavcic, D. Nakafuji.
\newblock Micro-grid State Estimation using Belief Propagation and Factor Graphs.
\newblock \emph{2010 Asia Pacific Signal and Information Processing Association Summit and Conference,}, Singapore, Dec. 2010.
\item
\newblock A. Kuh and D. Mandic.
\newblock Online Kernel Learning with Applications to Sensor Networks and Wind Prediction
\newblock \emph{2009 International Conference on 
Acoustics, Speech, and Signal Processing}, special session,  3581-3584, Taipei, Taiwan, Apr. 2009.

\end{Publications: Closely Related}


\begin{Publications: Other Significant}
\item
\newblock D.  Mandic, M. Golz, A. Kuh, D. Obradovic, and T. Tanaka , Editors.
\newblock \emph{Signal Processing Techniques for Knowledge Extraction}, Springer, 2008.
\item
\newblock A. Kuh and P. DeWilde,.
\newblock Coments on `Pruning Error Minimization in Least Squares Support Vector Machines'.
\newblock
\emph{IEEE Trans. on Neural Networks}, 18(2): 606-609,  2007.
\item
C.~Zhu, A.~Kuh, J~Wang, and P. DeWilde.
\newblock Analysis of an Evolving Email Network,
\newblock\emph{Physical Review E}. 74, 046109, 13 pages, Oct. 11, 2006.
\item
\newblock M. Gorman et. al.
\newblock Transforming the Engineering Curriculum: Lessons Learned from a Summer at Boeing.
\newblock\emph{Journal of Engineering Education}. 90(1), 143-150, Jan., 2001.
\item
\newblock X. Tian and A. Kuh.
\newblock Performance Bounds for Single Layer Threshold Networks when Tracking a Drifting Adversary.
\newblock \emph{Neural Networks}. 10(5) 897-906, Jul., 1997.
\end{Publications: Other Significant}


\begin{Synergistic Activities}

\item Member, {\it IEEE CIS Smart Grid Task Force}, 2010-.
\item Board of Governors, {\it Asia Pacific Signal and Information Processing Association}, 2009-2012.
\item Editor {\it Electrical and Computer Engineering Department Heads  Association
(ECEDHA)},  Newsletter, 2009-2010.
\item Member of Board of  {\it ECEDHA}, 2009-2012.
\item Technical Co-Chair, {\it International Conf. on Acoustics, Speech, and Signal Proc.}, Honolulu, 2007.
\item Program Committee, {\it IEEE Neural Networks for Signal Processing}, 1998-2001.
\item General Co-Chair {\it International Symp. on Nonlinear Theory and its Applications}, Honolulu, 1993.
\item Information Theory Representative. {\it IEEE Neural Network Administrative Committee}, 1990-1995.
\item Associate Editor, {\it IEEE Transactions on Circuits and Systems}, 1990-1991, 1995-1997, 2000-2001, 2005-2007 (neural networks and signal processing).

\end{Synergistic Activities}


\begin{Collaborators and Co-Editors}
\item A. Kavcic, A. Host Madsen, V. Lubecke, O. Boric-Lubecke, [Univ. Hawaii], D. Nakafuji [Hawaiian Electric Company Inc.], K. Poolla, P.P. Varaiya, F. Wu [UC Berkeley], K. Mani Chandy, 
S. Low [CalTech], D. Bakken, A. Bose, T. Yang [Washington State Univ.], P. Khargonekar 
[Univ. Florida], S. von Meier [Sonoma State], C. Zhu [Juniper Networks], D. Mandic, S. Javidi, J. Wang [Imperial College], S. L. Goh [Shell], K. Aihara [Univ. Tokyo], T. Tanaka [Tokyo Univ. Agriculture and Technology], M. Golz [Univ. of Applied Sciences Schmalkaden, Germany], D. Obradovic [Siemens AG], P. DeWilde [Heriot Watt Univ.].
\end{Collaborators and Co-Editors}


\begin{Graduate Advisors and Postdoctoral Sponsors}
\item Chaopin Zhu [Ph.D., 2007], Ying Hu [M.S. 2011], Nate Kowahl [M.S. 2010], Anand Sharma [M.S. 2008],  M. Sharif Uddin [Ph.D candidate], Navid Tafaghodi, Chris  Manloloyo [M.S. candidate], 16 other graduate students and postdocs.

\end{Graduate Advisors and Postdoctoral Sponsors}

\begin{Thesis Advisor and Postgraduate-Scholar Sponsor}
\item Bradley Dickinson (Princeton University)
\end{Thesis Advisor and Postgraduate-Scholar Sponsor}












