%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% project.workforce.minorities.tex -- 
%% Author          : Philip Johnson
%% Created On      : Fri Jan 13 07:58:21 2012
%% Last Modified By: Philip Johnson
%% Last Modified On: Mon Jan 23 10:03:57 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Education and workforce development}

{\em This project will not only include development of science and
  technology, but also curriculum materials. We will teach a variety of
  courses alongside the development of the microgrid which will enable our
  students to develop skills necessary to deploy these technologies. These
  curriculum materials will be made publically available.

  Contributions of this part of the research will include the development of curriculum
  materials and the creation of engineers and scientists skilled in this
  area.}

\paragraph{Integration of members of under-represented minorities}

This project will collaborate with the Native Hawaiian Science and
Engineering Mentorship Program (NHSEMP) to broaden the participation of
under-represented groups. NHSEMP is a successful program housed at the
University of Hawaii funded (in part) by the National Science Foundation
Louis Stokes Alliance for Minority Participation Program and the
U.S.~Department of Education Native Hawaiian Education Program. The program
is already very successful in attracting members of the Pacific Islander
minority into the undergraduate engineering program. Our next goal is to
achieve a smooth transition of the most talented minority students into the
best graduate programs nationwide. For example, we are already implementing
an REU exchange program with our collaborators (at MIT) in conjunction with
a joint project between MIT and University of Hawaii [NSF Grants
ECCS-0725555 and ECCS-0725649]. Throughout 2008 and 2009, MIT hosted
Hawaiian minority undergraduates from Professor Kavcic's group. Their
video-documented experiences can be viewed at {\bf
  http://www2.hawaii.edu/\verb+~+thanhvu/Videos.html}. We are also
presently in the process of selecting qualified minority undergraduates to
send to Pittsburgh, PA to participate in our joint project with Carnegie
Mellon University [NSF Grants ECCS- ECCS-1029081]. We will continue to draw
representatives of underrepresented into our research program in
conjunction with this project. UH minority undergraduates will spend their
semesters working as researchers with at University of Hawaii and their
summer/winter breaks as interns/researchers at HECO or visiting our
research partners on the mainland.
