\begin{acknowledgments}

Like any research project, this one could not have been completed without the help of a number of people. My thanks go to Prof. Philip Johnson, who has been my wonderful advisor and mentor for over 20 years. Thanks for going ``all in'' for sustainability: there are few advisors that decide to change their entire research program the way you did!
  
I'd like to thank my parents Bill and Ellen Brewer for their support both moral and professional.

I'd also like to thank past and present members of CSDL in chronological order: George Lee, Yongwen Xu, Michelle Kat\-chuck, and Carleton Moore. The Kukui Cup research project would not have happened without your long hours and hard work.

Actually running a Kukui Cup required an even bigger team. I'd like to thank Kaveh Abhari, Hana Bowers, Greg Burgess, Caterina Desiato, Risa Khamsi, Alex Young, and Chris Zorn for their work on the 2011 UH Kukui Cup. University of \Hawaii Student Housing was incredible in their financial and logistical support of the Kukui Cup from: Michael Kaptik, Michael Weakley, Shawn Patrick, Michiko Maggi, Randall Watanabe, and Roland Castillo.

Of course there would be little point in having a Kukui Cup Challenge if nobody participated. I'd like to thank all the residents who played the Kukui Cup; I hope you learned something of value.

I would like to thank my committee members Prof. Martha Crosby, Prof. Scott Robertson, Prof. Daniel Suthers, and Prof. Anthony Kuh for taking the time to read and evaluate this dissertation. In particular, Scott was generous with his time in helping me with questionnaire data analysis.

Several researchers from outside the University of \Hawaii have provided valuable feedback on the Kukui Cup, including Tawanna Dillahunt, Nicole Sintov, Prof. Jon Froehlich, Prof. Eric Paulos, Prof. Sam Joseph, Jim Cummings, and Dante Anderson.

This research is supported in part by grant IIS-1017126 from the National Science Foundation; the HEI Charitable Foundation; the Hawaiian Electric Company; and the State of Hawaii Department of Business, Economic Development and Tourism. We are also thankful for the support from the following organizations at the University of Hawai`i: Facilities Management, and the Department of Information and Computer Sciences. 

Funding from the Center for Renewable Energy and Island Sustainability (REIS) was essential to getting this research project off the ground. REIS has personally supported me with a research assistantship over the course of my work on the Kukui Cup, and provided initial seed funding before we obtained other grants. I've also appreciated the opportunity to interact with other students and faculty from REIS, which has exposed me to other perspectives on this research area.

Last but not least, I thank Yuka Nagashima, the love of my life, for all that she has done in support of me. You were a sounding board for my ideas, you always had insightful comments on the paper drafts you volunteered to read, and you even helped out with Kukui Cup events on your birthday. I could not have done it without you. Mahalo.

\end{acknowledgments}
