\chapter{In-Game Questionnaire}
\label{app:in-game-questionnaire}

This appendix details the contents of the questionnaire made available to participants via the in-game questionnaire during the Overall Round of the challenge.

When subjects filled out the questionnaire via the SurveyGizmo~\cite{surveygizmo} website, the questions were broken into pages. Each page provided subjects the ability to move forward to the next page in the questionnaire, but not back to previous pages. The topic of the pages of the survey were:

\begin{enumerate}
	\item Informed consent via email address
	\item Promotion and Events
	\item Prizes
	\item Adoption
	\item Website
	\item Gamification
	\item General feedback on Kukui Cup
	\item Open feedback on questionnaire
	\item Thank you page
\end{enumerate}

Questions that allowed multiple answer selections (i.e., checkboxes) are shown as ``\Square''. Questions that allowed only a single selection from a list of options (i.e., radio buttons) are shown as ``\Circle''. Questions that provided a text field for open answers do not have any special notation.

The questions were prefaced with the following instructions: ``The goal of this survey is to learn about your experiences during the Kukui Cup. Please answer honestly, we want to know how you really feel.''

\section{Promotion and Events}

Figuring out the best way to promote the challenge and events to residents was one of our major concerns. The goal of this section was to elicit feedback on how to improve our promotion efforts.

\vspace{5 mm}
\noindent
1. Did the banners in the tower lobbies stimulate your curiosity about the Kukui Cup?

\begin{radiobutton}
	\item Yes
	\item No
	\item Don't remember seeing banners
\end{radiobutton}

\vspace{5 mm}
\noindent
2. Did the banner in the Hale Aloha cafeteria during week 1 help you remember about events?

\begin{radiobutton}
	\item Yes
	\item No
	\item Don't remember seeing a banner
\end{radiobutton}

\vspace{5 mm}
\noindent
3. For the workshops and excursions that you attended, how did you like them? You can leave blank any events that haven't taken place by the time you fill out this survey.

[Each event was ranked on the following Likert-type scale:]

\begin{enumerate}
	\item Liked a lot
	\item Liked 
	\item OK 
	\item Boring
	\item Very Boring
	\item Didn't attend
\end{enumerate}

\noindent
Event list:

\begin{itemize}
	\item Kickoff party
	\item Play Outside the Cafe (any)
	\item Energy scavenger hunt
	\item Recycled fashion design
	\item Flashmob planning
	\item Kahuku wind farm
	\item Student Organic Farm Training (SOFT)
	\item Pedalpalooza
	\item Round 1 awards party
	\item Your sustainable future workshop
	\item Energy efficient eating
	\item Movie night
	\item Reppun farm (off-the-grid living)
	\item Kokua Market excursion
	\item Round 2 awards party
	\item High energy art and music
	\item Energy efficient chillaxation
	\item Beach cleanup excursion
\end{itemize}

%\begin{table}[htbp]
%	\centering
%	\begin{tabular}{| l | c | c | c | c | c | c |}
%		\hline
%		Event name & Liked a lot & Liked & OK & Boring & Very Boring & Didn't attend \\ \hline
%Kickoff party & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Play Outside the Cafe (any) & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Energy scavenger hunt & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Recycled fashion design & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Flashmob planning & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Kahuku wind farm & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Student Organic Farm Training (SOFT) & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Pedalpalooza & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Round 1 awards party & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Your sustainable future workshop & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Energy efficient eating & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Movie night & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Reppun farm (off-the-grid living) & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Kokua Market excursion & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Round 2 awards party & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%High energy art and music & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Energy efficient chillaxation & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%Beach cleanup excursion & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  \\
%\hline
%	\end{tabular}
%\end{table}

\vspace{5 mm}
\noindent
4. What can we do to improve attendance at workshops and excursions?

\vspace{5 mm}
\noindent
5. Did you sign up for any events via the website, but then end up not attending?

\begin{radiobutton}
	\item Yes
	\item No
\end{radiobutton}

\noindent
[If No] Why did you end up not attending?


\section{Prizes}

As an incentive to participate in the Kukui Cup, players could win a variety of prizes. In designing the challenge, we did our best to select prizes that would be appealing to participants. In this section we solicited feedback on what prizes participants wanted to see.

\vspace{5 mm}
\noindent
6. What prize (if any) did you find most motivating in the Kukui Cup?

\vspace{5 mm}
\noindent
7. If you had a budget of \$5/person for a prize to be given to everybody in a lounge, what would you want?

\vspace{5 mm}
\noindent
8. If you had a budget of \$10/person for a prize to be given to everybody in a lounge, what would you want?


\section{Adoption}

Understanding how residents find out about the Kukui Cup, and what motivated them to play could be helpful in designing future challenges. The questions in this section were written by Michelle Katchuck to gather pilot data for a potential dissertation proposal topic.

\vspace{5 mm}
\noindent
9. How did you first hear about the Kukui Cup? (choose all that apply)

\begin{checkbox}
	\item I saw a Kukui Cup banner
	\item I read about it in an email I received
	\item I attended one of the courtyard events
	\item A friend referred me
	\item My RA
	\item Other
	\item Yes
	\item No
\end{checkbox}

\vspace{5 mm}
\noindent
10. Prior to playing the Kukui Cup, were you interested in energy conservation?

\vspace{5 mm}
\noindent
11. Has the Kukui Cup increased your interest in energy conservation and sustainability?

\vspace{5 mm}
\noindent
12. Which of the following motivated you the MOST to keep playing the Kukui Cup?

\begin{radiobutton}
	\item to win prizes
	\item to be on the top of the scoreboard
	\item to play with my friends
	\item to learn more about energy and how to save energy
	\item attending events and excursions
	\item other: \underline{\hspace{4cm}}
\end{radiobutton}

\vspace{5 mm}
\noindent
13. What motivated you to participate in the online portion of the Kukui Cup?

\vspace{5 mm}
\noindent
14. What motivated you to participate in the ``real'' world activities of the Kukui Cup?

\vspace{5 mm}
\noindent
15. What has been the most fun and/or interesting activity of the Kukui Cup?

\vspace{5 mm}
\noindent
16. How would you like to hear about Kukui Cup events? (i.e. text message, email, RAs, posters, Facebook, Twitter, the website, etc)


\section{Website}

The 2011 UH Kukui Cup was the first time the challenge website had been used with a large number of participants. The questions in this section focus on the players' experiences with the website. George Lee wrote the questions in this section to collect data for his masters thesis~\cite{csdl2-11-01}.

\vspace{5 mm}
\noindent
17. What did you like about the website?

\vspace{5 mm}
\noindent
18. What did you find confusing about the website?

\vspace{5 mm}
\noindent
19. If you could add or change something in the website, what would that be?

\vspace{5 mm}
\noindent
20. Rate how much you agree with each statement below

[Each statement was ranked on the following Likert-type scale:]

\begin{enumerate}
	\item Strongly disagree
	\item Disagree
	\item Neutral
	\item Agree
	\item Strongly agree
	\item Not Applicable
\end{enumerate}

\noindent
Statements:

\begin{itemize}
	\item It was easy to find what I was looking for in the website.
	\item The website was responsive. I did not wait too long after I clicked on something.
	\item The website provided adequate help in teaching me how to play the game.
	\item I understood the rules of the game and how to play.
\end{itemize}

%\begin{table}[htbp]
%	\centering
%	\begin{tabular}{| l | c | c | c | c | c | c |}
%		\hline
%		Statement & Strongly disagree & Disagree & Neutral & Agree & Strongly agree & Not Applicable \\ \hline
%It was easy to find what I was looking for in the website. & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%The website was responsive. I did not wait too long after I clicked on something. & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%The website provided adequate help in teaching me how to play the game. & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%I understood the rules of the game and how to play. & \Circle  & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%\hline
%	\end{tabular}
%\end{table}

\vspace{5 mm}
\noindent
21. Have you accessed the Kukui Cup website from a smartphone?

\begin{radiobutton}
	\item Yes
	\item No
\end{radiobutton}

\noindent
[If Yes] How satisfied were you with the mobile website?

\begin{radiobutton}
	\item Very Satisfied
	\item Satisfied
	\item Neutral
	\item Dissatisfied
	\item Very Dissatisfied
	\item Not Applicable
\end{radiobutton}


\section{Gamification}

This section primarily covers questions about the game aspects of the challenge. Many of the questions in this section were written by Yongwen Xu to gather pilot data for a potential dissertation proposal topic.

\vspace{5 mm}
\noindent
22. Have you made any commitments through the website during the game?

\begin{radiobutton}
	\item Yes
	\item No
\end{radiobutton}

\noindent
[If Yes] Did you change your behavior during the competition based on the commitment(s) you made?

\begin{radiobutton}
	\item Yes
	\item No
	\item Not sure
\end{radiobutton}

\vspace{5 mm}
\noindent
23. Which of the followings Kukui Cup achievements would you want to share on Facebook? (choose all that apply)

\begin{checkbox}
	\item made a commitment
	\item participated in an activity
	\item attended an event or excursion
	\item earned a badge
	\item current leader in the scoreboard
	\item other
\end{checkbox}

\vspace{5 mm}
\noindent
24. How much time do you usually spend on the following activities?

[Options for each activity:]

\begin{enumerate}
	\item 3 or more hours a day
	\item about 1 hour a day
	\item about 1 hour a week
	\item 1 hour a month or less
	\item never
\end{enumerate}

\noindent
List of activities:

\begin{itemize}
	\item Playing games on a laptop computer
	\item Playing games on a game console (Xbox, PS3, Wii)
	\item Playing games on a handheld game device (DS3, PSP)
	\item Playing games on a mobile phone
	\item Checking Facebook
	\item Checking Twitter
\end{itemize}

%\begin{table}[htbp]
%	\centering
%	\begin{tabular}{| l | c | c | c | c | c |}
%		\hline
%		Activity & 3 or more hours a day & about 1 hour a day & about 1 hour a week & 1 hour a month or less & never \\ \hline
%Playing games on a laptop computer & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%Playing games on a game console (Xbox, PS3, Wii) & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%Playing games on a handheld game device (DS3, PSP) & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%Playing games on a mobile phone & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%Checking Facebook & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%Checking Twitter & \Circle  & \Circle  & \Circle  & \Circle  & \Circle \\
%\hline
%	\end{tabular}
%\end{table}

\vspace{5 mm}
\noindent
25. How would you describe the Kukui Cup? (check all that apply)

\begin{checkbox}
	\item Fun
	\item Educational
	\item So-so
	\item Boring
	\item Not useful
	\item Difficult
	\item Addictive
	\item Other
\end{checkbox}

\vspace{5 mm}
\noindent
26. The Kukui Cup website shows energy data updated every 15 seconds. Did you find this helpful in conserving energy?

\begin{radiobutton}
	\item not really, updating the data daily would be enough
	\item not really, updating the data hourly would be enough
	\item not really, I only care about the final result of the competition
	\item yes, it is helpful to see the energy usage changing in real time
\end{radiobutton}

\vspace{5 mm}
\noindent
27. Which of the following do you wish there were more of in the game? (choose all that apply)

\begin{checkbox}
	\item events
	\item excursions
	\item commitments
	\item videos
	\item social activities
	\item physical activities
	\item online activities
\end{checkbox}

\vspace{5 mm}
\noindent
28. On average, how many minutes a day did you spend on the Kukui Cup website?

\vspace{5 mm}
\noindent
29. On average, how many hours a week did you spend at Kukui Cup events?


\section{General Feedback}

This section covers feedback on the challenge overall.

\vspace{5 mm}
\noindent
30. What can we do to improve participation in the Kukui Cup website?

\vspace{5 mm}
\noindent
31. What was the best thing you liked about the Kukui Cup so far?

\vspace{5 mm}
\noindent
32. What was the thing you liked the least about the Kukui Cup so far?

\vspace{5 mm}
\noindent
33. If you were able to play the Kukui Cup next year, would you?

\begin{radiobutton}
	\item Yes
	\item I enjoyed it, but I wouldn't play again
	\item I didn't enjoy it, and I wouldn't play again
	\item No, because: \underline{\hspace{5cm}}
\end{radiobutton}

\vspace{5 mm}
\noindent
34. How likely would you be to recommend playing the Kukui Cup to a first year student in Fall 2012?

\begin{radiobutton}
	\item Very Likely
	\item Likely
	\item Neutral
	\item Unlikely
	\item Very Unlikely
	\item Not Applicable
\end{radiobutton}

\vspace{5 mm}
\noindent
35. Is there anything else you would like to tell us about your experience playing the Kukui Cup that this survey didn't ask?
