%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% costs.tex -- 
%% Author          : Philip M. Johnson
%% Created On      : Tue Nov 11 13:55:32 2003
%% Last Modified By: Philip M. Johnson
%% Last Modified On: Tue Nov 18 10:50:47 2003
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2003 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{Proposed Costs}

\subsection{Cost categories and amounts}

\begin{tabular}{|l|r|r|r|r|} \hline
{\bf Cost Category} & {\bf 2004} & {\bf 2005} & {\bf 2006} & {\bf Category Total} \\ \hline
Salaries (Senior Personnel)      & \$9,609 & \$9,609 & \$9,609 & \$28,827 \\ \hline
Salaries (Graduate Assistants)  & \$33,648 & \$33,648 & \$33,648 & \$100,944  \\ \hline
Fringe Benefits       & \$8,748 & \$8,748 & \$8,748 & \$26,244 \\ \hline
Travel                & \$4,000  & \$4,000  & \$4,000  & \$12,000 \\ \hline
Materials and Supplies & \$2,500   & \$2,500  & \$2,500  & \$7,500  \\ \hline
Indirect Costs         & \$21,237 & \$21,237 & \$21,237 & \$63,711  \\ \hline \hline
{\bf Yearly Total}            & \$79,741  & \$79,741  & \$79,741  & \$239,223 \\ \hline
\end{tabular}

\subsection{Cost category descriptions}

The following cost category descriptions are based upon co-funding
for the three years of this project from the National Science
Foundation under the Highly Dependable Computing and Communication Systems
program. Thus, the amount requested in this proposal for Graduate
Assistants is actually only half of the amount required to fully fund these
four positions during the project.

\subsubsection*{Salaries}

The project budget provides salary support for the principal investigator
(two summer months) and four graduate research assistants (11 months) for
each of the four years.  The principal investigator will perform or
supervise all major system design enhancements and empirical studies,
manage and train the graduate student assistants, and supervise activities
at JPL.

\subsubsection*{Travel}

The project budget provides funds for four trips per year from Hawaii to
the mainland.  These funds will be used by the principal investigator to
meet with JPL collaborators and/or attend relevant conferences (e.g., ICSE,
FSE, ISERN, or STAR), where he may report on the results of his own work
and obtain first hand information on related work. 

\subsubsection*{Materials and Supplies}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise SUN server and 8
workstations running Windows2000, along with a printer and 
networking peripherals.  The project budget provides for materials and services related
to this equipment.

\subsubsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
Summer salary is calculated as two additional months of the principal
investigator's annual nine month salary.   

\paragraph*{Research Assistants.}  
This cost category provides support for four graduate research assistants
for each of the four years.  

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 3.5\% for the principal
investigator and 25\% for the graduate assistants.

\paragraph*{Indirect costs.}  
University of Hawaii indirect (overhead) cost is calculated as 36.3\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).








