\begin{abstract}
Due to rising costs and the questionable future of our non-renewable energy reserves, individuals need to become aware of their energy usage. In order to instill these habits earlier, organizations have held energy competitions to promote the reduction of energy. This also has the side effect of reducing the energy cost to the organization holding the competition. Typically, these competitions are held in colleges and universities and there are companies that can provide hardware and software to support them. However, since such solutions can be expensive, we would like a free, open source solution that can be used by any organization.

We created Makahiki to be an open source framework for sustainability competitions. We also designed it to be a platform for researchers to investigate user behaviors during an energy competition. However, in order to validate our design, we need to evaluate and test our design. During the course of development, we had three evaluation phases. In the mockup phase, we validated our design before doing any implementation. In the onboarding phase, we investigated how individuals will interact with the system when they visit it for the first time. Finally, in the beta phase, we simulated the competition on a much smaller scale in order to observe how Makahiki might be used in an actual competition. Following these evaluations, Makahiki was used to support the 2011 Kukui Cup, which was held in mid-October.

In summary, we claim the following contributions:

\begin{enumerate}
  \item An open source system for creating serious games for energy competitions.
  \item A research platform on which researchers can observe user behavior during energy competitions.
	\item A methodology for evaluating and testing serious games that involve competitions over a period of time.
\end{enumerate}
\end{abstract}