%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 1-intro.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Apr  8 14:22:56 1998
%% Last Modified By: Philip Johnson
%% Last Modified On: Tue Feb  2 08:10:27 1999
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Philip Johnson
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{Introduction}
\label{sec:introduction}

\begin{quotation}
\noindent {\em The actual process is what you do, with all its omissions, mistakes, and
oversights. The official process is what the book says you are supposed to
do.} \cite{Humphrey95}
\end{quotation}

The Personal Software Process (PSP) was introduced in 1995 in the book, ``A
Discipline for Software Engineering'' \cite{Humphrey95}.  This text
describes a one-semester curriculum for advanced undergraduates or graduate
students in computer science that teaches concepts in empirically-guided
software process improvement. Since its introduction, experience with the
PSP has been reported on in several case studies
\cite{Ceberio-Verghese96,Ferguson97,Humphrey96,Humphrey97}.
Although empirically-guided software process improvement is a key feature
of other software engineering initiatives, such as the Capability Maturity
Model (CMM) \cite{Paulk95}, ISO-9000, and Inspection \cite{Gilb93}, the PSP
differs from these other approaches in important ways.

The CMM, ISO-9000, and Inspection discuss empirical software process
improvement in the context of a large organization.  Process improvement in
this context requires the gathering and analysis of large amounts of data,
within and across departments, generated by different people at different
times.  Indeed, inevitable personnel turnover means that the data collected
from the working procedures of one set of people tend to generate
measurements leading to process changes that affect the working procedures
of a potentially different set of people.  The substantial effort required
to collect, interpret, and introduce organizational change based upon the
measurements for a large organization leads to the need for an explicit
software engineering process group (SEPG) whose mission is to manage
empirically guided improvement. Although the utility of these approaches
have been repeatedly validated, they leave the unfortunate impression that
empirically-guided software process improvement is the sole province of
large organizations who can dedicate teams of people to its enactment.

The PSP provides an alternative, complementary approach in which
empirically guided software process improvement is ``scaled down'' to the
level of an individual developer.  In the PSP, individuals gather
measurements related to their own work products and the process by which
they were developed, and use these measures to drive changes to their
development behavior.  PSP focuses on defect reduction and
estimation accuracy improvement as the two primary goals of personal
process improvement. Through individual collection and analysis of personal
data, the PSP provides a novel example of how empirically-guided software
process improvement can be implemented by individuals regardless of the
surrounding organizational context and the availability of institutional
infrastructure support.

Since the PSP is a new technique, relatively little data exists on its use
and effectiveness.  Case studies typically report positive results, usually
based upon the data collected during enactment of the PSP curriculum.  One
typical case study conclusion is that ``during the course, productivity
improvements average around 20\% and product quality, as measured by
defects, generally improves by five times or more'' \cite{Ferguson97}.
Similarly, another study states that ``the improvement in average defect
levels for engineers who complete the course is 58.0 percent for total defects
per KLOC and 71.9 percent for defects per KLOC found in test''
\cite{Humphrey96}.  Indeed, our own
PSP data yields similarly positive measurements for process and products.

In this paper, we report on a case study performed to assess the quality of
PSP data---the measurements typically used to evaluate the effectiveness of
the PSP as illustrated above.  Our case study was motivated by our
experiences teaching and using the PSP, which led us to suspect that the
empirical measures gathered by the PSP may not, in all cases, reflect the
true underlying process or products of development.

We hypothesized that problems with the quality of process data collected
with the PSP could significantly change at least some of the measures
produced by the PSP that are commonly used to evaluate its effectiveness.
By ``significantly'', we mean something stronger than just a statistically
significant difference between the recorded measurements and the actual
underlying programmer behavior. We mean that the difference between
recorded measures and actual behavior would be sufficient, at least in some
cases, to lead developers to the wrong conclusion about how to improve
their process.

To test this hypothesis, one of us taught a modified version of the PSP
curriculum to a class of 10 students in the Fall of 1996.  The 
course was augmented
with features designed to improve the data quality of the raw
PSP data.  We then entered the PSP measures into a database and subjected
them to a variety of data quality analyses.  These analyses uncovered over
1500 errors in the PSP data used by the students to track their work and
motivate process improvements.  Additional analysis yielded a seven part
classification scheme for PSP data errors.  Although we were not always
able to generate corrected values for the data errors, partial correction
lead to ``significantly'' different values for certain PSP measurements,
confirming our hypothesis.

The remainder of the paper is organized as follows.  Section
\ref{sec:related} presents a description of related work, including an
overview of the PSP itself, case studies of the PSP, and additions and
enhancements to the basic approach.  Section \ref{sec:model} presents a
model of PSP data quality we devised to guide our investigation.  Section
\ref{sec:case-study} presents the case study; its design, instrumentation,
data collection, analysis, and threats to internal and external validity.
Section \ref{sec:results} presents the quantitative and qualitative results
we obtained from the study.  Section \ref{sec:discussion} presents our
interpretation of these results.  This section also introduces the concept
of ``measurement dysfunction'', which is important to our interpretation of
the results from this study and our recommendations for future research and
practice.









