\documentstyle[11pt,times,/export/home/csdl/tex/definemargins]{article}
\input{/export/home/csdl/tex/psfig/psfig}

\definemargins{1.0in}{1.0in}{1.0in}{1.0in}{0.3in}{0.3in}

\begin{document}

\title{A Critical Analysis of PSP Data Quality: Results from a Case Study}


\author{
        Philip M. Johnson\\
        Anne M. Disney\\
        Dept. of Information and Computer Sciences\\
        University of Hawaii\\
        Honolulu, HI  96822 USA\\
        johnson@hawaii.edu\\
        adisney@ilhawaii.net\\
       }

\maketitle

\begin{abstract}
  
  The Personal Software Process (PSP) is used by software engineers to
  gather and analyze data about their work.  Published studies typically
  use data collected using the PSP to draw quantitative conclusions about
  its impact upon programmer behavior and product quality.  However, our
  experience using PSP led us to question the quality of data both during
  collection and its later analysis.  We hypothesized that data quality
  problems can make a significant impact upon the value of PSP
  measures---significant enough to lead to incorrect conclusions regarding
  process improvement.  To test this hypothesis, we built a tool to
  automate the PSP and then examined 89 projects completed by ten subjects
  using the PSP manually in an educational setting.  We discovered 1539
  primary errors and categorized them by type, subtype, severity, and age.
  To examine the collection problem we looked at the 90 errors that
  represented impossible combinations of data and at other less concrete
  anomalies in Time Recording Logs and Defect Recording Logs.  To examine
  the analysis problem we developed a rule set, corrected the errors as far
  as possible, and compared the original and corrected data.  We found
  significant differences for measures such as yield and the
  cost-performance ratio, confirming our hypothesis.  Our results raise
  questions about the accuracy of manually collected and analyzed PSP data,
  indicate that integrated tool support may be required for high quality
  PSP data analysis, and suggest that external measures should be used when
  attempting to evaluate the impact of the PSP upon programmer behavior and
  product quality.

\end{abstract}

\subsection*{Keywords}

\noindent Personal software process, defects, 
empirical software engineering,
measurement dysfunction, automated process support\\

\newpage
\tableofcontents
\newpage

\input{1-intro}
\input{2-related}
\input{3-model} 
\input{4-case-study} 
\input{5-results}
\input{6-discussion}    

\section{Acknowledgments}

We gratefully acknowledge all of the students in all of the PSP classes at
the University of Hawaii.  Our colleagues in the Collaborative Software
Development Laboratory during the time of this research (Cam Moore, Robert
Brewer, Jennifer Geis, Joe Dane, and Russ Tokuyama) provided ongoing
support.  We would like to thank Watts Humphrey, James Over, and Will Hayes
of the Software Engineering Institute, and the anonymous reviewers, whose
comments sharpened the presentation of this research.  This research was
sponsored in part by grants CCR-9403475 and CCR-9804010 from the National
Science Foundation.


\bibliographystyle{plain}
\bibliography{/group/csdl/bib/psp,/group/csdl/bib/ftr} 

\end{document}



