%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% nsf93-bio.tex -- 
%% RCS:            : $Id: nsf93-bio.tex,v 1.2 93/09/01 16:09:06 johnson Exp $
%% Author          : Philip Johnson
%% Created On      : Wed Aug 11 14:09:06 1993
%% Last Modified By: Philip Johnson
%% Last Modified On: Mon Aug 30 17:15:28 1993
%% Status          : Unknown
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1993 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%% History
%% 11-Aug-1993		Philip Johnson	
%%    


%%	Command to define new categories:
\newcommand{\newcategory}[1]{\newenvironment{#1}
 {\sectionheading{#1}\begin{description}}{\end{description}\par}}
\newcommand{\sectionheading}[1]{\medskip\pagebreak[2]\par\noindent
 {\small\bf #1}\nopagebreak}

\newcategory{Degrees}
\newcategory{Five Closely Related Publications and Systems}
\newcategory{Five Other Significant Publications}




\section{Bibliographical Sketch}


\begin{center}
{\bf Philip M. Johnson}\\
Department of Information and Computer Sciences\\
University of Hawaii at Manoa\\
Honolulu, HI~~~96822\\
(808) 956-3489\\
(808) 956-3548 (fax)\\
Email: {\tt johnson@hawaii.edu}
\end{center}

\begin{Degrees} 
\item Ph.D.~in Computer Science, University of Massachusetts at Amherst, 1990. 
\item M.S.~in Computer Science, University of Massachusetts at Amherst, 1985.
\item B.S.~in Computer Science, B.S.~in Biology, University of Michigan, 1980.  
\end{Degrees}

\subsection*{Publications}

\begin{Five Closely Related Publications and Systems}

\item P.M.~Johnson, {\sl Experiences with EGRET: An Exploratory Group Work Environment},
Journal of Collaborative Computing, in submission.

\item P.M.~Johnson, D.~Tjahjono, D.~Wan, R.~Brewer, {\sl Experiences with CSRS: An Instrumented Software Review Environment}.
To appear in Proceedings of the Pacific Northwest Software Quality Conference, Portland, OR. October, 1993.

\item P.M.~Johnson, D.~Tjahjono, {\sl Experiences with CSRS: An Instrumented Software Review Environment}.
To appear in Proceedings of the Third European Conference on Computer Supported Cooperative
Work, Milan, Italy, September 1993.

\item P.M.~Johnson, {\sl Supporting Exploratory CSCW with the EGRET Framework},
in Proceedings of the ACM 1992 Conference on Computer Supported Cooperative
Work, Toronto, Canada. 1992.

\item {\em CSRS: A Collaborative Software Review System}.  CSRS is a 37 KLOC client-server, 
multi-user hypertext system for software review written in C++ and Lisp.  

\end{Five Closely Related Publications and Systems}

\begin{Five Other Significant Publications}

\item P.M.~Johnson, {\sl Type Flow Analysis for Exploratory Software Development}, 
Ph.D. Thesis, University of Massachusetts, 1990.

\item P.M.~Johnson, {\sl Structural Evolution in Exploratory Software Development},
in Proceedings of the 1989 AAAI Spring Symposium on AI and Software Engineering,
Stanford, CA, 1989.

\item P.M.~Johnson, D.~Hildum, A.~Kaplan, C.~Kay, and J.~Wileden, {\sl An Ada
Restructuring Assistant}, in Proceedings of the Fourth Annual Conference on Artificial
Intelligence and Ada, Fairfax, VA, 1988.

\item D.~Corkill, K.~Gallagher, and P.~Johnson, {\sl Achieving Flexibility, Efficiency, 
and Generality in Blackboard Architectures},in Proceedings of the Sixth National Conference
on Artificial Intelligence (AAAI-87), Seattle, WA., 1987.  Also appearing in {\em Readings
in Distributed Artificial Intelligence}, Alan Bond and Les Gasser, editors, Morgan-Kaufman, 1988.

\item P.~Johnson and W.~Lehnert, {\em Beyond Exploratory Programming: A Methodology and 
Environment for Natural Language Processing}, in Proceedings of the Fifth National Conference
on Artificial Intelligence (AAAI-86), Philadelphia, PA, 1986.  Also appearing in 
{\em Artificial Intelligence and Software Engineering}, D. Partridge, editor, Ablex 1990.

\end{Five Other Significant Publications}


\subsection*{Other Collaborators}

None.

\subsection*{Graduate Advisors}

Victor R. Lesser, Jack C. Wileden, Wendy G. Lehnert, Daniel D. Corkill. 

