%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% uhtest-appendix.tex -- 
%% Author          : Robert Brewer
%% Created On      : Fri Oct  2 16:31:12 1998
%% Last Modified By: Robert Brewer
%% Last Modified On: Mon Oct  5 14:41:05 1998
%% RCS: $Id: uhtest-appendix.tex,v 1.1 1998/10/06 02:07:03 rbrewer Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\documentclass[11pt]{article}
%%% Load some useful packages:

\usepackage{times}

% make margins smaller
\usepackage[left=2.5cm,top=2cm,right=2.5cm,bottom=2cm,nohead,nofoot]{geometry}

%% Provides customization of lists
\usepackage{enumitem}

%% Now define multiple choice answer list type
\newlist{answer}{enumerate}{1}
\setlist[answer]{label=\alph*)}

%%% End of preamble
\begin{document}

\title{Energy Literacy Quiz}
%\author{Robert S. Brewer}
\date{March 19, 2010}

\maketitle

%%% Body text goes here

\section{Quiz}

\subsection{The watt is a unit of:}

\begin{answer}
	\item energy
	\item power
	\item distance
	\item force
\end{answer}

\subsection{Power is commonly measured in units of:}

\begin{answer}
	\item BTU
	\item joule
	\item kilowatt-hour
	\item watt
\end{answer}

\subsection{The watt is abbreviated as:}

\begin{answer}
	\item wt
	\item Wh
	\item W
	\item tt
\end{answer}

\subsection{The watt-hour is a unit of:}

\begin{answer}
	\item energy
	\item power
	\item distance
	\item force
\end{answer}

\subsection{Electrical energy is commonly measured in units of:}

\begin{answer}
	\item BTU
	\item joule
	\item watt-hour
	\item watt
\end{answer}

\subsection{The watt-hour is abbreviated as:}

\begin{answer}
	\item Wh
	\item wth
	\item W
	\item erg
\end{answer}

\subsection{A compact fluorescent lightbulb uses 13 W. If it is run for 2 hours, how much energy does it use?}

\begin{answer}
	\item 7.5 Wh
	\item 13 Wh
	\item 26 Wh
	\item 52 Wh
\end{answer}

\subsection{A compact fluorescent lightbulb (CFL) used 26 Wh after running for 2 hours. How much power does the bulb consume?}

\begin{answer}
	\item 7.5 W
	\item 13 W
	\item 26 W
	\item 52 W
\end{answer}

\subsection{Roughly how much power does a normal compact fluorescent lightbulb (CFL) use when running?}

\begin{answer}
	\item 20 mW
	\item 3 W
	\item 60 W
	\item 13 W
\end{answer}

\subsection{Roughly how much power does an electic oven use when turned to its highest setting?}

\begin{answer}
	\item 100 W
	\item 500 W
	\item 1000 W
	\item 4000 W
\end{answer}

\subsection{What is the approximate maximum power generated from a single standard rooftop solar panel?}

\begin{answer}
	\item 25 W
	\item 50 W
	\item 200 W
	\item 800 W
\end{answer}

\subsection{What is the approximate maximum electrical power demand for the entire island of Oahu?}

\begin{answer}
	\item 560 kW
	\item 3800 kW
	\item 11.8 GW
	\item 1.2 GW
\end{answer}

\subsection{On average, how much electrical energy does a home in Hawaii use per day?}

\begin{answer}
	\item 13 kWh
	\item 4 kWh
	\item 57 kWh
	\item 328 kWh
\end{answer}

\subsection{If your game console uses 100 W when turned on, how much energy would it waste if you left it on all weekend while you were away?}

\begin{answer}
	\item 15000 Wh
	\item 100 Wh
	\item 480 kWh
	\item 4.8 kWh
\end{answer}

\end{document}
