%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% budget.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Jul 27 13:31:23 1994
%% Last Modified By: Philip Johnson
%% Last Modified On: Fri Apr 03 14:58:31 2009
%% Status          : Unknown
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 


\section*{Budget Justification}

\pagenumbering{arabic}
\renewcommand{\thepage} {F--\arabic{page}}

\subsection*{Salaries}

The project budget provides salary support for the principal investigator
(two summer months) and four graduate research
assistants (11 months) for each of the four years.  The principal
investigator will perform or supervise all major system design enhancements
and empirical studies, manage and train the graduate student
assistants, and supervise activities at JPL.

\subsection*{Travel}

The project budget provides funds for four trips per year from Hawaii to the
mainland.  These funds will be used by the principal investigator to attend
two relevant conferences (e.g., ICSE, FSE, ISERN, or STAR) each year, where
he may report on the results of his own work and obtain first hand
information on related work. The remaining two trips per year will be to JPL. 

\subsection*{Computer Services}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise SUN server and 8
workstations running Windows2000, along with a printer and 
networking peripherals.  The project budget provides funds of \$5,000
per year for each year of the grant to support computer services related
to this equipment.

\subsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
Summer salary is calculated as two additional months of the principal
investigator's annual nine month salary.   

\paragraph*{Research Assistants.}  
This cost category provides support for four graduate research assistants
for each of the four years.  The salary for each research assistant is
based upon RA-Step 3.

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 3.5\% for the principal
investigator and 25\% for the graduate assistants.

\paragraph*{Overhead.}  
University of Hawaii overhead cost is calculated as 36.3\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).





