%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Plain-Tex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%
%% thesis-acknowledgment.tex -- 
%% Author          : Dadong Wan
%% Created On      : Thu Apr  7 01:04:21 1994
%% Last Modified By: Jennifer Geis
%% Last Modified On: Tue Apr  7 15:30:21 1998
%% Status          : Unknown
%% RCS: $Id: acknowledgments.tex,v 1.5 94/04/28 15:52:00 dxw Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgements{

Some people may think that the writing of a thesis is a solitary journey.
It is, in the sense that you are stuck doing the work, but without the
support of others, I imagine that few would ever finish the job.  I am very
fortunate, for I have many people who have helped me along the way.

To Philip Johnson, my advisor.  I can't begin to recall all the times you
acted as my cheering section, urging me on to accomplish things I never
would have dreamed of just a few years ago.  I was muddling through the
undergraduate program when you held out the prospect of joining CSDL.  With 
some trepidation, I accepted, not having a clue what was in store for me.
You have changed the way I view education, software development, and my
future.  I cannot thank you enough. 

To all past and present members of CSDL.  You have critiqued my work and
suggested many improvements which I would not have thought of myself.  This 
thesis has been greatly enhanced as a result of your willingness to give of
your time and experience.  

To the person who started me on JWiz, Bruce Foster.  Who would of thought
that a summer in New Hampshire would lead to this?  I thank you for taking
the gamble of having me as an intern, and for offering me the opportunity
to continue my efforts back in Hawaii.  

To Dale Skrien, Jeremy Lueck, and Mitchell Goodman.  You provided me with
the underlying functionality on which I run JWiz.  Thank you for
considering my suggestions and requests for your system.  

To Bill McKeeman, Will Kling, Steve Rodgers, and all the people at Digital
who offered their ideas and support.

%To Michael O'Sullivan.  You were patient and understanding.  You took my
%complaints and frustration in stride and offered a sympathetic ear.  I am
%so lucky to have you in my life.

I would like to thank my mother and father for attempting to look
interested when I talked about JWiz, sympathizing with me when I was
worried, and making an effort to understand my topic.  I realize much of
what I said probably had little meaning to you, but I appreciate your
letting me use you as sounding boards.  I would also like to thank you both
for letting me set my own path through life and supporting the choices I
make. For these reasons, this thesis is for you.  

% people to thank: Mom, Dad, Mike O, Johnson, all members of CSDL, Digital, 
% Bruce Foster, Will Kling, Steve Rodgers, Bill McKeeman, Dale Skrien, 
% Jeremy Lueck, Mitchell Goodman, Dr. Martha Crosby, Dr. Jay Corbett,
% Feng Gao 
}











