\section{Discrete Fourier Transform (DFT)}
Agrawal et al. in \cite{citeulike:3973409} proposed a DFT-based method for the time-series indexing and similarity queries processing. Their idea based on the observation of the fairly good approximation of the time-series by only few ``strong'' frequencies and on the Parseval's Theorem (aka Rayleigh's energy theorem). 

For each of the time-series $x$ of length $N$ authors proposing to extract only $c$ first DFT coefficients (frequencies), where $c<<N$ and $f_{c}$ is the ``cutoff frequency''. Thus each of the time-series is mapped into the low $c$ dimensional space and stored in the $F$-index, whether the search over the index implemented by using the $R^{*}$ tree \cite{citeulike:343069}. Authors arguing that the approach taken is characterized by the ``completeness of the feature-extraction'' and it is ``efficiently dealing with the dimensionality curse''. 

The $n$-point DFT transform of the time-series $X=(x_{0}, x_{1}, ... , x_{n-1})$ is defined as:
\begin{align}
& X_{f} = 1/\sqrt{(n)}\sum_{t=0}^{n-1} x_{t} \exp(-j2 \pi f t/n),\; t=0,1,...,n-1, \; j=\sqrt{(-1)} \\
& \text{also the inverse transform:} \nonumber \\
& x_{t} = 1/\sqrt{(n)}\sum_{f=0}^{n-1} X_{f} \exp(j2 \pi f t/n),\; t=0,1,...,n-1 
\end{align}
and the fundamental observation of the Paseval's theorem is
\begin{equation}
\sum_{t=0}^{n-1} \left| x_{t} \right| ^{2} = \sum_{f=0}^{n-1} \left| X_{f} \right| ^{2}
\label{eq:parseval}
\end{equation}
that the energy of the sequence in the time domain equals the energy in the frequency domain.

Narrowing further, authors show that Discrete Fourier Transform inherits linearity and preservance of amplitude coefficients during the time-shifts from the Continuous Fourier Transform. Taking this properties and Parseval's theorem in account authors stating that
\begin{equation}
\left\| x - y \right\| ^{2} \equiv \left\| X - Y \right\| ^{2}
\label{eq:dft_similarity}
\end{equation}
which implies the equivalence of the Euclidean distances between two sequences $x$ and $y$ in the time and frequency domains. 

Agrawal et al. introducing the distance function $D$ where distance between two sequences $x$ and $y$ is the square root of the energy of the difference:
\begin{align}
D(x,y) \equiv \left( \sum_{t=0}^{n-1} \left| x_{t} - y_{t} \right| ^{2} \right) ^{1/2}
       \equiv \left( E(x-y) \right) ^{1/2} \\
\text{where } E(x) \; \text{is the energy of the sequence:} \nonumber \\
E(x) \equiv \left\| x \right\| ^{2} \equiv \sum_{t=0}^{n-1} \left| x_{t} ^{2} \right|
\label{eq:dft_distance}
\end{align}
and implementing the similarity relation between two sequences by using a user defined threshold $\epsilon$ - i.e. if distance between two sequences falls below threshold they are similar:
\begin{equation}
D(x,y) \leq \epsilon \; \Rightarrow \text{$x$ similar to $y$}
\label{eq:dft_similarity_dft}
\end{equation}

Continuing the discussion authors comparing their DFT-based method to the ``sequential scanning'' method. They used the same $R^{*}$ tree for the sequential indexing and implemented ``early abandoning'' as soon as the calculated Euclidean distance exceeds $\epsilon^{2}$ declaring two sequence dissimilar. It was shown by comparing two approaches that with the growth of the dataset and the length of the sequences the DFT-based approach outperforms the sequential search. Another interesting point shown is that in general $f_{c} \approx 3$ is enough to capture time-series features and build index which provides satisfactory performance and has a contractive property, i.e. ensures no false-dismissal.
