%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 03-05-abstract.tex -- 
%% Author          : Joy M. Agustin
%% Created On      : Fri Jun  9 09:43:42 1995
%% Last Modified By: Aaron Kagawa
%% Last Modified On: Wed Nov 19 14:01:26 2003
%% Status          : Unknown
%% RCS: $Id: thesis-abstract.tex,v 1.1 1998/09/19 01:24:42 jagustin Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1995 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 


\begin{abstract}

Managing large software projects is intrinsically difficult.  Although, high
software quality is a definite must, other issues like time and cost play
major roles in large software development.  For example, if a software
company can produce the highest quality products but cannot predict how
long and how much it is going to cost, that company will not have any
business.

For large scale programming projects is it possible to use the system's
build process to understand, predict, and prevent problems in quality and
productivity?  I will address this research question through the
development of specific measurements and analyses of a large scale software
project: the Mission Data System (MDS) at the Jet Propulsion Laboratory.  I
have created the Hackystat Jet Propulsion Laboratory Build (hackyJPLBuild)
system which contains an initial set of measurements and analyses of the
build process of MDS.

Using the hackyJPLBuild system, I will address the problem of predicting
software problems in the Mission Data System.  To accomplish this I will
conduct three case studies: (1) can the hackyJPLBuild accurately represent
the build process of MDS, (2) can threshold values indicate problematic
issues in the MDS system, and (3) can the hackyJPLBuild predict future
problematic issues in the MDS system.

Initial results of case study 1 indicate that hackyJPLBuild can accurately
represent the build process of MDS.  In fact, the hackyJPLBuild system has
already identified some potential flaws in the MDS build process.  Case
studies 2 and 3 have not been conducted yet.

If this thesis project is successful, I will be able to identify the types
of build process metrics that are applicable to identify, predict, and
prevent software quality and productivity problems.

\end{abstract}
