

\chapter {Questionnaire For Individuals (EIAM) --ICS411}

\begin{enumerate}

\item  My understanding of the source code before the review was: 
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.1)

\item  My understanding of the source code after the review was: 
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill
Very high\\
(Score: 3.6)

\item  My understanding of C programming language was improved after
this review.  
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.1)

\item  In general, I found it easy to understand the logic of
the code. 
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.2)

\item  For this review, I would have preferred working alone rather 
than working in a group (if I have had a choice).
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 2.4)

\item  For this review, I felt more confidence about the issues I raised when 
working alone rather than working in a group.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.0)

\item  The training session for individual review was sufficient.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.0)

\item   There was sufficient time to work on this review.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.1)

\item   I was motivated to do this review project.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.6)

\item  My overall confidence in the quality of my review was
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.3)

\item  I felt comfortable doing this review.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.5)

\item  It took me a good while to decide whether a particular program
segment contained valid issues.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.7)

\item  I often wanted to change or delete issues I just created.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 2.7)

\item  My overall satisfaction with individual review (EIAM) process as
outlined in the FTR guideline was:
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.8)

\item  I believe EIAM system made my review more productive (i.e., find
lots of issues) 
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.4)

\item  I believe EIAM system made my review more effective
(i.e., find lots of ``good'' issues in a relatively short time)
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.3)

\item  EIAM system is easy to use.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.3)

\item  EIAM system is useful. 
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.8)

\item  My overall satisfaction with EIAM system was:
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.9)

\item  Problems that I had with EIAM system (please explain): 
\begin{itemize}
\item I wanted to search but I didn't know what key combinations to do.
Maybe a menu item for that would be nice.
\item none
\item Was that there were no manuals to consult in aiding the review
process.  It is often difficult to memorize an entire C/C++
reference manual, and without the luxury of knowing the full extent of
the C/C++ language, the review process was affected.  The entire
purpose of the review process, as I see it, was to raise certain
issues concerning the program.  This I feel was more or less
accomplished, though not very precisely.

When writing a program, programmers often refer to programming manuals
for language syntax and command usage.  During the review process,
this was a luxury that weren't available.  Without knowing the
language, the issues that arose may not be correct simply because the
reviewer did not understand what the command in the procedure was for
or if it was the proper syntax.  This is unfortunate and affects the
reliability of the review process.


\item Trying to remember which mouse button to press.
\item  not familiar to use at the first time, after several trials, it's
 wonderful!

\item Not familiar with openwin...
\item I never really had any problems with the EIAM system, since it was
pretty general, and also it was quite efficient to make me work
faster.

\item Looking at the screen for long periods of time.
\item It takes too long to do this experiment.  The functions are not in any
logical orders which makes it very diffcult to view.  Maybe we should
have a printout of all the codes went we doing this experiment. 
\item The problems with this  system was the environemnt
\item controlling the environment
pointing here and there

\item     I don't debug programs entirely from memory without doing test
compiles to find the errors.  I open my K \& R C book and compile the
code to catch all the obvious errors, such as spelling, forgotten
semi-colons, etc.  The "true" programming way may be to understand the
code 100% without compiling like they did in the olden days, but I
cannot adapt to that system.  I know that if you took away my C book
and compiler, I would NEVER be able to turn in any working programs longer
than "hello world."  
    As such, I think that this process could be much improved
if it could compile code ( just to find obvious spelling and structure
errors ).  Doesn't it make sense to use a spell checker on a term
paper?  Even the brightest English student won't find all typos in his
essays.  Why should computer users be held to higher standard?
\item I prefer to work in groups
\item I did not have many problems with EIAM.  I like the format and it does
make the review process much easier.
\item The scrolling bar for the review screen should be on the right not
left because the mouse is on the right.
\item a little confusing at the start knowing which button on the mouse to
use to do a certain function 
\item none
\item sometimes it gets stuck and had to start over
\item I was expecting to do a group review first and I have not prepared for
doing individual review psychologically.  It scared me almost freezed.
\item none
\item NOne
\end{itemize}

\item  Problems that I had with EIAM process (please explain): 
\begin{itemize}
\item my limited knowledge of 'C' was an impediment in the review process.
sometimes I was not sure what the function was doing
\item none
\item none
\item Read Above

\item none
\item 1. the confidence part is not that important at all. 
2. not enough on line help for the first time users.

\item 

\item Not really sure if it's any more efficient than pouring thru code
that's printed or jus' using a debugger...
\item There were no problems with the process
\item 	     .......................
I think that I summed it up in the last comment.  I've just been
spoiled by the online compilers in Borland 4.0 and Microsoft Visual
C++.  How can EIAM compete will compilers that contain online manuals?


\item None
\item The mouse is not very responsive.  The system is slow.
\item referring back to other source nodes was a little confusing at times
\item none
\item I understood the C++ codes that I turned in, but not the C codes that
you gave me to do.  I arrange the codes differently.  I bid I loose
lots of points of this review by programing in C++ instead of C 
\item none
\item It was opening some scratch file.
I did not know what it was.
\end{itemize}

\item  Suggestions on how to improve EIAM system (please explain):
\begin{itemize}
\item suggested in item 20.
\item Perhaps incorporate some sort of reference concerning the language
being used for the program.  After all the review process was to bring
up issues concerning the program's functionality.  Without knowing how
the language, in this case C/C++, works.  The review process has
become one which dealt with the syntax of the language rather than the
program it was used to to create.

\item Need the users practice using EIAM system, to get use to the buttons
and mouse controls.
\item run it on pc and make it more user friendly.
\item It doesn't really replace an active debugging program.
\item This is just minor, but rather than scrolling page by page, how about
line by line, meaning the top line of the page disappears, and the
bottom line incrents up by one, and a new line in the bottom
appears. This would be easier on my part in reviewing the code, rather
than having to flip page by page.
\item More days to debug
\item it is good
very easy to use
maybe let the user see more! or have a debugger 
\item I know that EIAM is NOT designed to compete directly against Borland
or Microsoft, but I can't see how it would improve my coding over
using such a compiler.  Personally, I know that I find a hell of a lot more
errors with gcc than with this, because gcc tells me where the errors are.
Yes, it is sloppy to do things that way, but if I didn't, I would
never be able to turn in any assignments.

As this is the age of hypermedia, I think that the system could be
improved if it offered hypertext links to help files.  For example,
clicking on the word "int" in the code being reviewed would take the
user to a help file on the "int" type.  This would be helpful for
people like me who don't know 100% of the C++ commands by heart.

\item maybe a hot key to automatically move hightlight to issue form.  Mouse
are a bit difficult to use.
\item include sample input and output for the functions and the program.
\item it is very good now.  I wish I could've spent more time using it for
all SE documents
\item I would prefer to have more windows for me to put more related
function on the screen at the same time.  Moreover, the room was full
of noise while I was taking the review.  The noising enviornment really distracted me
\item none
\item 1. Short demo before the actual thing.
2. Dynamic loading of the code for review
3. Better graphical user interface(Certainly better than emacs)

\end{itemize}


\item  Suggestions on how to improve EIAM process (please explain):
\begin{itemize}
\item online-help for 'C' functions do would be a great help.  The
reviewers would be able to avoid making issues which are not based on
clear understanding of the actual process going on in the program.
\item none

\item Read Above

\item 	     .......................
if there is an on line help, that's great!

\item Not really sure but I preferred the group process.
\item If we were allowed to take a short break.
\item No suggestions
\item I think I covered it in the last comment...
\item So far so good!
\item provide a set of procedures and suggestions on how to review more productively
\item ......................
I would prefer to do the group review first.  I quess I am pretty much
a chicken.  
\item none
\item It is fine as it is
\end{itemize}


\item  Other comments:
\begin{itemize}
\item none
\item The ics 411 project due date is very close to this CRCS which is very
good because I got to read my code in trying to prepare for this
review.  For future reviews, I think that we should start the review
after the project due date, but do the review as soon as possible.
\item 1.  the handout that Danu gave us is very detail and is a really good
    guideline for our first time users.
2.  the training session is good too. I gave me a rough idea of what
    this program is and how to use it.
3.  for the sample training program, if there is clues showing the
    users where the errors are, it will be more helpful for the users
	 to understand how to found out errors.
 
\item None.
\item This was really interesting!
\item The assignment should have been base on a different program, not the
one already done in class.
\item I don't mean to cut the system down out of spite.  It's just that I
cannot learn to adapt to this style of debugging.  I don't think that
it's safe to program entirely from memory, as I cannot remember 100%
of the possible C++ commands.  I need a manual, online help, and as
much assistance as possible.  
\item 
FIRST FAMILIAR WITH THE PROGRAM

\item The producer of the EIAM is very friendly and helpful!
\item I like how this system works, although i think it can be more
effective if there was a book of C programming language nearby
which would allow me to confirm my errors.
\item it would help to be prepared before coming in
\item it was well worth the time.  its a good system
\item The reviewing schedules are not flexible.  I bid you next time I will
do a better job.  Think you for reviewing my comments and have a
wonderful day!
\item this was an interesting reivew.  i needed to be dependant on only
myself and not anyone else.  this really made me concentrate harder and try
to understand the code a lot better.  i wish i was more prepared so i
would feel more confident in creating the issues.
\item None
\end{itemize}

\end{enumerate}

