
\chapter{Final Questionnaire (ICS411)}

\begin{enumerate}

\item  I would rather review source code manually using pencil and paper than
using CSRS (on-line Collaborative Software Review System).
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 1.8)

\item  In general, CSRS is useful.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.2)

\item  Overall, I enjoyed using CSRS.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5
\hfill Very true\\
(Score: 4.1)

\item I would:
   \begin{enumerate}
   \item[(1)]Strongly prefer using EGSM
   \item[(2)]Somewhat prefer using EGSM
   \item[(3)]Equally prefer using EGSM or EIAM
   \item[(4)]Somewhat prefer using EIAM
   \item[(5)]Strongly prefer using EIAM
   \end{enumerate}
(Score: 2.1)

\item I am:
   \begin{enumerate}
     \item[(1)] Much more productive using EGSM
     \item[(2)] Somewhat more productive using EGSM
     \item[(3)] Equally productive with EGSM or EIAM
     \item[(4)] Somewhat more productive using EIAM
     \item[(5)] Much more productive using EIAM
    \end{enumerate}   
(Score: 2.0)

\item  I prefer working:
  \begin{enumerate}
     \item[(1)] Alone 
     \item[(2)] Group
  \end{enumerate} 
(Score: 1.8)

\item  Reasons for preferring working in a group: 
\begin{itemize}
\item have more input as a group and we could talk.
Also, more heads are better than one

\item Comparison of my performance in each situation should speak for itself.
I don`t like programming alone.
\item Because there is no output, it is very hard to look for specific
errors.  I'd rather trace a program then review it's code any day.

\item It's more easy to find errors in group than individal.
\item If I have to work on code myself then I'd rather use a debugger or
simply look through the code via hardcopy but if working in a group I
find that the group discussion and consensus process allows me to find
more errors and be more critical because I can ask questions that
hopefully someone will have answers to.
\item I'm a group oriented type of person however at times I do like to work
on my own.

\item In group work even though you are not confident,
when you say and issue others may recognize and they will offer
support to the issue
\item In a group situation, there are things that others may see that I will
not see, and vice versa.  Also, the issues that a group member raises
may not be an issue and it can be determined in a group discussion.  I
think a group setting, with a group that works, is better and more
reliable than individual work.
\item collaborative review leads to higher quality and quality is job \#1
\item We can discuss the error before create a issue.  That's the good way
to eliminate unnecessary time to find errors.

\item The CSRS is a very useful system.  It would be a lot harder or
inefficient to review source code manually using pencil and paper. The
system was a neat way for reviewing code and I think improves the
amount and types of errors we find in a program.  Although it takes a
lot more effort and reliability on the individual effort, I would
prefer working in a group.  Groups are good because I get to see how
other people work out problems and working together, to me is more
efficient in finding errors. 
\item Working in group helped me to find more number of errors than the
individual review.

\item sometimes you get ideas from others that you never could thought of
you feel less stressed out


\item group is better than individual
because the presenter's explanation was very helpful to think about
the questions



\item I am a pepole person
\item Can discuss issues with other people and use them as a sounding board.

\item I think whether working in a group or as an individual depends upon
the assignment, group members, and the overall understanding of the
issues at hand.
\item Because if I did not know or understand some code then I could get
help and support from the other members of my group.
\item feel more confident about making decisions;

\item more chance of finding errors.
\item We can discuss and support by others with suggestionsIssue
\item it is much easier to 'bounce' ideas off of each other and see things
i never could see before. This makes it much more efficient to debug code.

\item It has brought our own individually ideas as to how things work and
how they should work.  The presentation of ideas allows us to
personally evaluate the issue, and see if it is true or not.
Sometimes a person may think of an issue, and another may not, and
vice versa.
\item I think this is a terrific system.  I do have a good fun in doing
the review.  I prefer to do the review in group instead of individual.
I wish you have a wonderful summer.
\item working in a group can stimulate more ideas and is far more effective
than doing alone. Also, if one gets an error, other group member can
point it out immediately to avoid unneccessary mistakes. moreover, I
found that working in a group can help solving some of the C syntax
questions too if one is not that familiar with certain functions in C.


\item It was not easy to look at all the codes and gets kind of boring after
an while.
\item Interaction for code is good !
\item Three eyes looking at one code will definitely
find errors, one way or the other.  Also we have
known each other for a long time and we worked
on other projects together.
\item my current limited knowledge of 'C' makes me less confident about the
issues I make.  I am sure I am missing out many issues.

\end{itemize}

\item  Reasons for preferring working alone:
\begin{itemize}
\item It's hard to work efficiently as a group.
I think all members should well familiar with each other, and
have good understanding about program.
\item Better concentration.
\item In a group, more time is spent for discussion and less time to
concentrate to find errors.  However, if we have a good group; it will
be better to do in the group because you can learn more from other findings.
\item Don't need to debate the issue.  It easier to work alone.
\item Group working always involves lots of disscutions and
debates. Sometimes, I feel so sick of those timeless arguments.
But, working in a group gives more productivity when working on a
large scale of project. That is why we software developers need an
efficient program like CSCR. Thanks!

\end{itemize}

\end{enumerate}

