\chapter{Background work}
This chapter presents a research background and a literature review of work related 
to the research problem.
The relevant theory is examined for each domain: software engineering, process mining,
timeseries analysis and activity patterns. Each section provides a concise review 
of related work, discussion of constructs and summary. 

Section \ref{software.processes} includes the general principles and properties of 
software process models, as well as a discussion of the current research in the area.
Section \ref{oss.processes} reviews studies in Open-Source software processes
explaining foundations for OSS processes and their recovery.
Section \ref{activity} includes a discussion of the activities performed within software
process, their meaning, and their relation to recurrent behaviors.

\section{Software process}\label{software.processes}
There are different approaches for software development which were designed in order to 
facilitate the creation of software systems. These approaches provide means for 
structuring of development activity. The main goal of imposing such a structure on the 
development process is to organize production of code in manageable way. 
According to the research, structuring software process yields a number of benefits:
\begin{itemize}
 \item whole software development cycle can be broken down onto number of one-step pieces;
 \item which, in turn, helps to keep clear focus on what must be delivered and when, during each step;
 \item it clarifies the project scope and improves time, effort, and cost estimates;
 \item it provides ability to measure the progress;
\end{itemize}
It is strongly advocated, that the use of established and well structured process is 
essential for the complex projects in order to orchestrate collaborative effort 
of multiple teams. 

\subsection{Software process models}
Structuring software process includes the use of a software process model and following 
a software method, also known as methodology. While latter is used primarily to navigate 
through the development process: determining a number of functional points, 
designing data flow diagrams etc., the model provides developers with guidance about their 
tasks and the activities that should be undertaken during development. 

Software models of development include three generic phases corresponding to software
life cycle: definition phase, development phase and a maintenance phase. 
The definition phase includes the initial planning of the future system and 
requirements collection: developers identify data need to be processed by the system, 
its functionality, behavior, and what constraints must be placed on the system design 
and development. 
The development phase focuses on the system implementation and testing: developers 
write the system code, test if the system satisfies to user requirements, 
has planned behavior and produces needed output. 
The last, maintenance phase, focuses on the post-development activities: 
system deployment and its operational support. 

Software models can be characterized by the series of distinct phases, in fact, the 
determination of phases, its ordering and the definition of the phase-transitioning 
criteria are the core of model components.
Each of these phases is executed with a particular goal: some will provide a part of the 
software system, or its validation, while other will deliver an engineering documentation 
or a user manual. Examples of such phases are the requirements collection, 
user manual writing, coding of a a functional module, etc.
As mentioned before, the definition and ordering of carried activities facilitates
a framework for estimation of resources, defines major milestones, and provides 
means for time and effort monitoring and management. 
\todo[inline]{Definitions!}

\subsection{Software process modeling research}
The research area dealing with Software Process Modeling (SPM) was receiving significant 
attention over the years and is recognized as one of the powerful technologies
in software process engineering. The empirical approach techniques play a critical role
in SPM and most of the work is based on the case studies and action research. 
However, this approach has an intrinsic flaw - it is hardly feasible since
SPM research activities are often limited by organizational context, and also extremely 
expensive - some of the known reports required decades to obtain an empirical evidence.
Thus, in most cases, it is unfeasible to perform an explanatory research in full rigor,
and as pointed in \cite{citeulike:11079867}, ``Currently the empirical studies in SPM were 
mostly exploratory in nature, whose strengths of empirical evidence were relatively weak...''
\todo[inline]{process modeling built up-down and it is expensive!}

\subsection{Software process conformance}
Another area of research related to my work is the area of the software process conformance. 
the research in this area is dealing with methods and theories for formulating, identifying and
investigating violations in the execution of software processes. These methods heavily rely
on the process elicitation step which is of most interest \todo[inline]{very similar}. 

\todo[inline]{Paragraph about ``classical'' software process research - heavyweight, expensive
mostly top-down oriented: first designed, secondly tested or confirmed}.

\section{Open Source Software processes}\label{oss.processes}

\subsection{Death of distance and cyberspace}
\subsection{Relaxation of space-time constraints and displacement}
\subsection{The Bazaar model}
\subsection{OSS process research}
\subsection{Mining Software Repositories}\label{background.msr.summary}
\input{background_msr_summary}

\section{Evidence of recurrent behaviors}

\section{Activity fragmentation, Activity-based modeling}\label{activity}

\section{Release early, Release often, and impact of small changes on process visibility}

\section{Literature search overview}

\section{Process mining}
The recognition of interest and importance of various processes 
According to the IEEE Task Force on Process Mining, established in 2009, ``Process mining is 
a relatively young research discipline that sits between computational intelligence and data 
mining on the one hand, and process modeling and analysis on the other hand'' \cite{citeulike:11077707}.
This group promotes the topic of process mining,

\subsection{Workflow mining, Business process mining}\label{mackground.bpm}
\input{background_business}

\subsection{Software process mining and discovery}\label{mackground.bpm}
\input{background_msr_details}

\section{Temporal data mining}
\subsection{Survey of temporal patterns mining}
\subsection{Symbolic aggregate approximation}

\section{Summary of literature review}

\section{Hypothesized relations between activity patterns and CSDL cycle}

\section{Activity Patterns Frequency and Activity Patterns Entropy as metrics}

\section{Research objectives}