\documentclass[12pt,oneside]{report}

%%% Load some useful packages:
%% "New" LaTeX2e graphics support.
\usepackage{graphicx}
%%	using final option to force graphics to be included even in draft mode
%\usepackage[final]{graphicx}
%% Tell graphicx the default directory for all figures
\graphicspath{{figures/}}

% Enable subfigure support
\usepackage{subfigure}

%% Make subsubsections numbered and included in ToC
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}

%% Package to linebreak URLs in a sane manner.
\usepackage{url}

%% Define a new 'smallurl' style for the package that will use a smaller font.
\makeatletter
\def\url@smallurlstyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{smallurl}

%% Define 'tinyurl' style for even smaller URLs (such as in tables)
\makeatletter
\def\url@tinyurlstyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\scriptsize\ttfamily}}}
\makeatother

%% Provides additional functionality for tabular environments
\usepackage{array}

%% Puts space after macros, unless followed by punctuation
\usepackage{xspace}

%% Make margins less ridiculous
\usepackage{fullpage}

%% Allows insertion of fixme notes for future work
\usepackage[footnote, nomargin]{fixme}

%%%% Turned off for tech report, should be turned on for research portfolio
%% Turn on double spacing
\usepackage{setspace}
\usepackage{mdwlist}
\doublespacing

%% Make URLs clickable
%\usepackage[colorlinks, bookmarks=false]{hyperref}
\usepackage[colorlinks, bookmarks=true]{hyperref}

%% Since I'm using the LaTeX Makefile that uses dvips, I need this
%% package to make URLs break nicely
\usepackage{breakurl}

\usepackage{amsmath,amsfonts}
\numberwithin{equation}{subsection}
%%\usepackage{nonfloat}
\usepackage{bbm}
\usepackage{setspace}
\onehalfspacing
\usepackage{tabularx}

%%% End of preamble
\begin{document}

\input{title.tex}

%% Philip suggests it needs a ToC
\tableofcontents

\begin{abstract}
%Abstract goes here if needed.
Software process, team (TSP) or personal (PSP) is a structured activity resulting in a software. Many well-defined software processes were designed. However, they are not proven to deliver consistently and fail with equal probability. This ``software crisis'' phenomena was coined in 1968 and since then studied by generations of researchers. Over the years, the research in software process matured and yielded a number of standards and principles, but overall, the picture changed a little. As an alternative opinion - the ``software development process as a craft'' idea emerged and found some strong supporters in the industrial and research communities. While both approaches for the software process design are supported by excellent research work and industrial success stories, in general, software processes remain designed ``top-down'' - i.e. first someone has to invent a process and to design and to implement its building blocks after. In my work I explore an alternative approach for the process design - through the discovery of recurrent behaviors from the software process artifacts trails, which may shed light on the larger building blocks of software process. In this chapter I will analyze various ways of data collection, aggregation and its informational content.
\end{abstract}

% my main text chapters
\input{vc-intro}
\input{eclipse-report}

%%% Input file for bibliography
\bibliography{seninp}
%% Use this for an alphabetically organized bibliography
\bibliographystyle{plain}
%% Use this for a reference order organized bibliography
%\bibliographystyle{unsrt}
%% Try using this BibTeX style that hopefully will print annotations in
%% the bibliography. This will allow me to make notes on papers in the
%% BibTeX file and have them readable in the references section until
%% I turn them into a conceptual literature review 
%\bibliographystyle{annotation}

\end{document}