%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% AFTR.tex -- 
%% Author          : Carleton Moore
%% Created On      : Fri Oct 18 16:21:58 1996
%% Last Modified By: Carleton Moore
%% Last Modified On: Tue Nov 19 15:26:31 1996
%% RCS: $Id: AFTR.tex,v 1.5 1996/11/20 01:26:51 cmoore Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1996 Carleton Moore
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{AFTR:  An Automated Formal Technical Review Method}
\label{sec:AFTR}



What would an automated FTR tool look like that did not allow measurement
dysfunction?

\begin{itemize}
\item{Individual reviewers will control where their data is stored.  e.g. a
private disk or access controlled file.}
\item{System will aggregate the individual data and not store individual
data.  The aggregate will be accessible only to the group/review members.
The data will have mean and standard deviation so members will be able to
see how they are doing vs the group.}
\item{Number and severity of defects found will only be aggregated into
group numbers then deleted.  Only the author will have a record of the
defects and locations.  Once the review meeting is finished the ``scribe''
sheet will be only accessible to the author.}

\end{itemize} 



To remove dysfunction all personal review data will be kept private. 

Some of the more important implications of this principle are: 

      Individual preparation times are private. 
      Individual defect lists generated during preparation are private. 
      Rework times are private. 

Personal data is always reported in aggregate and anonymously.


Some implications of this principle are: 

      Preparation times are recorded as an aggregate value over all reviewers. No data
      accessible to the organization records individual preparation times. 
      Participation in reviews is private data. No data accessible to the organization records
      which individuals participated in which reviews. 
      Defect lists are recorded as an aggregate of all individual lists and without author
      attribution. No data accessible to the organization records which individuals uncovered
      which defects. 

Individual process improvement is an individual matter.

Development group review data is private.


Group review data is always reported qualitatively.


Group process improvement is a group matter.




\subsection{Design}

\subsection{Architecture}

\subsection{Metrics Collected}