%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% experiment.tex -- 
%% Author          : Carleton Moore
%% Created On      : Fri Oct 18 16:22:32 1996
%% Last Modified By: Carleton Moore
%% Last Modified On: Wed Nov 20 16:42:08 1996
%% RCS: $Id: experiment.tex,v 1.4 1996/11/21 02:42:20 cmoore Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1996 Carleton Moore
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{Experimental design}
\label{sec:experiment}

I will conduct an experimental case study to test the following hypotheses
regarding the automation of a Formal Technical Review process.
\subsection{Hypotheses}
\begin{enumerate}
\item[H1.1:]{There will be observable differences in review
effectiveness between the manual process and the automated process.}
\item[H1.2:]{There will be a observable difference in cost between
the manual FTR process and the automated FTR process.}
\item[H1.3:]{The organization will adopt AFTR as their FTR tool.}
\item[H1.4:]{The participants will prefer to use the automated FTR
process over the manual FTR process.}
\item[H2.1:]{Analysis of the process defects will lead to Process
Improvement Proposals (PIPs).}
\item[H2.2:]{Process Improvement Proposals (PIPs) from the method will
provide insight to improving the technology transfer method.}
\end{enumerate}
\subsection{Related Objectives}

A related goal of this experiment is to improve the software development
process  of Tektronix.  By gaining insight into the FTR process we will
better understand how to develop high quality software.

\subsection{Experimental Method}
The manual method has been introduced into the Instrument Business Unit of
Tektronix in Beaverton.  This group will conduct reviews of their
requirements, designs, and source code using the TekInspect method from
October, 1996 through May, 1997.  During this period they will record
several metrics;
\begin{itemize}
\item{number of participants}
\item{time spent in each phase per participant}
\item{size of work product}
\item{number and severity of defects found in work product}
\item{deviations from the base method}
\item{time spent fixing defects}
\item{pass or fail judgment of review team}
\end{itemize}

During the summer of 1997 we will introduce the automated FTR tool
described in Section \ref{sec:AFTR}.

\subsection{Experimental Procedures}

\subsection{Pilot Study}

I plan on using CSDL as a pilot to evaluate AFTR's usability, the user
questionnaire, the metrics collection, and the metrics analysis.  We are
currently using TekInspect as CSDL's review method.  Between October, 1996
and December, 1997 we should conduct many reviews of different work
products in CSDL.  




\subsection{Experimental Limitations}
\subsection{Anticipated Results}
