%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% personnel.tex -- 
%% Author          : Philip M. Johnson
%% Created On      : Tue Nov 11 13:31:41 2003
%% Last Modified By: Philip M. Johnson
%% Last Modified On: Tue Nov 11 13:47:45 2003
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2003 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 


\newcommand{\newcategory}[1]{\newenvironment{#1}
 {\sectionheading{#1}\begin{list}{}{\setlength{\labelwidth}{0cm} \setlength{\labelsep}{0cm} \setlength{\itemsep}{0ex plus0.2ex} \setlength{\itemindent}{-1cm} \setlength{\leftmargin}{1cm} \setlength{\parsep}{0ex plus0.2ex}}}{\end{list}\par}}
\newcommand{\sectionheading}[1]{\medskip\pagebreak[2]\par\noindent
 {\small\bf #1}\nopagebreak}

\section{Personnel}

\subsection{Philip Johnson}

\newcategory{Degrees}
\begin{Degrees}
\item Ph.D.~in Computer Science, University of Massachusetts. 1990 
\item M.S.~in Computer Science, University of Massachusetts.  1985
\item B.S.~in Biology, B.S. in Computer Science, University of Michigan. 1980
\end{Degrees}

\newcategory{Academic Appointments}
\begin{Academic Appointments}
\item {\em Professor},  Information and Computer
  Sciences, University of Hawaii.  2001---present

\item {\em Associate Professor},  Information and Computer
  Sciences, University of Hawaii.  1995---2001

\item {\em Senior Research Fellow},  Distributed Systems Technology Centre,
University of Queensland, Australia, 1997.

\item {\em Assistant Professor},   Information and Computer
  Sciences, University of Hawaii.  1990---1995

\end{Academic Appointments}


\newcategory{Selected Publications}
\begin{Selected Publications}
  
\item P.~M.~Johnson and C.~A.~Moore and J.~A.~Dane and R.~S.~Brewer, {\em
    Empirically Guided Software Effort Guesstimation}.  IEEE Software, Vol.
  17, No. 6, December 2000.
  
\item P.~M.~Johnson and A.~M.~Disney, {\em A Critical Analysis of PSP Data
    Quality: Results from a Case Study}.  Journal of Empirical Software
  Engineering, Volume 4, December, 1999.
  
\item P.~M.~Johnson, {\em Reengineering Inspection}.  In Communications of
  the ACM, Volume 41, No. 2, February, 1998.
  
\item P.~M.~Johnson and D.~Tjahjono, {\em Does Every Inspection Really Need
    A Meeting?}, In Journal of Empirical Software Engineering, Volume 4,
  No. 1, January 1998.
  
\item A.~A.~Porter and P.~M.~Johnson, {\em Assessing Software Review
    Meetings: Results of a Comparative Analysis of Two Experimental
    Studies}.  In IEEE Transactions on Software Engineering, vol. 23, no.
  3, March 1997.
  
\item P.~M.~Johnson and A.~M.~Disney, {\em The Personal Software Process: A
    Cautionary Case Study}.  In IEEE Software, Volume 15, No. 6, November,
  1998.

\item P.~M.~Johnson, {\em Design for Instrumentation: High Quality
    Measurement of Formal Technical Review}.  Software Quality Journal,
  Volume 5, March, 1996.

\end{Selected Publications}

\newcategory{Professional Activities}
\begin{Professional Activities}
\item {\em Editorial Board}, IEEE Transactions on Software Engineering,
2000-present.

\item {\em Program Chair}, International Software Engineering Research Network
 Annual Meeting, 2000.

\item {\em Advisory Board Member}, 
  International Journal of Computer Supported Cooperative Work

\item {\em Program Committees}: \\
  XP/Agile Universe,  2004. \\
  International Software Metrics Symposium, 2002, 2003, 2004. \\
  International Symposium on Empirical Software Engineering, 2002.\\
  European Conference on Computer Supported Cooperative Work, 1997, 1999.\\
  Software Architectures for Cooperative Systems Workshop, 1994 \\
  CSCW Tools and Technologies Workshop, 1992, 1993 \\

\end{Professional Activities}


\subsection{Other personnel}

In addition to the Principal Investigator, other personnel associated with
this project include:

\begin{itemize}

\item {\em Dan Port}, Assistant Professor of Information Technology
Management, University of Hawaii College of Business Administration.

\item {\em 4 Graduate Students, Information and Computer
Sciences}. Currently we have two Ph.D. students and two M.S. students
working on this project.

\end{itemize}


