%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Plain-Tex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%
%% thesis-acknowledgment.tex -- 
%% Author          : Dadong Wan
%% Created On      : Thu Apr  7 01:04:21 1994
%% Last Modified By: Carleton Moore
%% Last Modified On: Thu Jul  6 18:20:19 1995
%% Status          : Unknown
%% RCS: $Id: acknowledgments.tex,v 1.5 94/04/28 15:52:00 dxw Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgements{

The last year has been quite an adventure.  Designing, implementing,
using, and evaluating AEN has been a long journey.  With out the invaluable
support of the following people this thesis would not have been possible.

Philip Johnson whose advice, critiques, and support helped me make it
through this journey.  I thank you for providing an environment that
fosters world class programming and research, for your time and effort into
reviewing this thesis, for the many discussions on the bus going home.

I am grateful to other members of my committee, Martha Crosby and James
Corbett for their insight and perspectives.  They opened my eyes to new
ways of viewing AEN.

Rosemary Andrada, who while writing her own thesis, was a great sounding
board for ideas.  Thank you for your friendship during the long hours in
Keller 304.

I feel blessed by being a part of the Collaborative Software Development
Laboratory (CSDL).  I thank the members of CSDL both active and emeritus
for their support, their discerning input, reviews of code and
documentation, and helping with the implementation of AEN.

Finally, I thank my parents and family.  With out their support I would
never have gotten this far.  So, I dedicate this thesis to them.

}



