\subsection{Summary}

While attempting to understand the nature and benefits of inspection 
meetings we uncovered anecdotal evidence that pointed to two possible
explanations of meeting performances. One explanation was that different 
inspection processes have different goals for the first and second inspection 
phases. Another explanation is that faults found at the meeting might 
result from factors other than meeting synergy or teamwork.

These observations led us to investigate whether two different 
inspection processes (PI and DC methods) had different performances. 
We also investigated whether the benefits of an inspection meeting could 
be achieved by simply allowing reviewers to conduct further individual analysis
(DD method). 

Our results suggest that the inspection method used is a significant 
factor in the inspection performances. We found that there was no
significant difference in the defect densities found by the DC and PI
method. The DC method may produce more non-defect issues and false positives 
than the PI method, but the difference is not yet statistically significant. 
We found that DD method inspections detected more new faults in the second 
phase and more total faults than either the DC or PI method. Finally, we found
no evidence that the PI and DC methods found specific faults with greater
frequency than the DD method.

These results suggest that (even well-prepared for) meetings do not find more 
defects than individuals working alone. Furthermore, it appears that meetings
function more as a filtering process (discarding false positives) than as 
a defect discovery process.

From the perspective of the software practitioner, this outcome suggests that 
meetings are not necessarily the most effective way to discover large numbers of defects.
On the other hand, meetings suppress false positives (and some true defects) and
allow groups to make decisions and coordinate activities.  Therefore, inspection meetings 
should be held if the needs of the organization warrant them, but they can not be justified
on the basis of defect discovery. 

