%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% proposal.tex -- 
%% Author          : Jennifer Geis
%% Created On      : Fri Sep  5 13:50:18 1997
%% Last Modified By: Jennifer Geis
%% Last Modified On: Tue Sep 30 14:03:26 1997
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1996 Jennifer Geis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\documentstyle[nftimes,11pt,/group/csdl/tex/definemargins,
/group/csdl/tex/lmacros]{report} 
\input{/group/csdl/tex/psfig/psfig}
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\begin{document}


\title{JavaWizard: An Automated Code Checker For Java} \author{Jennifer
  Geis\\ Collaborative Software Development Laboratory,\\ Department of
  Information and Computer Sciences\\ University of Hawaii, Manoa\\ 
  Honolulu, Hawaii 96822\\ {\tt jgeis@uhics.ics.hawaii.edu}} \maketitle

\tableofcontents

\chapter{Introduction}
\section{Thesis Statement} 
\section{Objective}

\chapter{Related Work}
\section{Automated Debugging}
Automated debugging involves using a computer program that scans through
code and looks for possible problem areas. Usually the program just
notifies the developer of the potential problem and the developer decides
whather or not it really is a defect.

\subsection{Lint-type tools}
LINT's original purpose was to locate bugs and inefficiencies in C source
code.  Later, it was expanded to handle portability issues.  The
portability problems arise from programs including system dependent
operations.  LINT shows where you have written system-dependent
(non-portable) code.

\subsection{AI tools}
\subsection{HCI tools}
\subsection{Program Comprehension}

\section{Non-Automated Debugging}
\subsection{Individual}
Non-Automated individual debugging typically involves the developer acting
as reviewer of their own product.

\subsubsection{Personal Software Process}
The Personal Software Process is a process of developing software designed
by Watts Humphrey of the Software Engineering Institute.  The method
includes doing personal reviews of both design and code.  The purpose of
these reviews is to identify defects before the next stage of development.
The idea is that the longer a defect stays in the product, the more costly
it is to remove it.  By performing reviews, the developer can identify the
problem areas and fix the defects prior to the next phase.  Through the use
of the Personal Software Process developers have managed to decrease the
number of defects found in compile and test by 50 percent or more.





\subsection{Group}
\subsubsection{Formal Technical Review}

\chapter{JavaWizard}

\section{Motivation}
\subsection{Relation to Automated/Non-Automated Debugging}
\subsection{Preliminary Case Study}
\subsubsection{Data Collection}
\paragraph{ICS414 and CSDL}
\subsubsection{Initial Observations}
\paragraph{Common Mistakes}
\paragraph{Testing Time}
\paragraph{Defect Frequency}

\section{System Requirements}
\subsection{Cut Testing Time}
\subsection{Fewer Defects in Test (For which Jwiz is Designed)}
\subsection{Possible benefits}
\subsubsection{Shorter overall development cycle}
\subsubsection{LOC/HR increase}
\subsubsection{Test Time as a Smaller Percentage of Overall Cycle}

\section{Design}
\subsection{Java}
\subsubsection{Platform Independent}
\subsection{JavaCC}
\subsection{Symbol Table}
\subsection{Design Pattern}
\subsection{How Jwiz Works}
\subsubsection{Examples}
\paragraph{String == String}
\paragraph{Missing GUI Components}
\subsection{Current State}
\subsection{Future Additions/Modifications}
\subsubsection{Defect Checks}
\subsubsection{Extensibility: Adding New Tests}
\paragraph{JavaCC Abstraction}
\subsection{Defect Toggling/Selection}

\section{Development Issues}
\subsubsection{Extensibility}
\subsection{Performance}
\subsection{What Makes a Defect?} Not always a defect
\subsection{Complexity: Number of Tests}

\chapter{Research Plan}
\section{Personal Thesis Process (PTP)}
\section{What Makes a Defect?} 
How to determine if a defect should be checked for by an automated
code-checker.
\subsection{Process of Determining a Defect}
\subsection{Frequency}   Obscure?
\subsection{Commonality} How many people make it (one or everyone)?
\subsection{Performance Impact}
\subsubsection{Automated vs. FTR}
\subsubsection{Automated vs. Personal Review}

\section{Data Collection}
\subsection{Experiment}
\subsubsection{ICS 613}
\paragraph{PSP}
\subparagraph{Manual Defect Recording} \subparagraph{Personal/Individual
  Review vs. Jwiz} \subparagraph{Automated Data Tracking} (w/in Jwiz)
\subsubsection{CSDL}
\paragraph{PSP}
\subparagraph{Manual Defect Recording} \subparagraph{Personal/Individual
  Review vs. Jwiz}
\paragraph{FTR}
\subparagraph{Group Review vs. Jwiz}
\paragraph{Automated Data Tracking}

\section{Data Analysis/Evaluations}
\subsection{Possible Outcomes}
\subsubsection{Comparison to PSP/FTR}
\paragraph{Time/Defect}
\paragraph{Efficiency}
\paragraph{Cost Savings}
\paragraph{Reliability}
\subsubsection{Explanations/Hypothesis}

\chapter{Timeline}
\section{Monthly Schedule}

\subsection{September '97}
\subsubsection{Package}
\subsubsection{Outline}
\subsubsection{Research}
\subsubsection{Revise PSP Defect Recording Form}
\subsubsection{Chapter 2 (Related Work) Rough Draft}

\subsection{October '97}
\subsubsection{GUI toggling}
\subsubsection{Chapter 2 Revision}
\subsubsection{Chapter 4 (Research Plan) Rough Draft}
\subsubsection{Add Automic Recording/Reporting Mechanism to Jwiz}
\subsubsection{Jwiz Release}
\subsubsection{Defect Data Collection}
\subsubsection{Eight New Defect Checks}

\subsection{November '97}
\subsubsection{Chapter 4 Revision}
\subsubsection{Chapter 3 (JavaWizard) Rough Draft}
\subsubsection{Defect Data Collection}
\subsubsection{Eight New Defect Checks}

\subsection{December '97}
\subsubsection{Chapter 3 Revision}
\subsubsection{Chapter 1 (Introduction) Rough Draft}
\subsubsection{Defect Data Collection}
\subsubsection{Eight New Defect Checks}

\subsection{January '98}
\subsubsection{January 23, Last day to file Spring degree applications}
\subsubsection{Chapter 1 Revision}
\subsubsection{Defect Data Collection}
\subsubsection{Defect Data Analysis}
\subsubsection{Chapter 5 (Conclusion) Rough Draft}
\subsubsection{Eight New Defect Checks}

\subsection{February '98}
\subsubsection{Chapter 5 Revision}
\subsubsection{Date ? - Give Committee copy of thesis}
\subsubsection{Eight New Defect Checks}

\subsection{March '98}
\subsubsection{Date ? - Defense of Thesis}
\subsubsection{Eight New Defect Checks}

\subsection{April '98}
\subsubsection{April 8, Last Day to Submit Thesis to Graduate Division}
\subsubsection{Eight New Defect Checks}

\subsection{May '98}
\subsubsection{May 17, Commencement}

\end{document}








