%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 03-04-ch1.tex --
%% Author          : Hongbing Kou
%% Created On      : Fri Sep  5 13:50:18 1997
%% Last Modified By: Hongbing Kou
%% Last Modified On: Sat Sep 25 23:24:09 2004
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%

\chapter{Related Work}
``Test-Driven Development (TDD) is a software development practice in which
test cases are incrementally written prior to code
implementation.''\cite{George_2003} Here test case is the unit test, which
is a piece of code written by a developer that exercises a very small,
specific area functionality of the code being tested. The rational of TDD


is to ``Analyze a little, test a little, code a little, and test a
little,repeat.'' The goal of TDD is to write ``clean code that
works''\cite{Beck_TDD_2003} and it has two basic rules.
\cite{Beck_TDD_2003}:

\begin{enumerate}
\item Write new code only if an automated test has failed.
\item Eliminate duplication (Refactoring).
\end{enumerate}

\begin{description}
\item They imply an order to do programming task \cite{Beck_TDD_2003}:
\begin{enumerate}
\item \emph{Red}\newline Write a little test that does not work, and perhaps
does not even compile at first.
\item \emph{Green}\newline Make the test work quickly, committing whatever
sins(for example, constant, fake implementation) necessary in the process.
\item \emph{Refactor}\newline Eliminate all the duplications created in
merely getting test work.
\end{enumerate}
\end{description}

In book ``Test-Driven Development by Example'', Ken Beck claimed that
red/green/refactor rhythm is the mantra of TDD, once you have an automated
suite of tests you will never go back. It gives incredible confidence in
your code. He also said TDD will naturally generate 100\% code coverage and a
coverage tool is not necessary at all if you do TDD perfectly.
Unit tests are very important to software development because it makes it
easy to check whether the system functions properly or not. It allows the
programmer to refactor code at a later date, and be sure the model is still
functioning properly\cite{UnitTest}.Also the automatic unit test suite created
in the development process could be used as regression tests. And ``a unit
test behaves as executable documentation, showing how you expect code to
behave under the variation condition you've considered.''\cite{Andy&Dave_2003}

In common software development unit test is not a must and is usually done
as an afterthought. Occasionally no tests are created at all, especially
with tight schedules. With TDD, before writing implementation,the automated
unit tests are written first and they are created on the ground of
requirement analysis, which leads to better and effective test suites.
  
However, in reality, software developers are educated to program first and
walk though the program to make sure the system works well to the desired
scenarios. As a new and counter-intuitive approach TDD requires discipline,
training and good tool support. Kent Beck suggested the ``xUnit''
framework. ``xUnit'' has many variances and it has been ported to more than
30 languages' support\cite{XPSoftware}. xUnit is a very light-weight tool
and it can be mastered by a novice programmer in several minutes. But to
applications like graphic user interface, socket programming, parallel
computing, database querying etc. It is hard to write small unit test to
deploy them or not practical. TDD practitioners suggest mocking technique
to mock up the real operations for test purpose. There are two widely
accepted mock tools called MockObject and EasyMock. Mock technique is
still under development and is welcomed by many practitioners in daily
programming work.

Boby George's analysis and quantification of test-driven development
concluded that both students and professional TDD developers appear to have
higher code quality\cite{George_2002}. E. Michael Maximilien and Laurie
A. Williams found that defect rate of project IBM Retail Store
Solution was reduced by 50\% compared to another system built with ad-hoc
unit test approach \cite{Maximilien_2003}, and TDD developers passed 18\% more
functional black-box tests than non-TDD developers. 











