\documentstyle[times]{acmconf}
\input{/group/csdl/tex/psfig/psfig}

\begin{document}

\title{Investigating Data Quality Problems in the PSP \\ \par \medskip 
       {\small Experience Paper}}

\author{
        Anne M. Disney\\
        Philip M. Johnson
        Dept. of Information and Computer Sciences\\
        University of Hawaii\\
        Honolulu, HI  96822 USA\\
        anne@ics.hawaii.edu\\
        johnson@hawaii.edu}
       }

%\toappearbox{Submitted to the Sixth International Symposium on the
%Foundations of Software Engineering (FSE-6), Orlando, Florida, November, 1998.}

\maketitle

%PJ -- abstract modified
\begin{abstract}
  
  The Personal Software Process (PSP) is used by software engineers to
  gather and analyze data about their work.  Published studies typically
  use data collected using the PSP to draw quantitative conclusions about
  its impact upon programmer behavior and product quality.  However,
  our experience using PSP in both industrial and academic settings
  revealed problems both in collection of data and its later analysis.
  We hypothesized that these two kinds of data quality problems could make a
  significant impact upon the value of PSP measures.  To test this
  hypothesis, we built a tool to automate the PSP and then examined 89
  projects completed by ten subjects using the PSP manually in an
  educational setting.  We discovered 1539 primary errors and categorized
  them by type, subtype, severity, and age.  To examine the collection
  problem we looked at the 90 errors that represented impossible
  combinations of data and at other less concrete anomalies in Time
  Recording Logs and Defect Recording Logs.  To examine the analysis
  problem we developed a rule set, corrected the errors as far as possible,
  and compared the original and corrected data.  This resulted in
  significant differences for measures such as yield and the
  cost-performance ratio, confirming our hypothesis.  Our results raise
  questions about the accuracy of manually collected and analyzed PSP data,
  indicate that integrated tool support may be required for high quality
  PSP data collection and analysis, and suggest that external measures
  should be used when attempting to evaluate the impact of the PSP.

\end{abstract}

\subsection*{Keywords}

\noindent Personal software process, defects, 
empirical software engineering,
measurement dysfunction, automated process support\\


\input{intro}
\input{overview}
\input{problems} 
\input{case-study}
\input{results}    
\input{conclusions}  

\bibliographystyle{plain}
\bibliography{/group/csdl/bib/psp,/group/csdl/bib/ftr} 


\end{document}



