\begin{acknowledgments}
%\begin{acknowledgements}

I would like to express my deep and sincere gratitude to my advisor, Professor Philip Johnson. Your wide knowledge and creative way of thinking have been a great value of me. Thank you for spending numerous hours on proof-reading this manuscript. This thesis would never have become possible without you.

I would like to thank Dr. Hakan Erdogmus from National Research Council of Canada. With your guidance and mentoring, August 2006 has become the most productive month in my thesis research.

I would like to thank Dr. Daniel Suthers, Dr. Dan Port and Dr. David Pager. Your advising and support have been priceless in my graduate studies. 

I would like to thank my thesis committee members: Dr. Wes Peterson and Dr. Kim Binsted. Your time and support are precious to me.

I would like to thank my fellow CSDL hackying buddies: Qin Zhang, Joy Augustin, Aaron Kagawa, Julie Sakuda, Takuya Yamashita, Mike Paulding, Robert Brewer, Pavel Senin, Austen Ito, Jitender Miglani and Christopher Lofi. 

I would like to thank Wesley Sugimoto and Janice Oda-Ng from the Department of Information and Computer Sciences. Thank you two for spending so much time on the tedious paper work for me.

Lastly, I would like to thank my friends Xin Zhao, Qin Guo, Ming Liu, Hao Zhou, Hu Li (Tiger), Decheng Yang, Hongbo Ding, Ying Zhang, Shidong Kai, Yangheng Zheng, Yuhuan Li, Rui Xue, Xiangli Xu and many others. Life would not be so wonderful without you guys.
\end{acknowledgments}
%\end{acknowledgements}








