%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% bio.tex -- 
%% Author          : Philip Johnson
%% Created On      : Tue Nov  4 10:26:48 1997
%% Last Modified By: Philip Johnson
%% Last Modified On: Tue Mar 31 11:15:28 2009
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1997 Philip Johnson
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 


%\section*{Biographical Sketch}

%%      Command to define new categories:
\newcommand{\newcategory}[1]{\newenvironment{#1}
 {\sectionheading{#1}\begin{list}{}{\setlength{\labelwidth}{0cm} \setlength{\labelsep}{0cm} \setlength{\itemsep}{0ex plus0.2ex} \setlength{\itemindent}{-1cm} \setlength{\leftmargin}{1cm} \setlength{\parsep}{0ex plus0.2ex}}}{\end{list}\par}}
\newcommand{\sectionheading}[1]{\medskip\pagebreak[2]\par\noindent
 {\small\bf #1}\nopagebreak}

\begin{center}
Philip M. Johnson\\
Information and Computer Sciences \hfill (808) 956-3489

University of Hawaii              \hfill fax: (808) 956-3548

1680 East-West Road               \hfill johnson@hawaii.edu

Honolulu, HI~~~96822              \hfill http://www.ics.hawaii.edu/$\sim$johnson/

\end{center}

\newcategory{Professional Preparation}
\begin{Professional Preparation}
\item Ph.D.~in Computer Science, University of Massachusetts. 1990 
\item M.S.~in Computer Science, University of Massachusetts.  1985
\item B.S.~in Biology, University of Michigan. 1980
\item B.S.~in Computer Science, University of Michigan. 1980
\end{Professional Preparation}

\newcategory{Appointments}
\begin{Appointments}
\item {\em Professor},  Information and Computer
  Sciences, University of Hawaii.  2001---present

\item {\em Associate Professor},  Information and Computer
  Sciences, University of Hawaii.  1995---2001

\item {\em Senior Research Fellow},  Distributed Systems Technology Centre,
University of Queensland, Australia, 1997.

\item {\em Assistant Professor},   Information and Computer
  Sciences, University of Hawaii.  1990---1995

\end{Appointments}




\newcategory{Publications: Closely Related}
\begin{Publications: Closely Related}
  
\item P.~M.~Johnson and C.~A.~Moore and J.~A.~Dane and R.~S.~Brewer, {\em
    Empirically Guided Software Effort Guesstimation}.  IEEE Software, Vol.
  17, No. 6, December 2000.
  
\item P.~M.~Johnson and A.~M.~Disney, {\em A Critical Analysis of PSP Data
    Quality: Results from a Case Study}.  Journal of Empirical Software
  Engineering, Volume 4, December, 1999.
  
\item P.~M.~Johnson, {\em Reengineering Inspection}.  In Communications of
  the ACM, Volume 41, No. 2, February, 1998.
  
\item P.~M.~Johnson and D.~Tjahjono, {\em Does Every Inspection Really Need
    A Meeting?}, In Journal of Empirical Software Engineering, Volume 4,
  No. 1, January 1998.
  
\item A.~A.~Porter and P.~M.~Johnson, {\em Assessing Software Review
    Meetings: Results of a Comparative Analysis of Two Experimental
    Studies}.  In IEEE Transactions on Software Engineering, vol. 23, no.
  3, March 1997.

\end{Publications: Closely Related}


\newcategory{Publications: Other Significant}
\begin{Publications: Other Significant}

\item P.~M.~Johnson, M.~L.~Moffett, and B.~T.~Pentland, {\em Lessons
Learned from VCommerce: A virtual environment for interdisciplinary
learning about software entrepreneurship}, To appear in Communications of
the ACM, 2002.

\item P.~M.~Johnson, {\em Leap: A ``Personal Information Environment'' for
Software Engineers}. In
  Proceedings of the 1999 International Conference on Software Engineering,
  Los Angeles, CA., May 1999.
  
  
\item P.~M.~Johnson and A.~M.~Disney, {\em The Personal Software Process: A
    Cautionary Case Study}.  In IEEE Software, Volume 15, No. 6, November,
  1998.

\item P.~M.~Johnson, {\em Design for Instrumentation: High Quality
    Measurement of Formal Technical Review}.  Software Quality Journal,
  Volume 5, March, 1996.

\item P.~M.~Johnson, {\em An Instrumented Approach to Improving Software
  Quality through Formal Technical Review}. In Software Inspection: 
  An Industry Best Practice.  David A. Wheeler, Bill Brykczynski, and
  Reginald N. Meeson, Jr., Editors.  IEEE Computer Society Press. 1996

  Also appearing in the Proceedings of the 16th International Conference on
  Software Engineering, Sorrento,  Italy. 1994.


\end{Publications: Other Significant}


\newcategory{Synergistic Activities}
\begin{Synergistic Activities}
\item {\em Editorial Board}, IEEE Transactions on Software Engineering,
2000-present.

\item {\em Program Chair}, International Software Engineering Research Network
 Annual Meeting, 2000.

\item {\em Chair}, Hawaii Java Users Group, 1996-present

\item {\em Advisory Board Member}, 
  International Journal of Computer Supported Cooperative Work

\item {\em Program Committees}: \\
  International Symposium on Empirical Software Engineering, 2002.\\
  European Conference on Computer Supported Cooperative Work, 1999.\\
  European Conference on Computer Supported Cooperative Work, 1997.\\
  Software Architectures for Cooperative Systems Workshop, 1994 \\
  CSCW Tools and Technologies Workshop, 1993 \\
  CSCW Tools and Technologies Workshop, 1992

\end{Synergistic Activities}



\newcategory{Collaborators}
\begin{Collaborators}
\item Simon Kaplan, University of Queensland
\item Ken King, Digital Equipment Corp.
\item Adam Porter, University of Maryland
\item Allen Sampson, Tektronix, Inc.
\item Joe Van Ryzin, Makai Ocean Engineering, Inc.
\end{Collaborators}


\newcategory{Thesis Advisor and Postgraduate-Scholar Sponsor}
\begin{Thesis Advisor and Postgraduate-Scholar Sponsor}
\item Carleton Moore, Orincon, Inc.
\item Danu Tjahjono, Aloha Networks, Inc.
\item Dadong Wan, Accenture, Inc.
\item Robert Brewer, LavaNet, Inc.
\item Mette Moffett, hotU, Inc.
\item Monir Hodges, Honolulu Community College.
% CIS Ph.D. students (D. Wan, D. Tjahjono, R. Tokuyama), 
% MS students (J. Geis, A. Disney, R. Brewer, C. Moore, R. Andrada, K. Ram, J. Johnson, F. Palit)
% EE Masters student (W. Kwan)

\item {\em Total graduate students advised:} 23

\end{Thesis Advisor and Postgraduate-Scholar Sponsor}

\newcategory{Graduate and Postdoctoral Advisors}
\begin{Graduate and Postdoctoral Advisors}
\item Jack Wileden, University of Massachusetts
\item Victor Lesser, University of Massachusetts
\item Wendy Lehnert, University of Massachusetts
\item David Stemple, University of Massachusetts
\item Daniel Corkill, Blackboard Technology Group
\end{Graduate and Postdoctoral Advisors}











