%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% budget.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Jul 27 13:31:23 1994
%% Last Modified By: Philip Johnson
%% Last Modified On: Thu Apr 16 09:22:14 2009
%% Status          : Unknown
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 


\section*{Budget Justification}

\pagenumbering{arabic}
\renewcommand{\thepage} {F--\arabic{page}}

\subsection*{Salaries}

The project budget provides salary support for the principal investigator
(two summer months) and two graduate research
assistants (11 months) for each of the three years.  The principal
investigator will perform or supervise all major system design enhancements
and empirical studies, and manage and train the graduate student
assistant.

\subsection*{Travel}

The project budget provides funds for two trips per year from Hawaii to the
mainland.  These funds will be used by the principal investigator to attend
one relevant conference (e.g., ICSE, FSE, ISERN, or STAR) each year, as well 
as attend one CPATH meeting per year. 

\subsection*{Equipment}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise server and several
workstations, along with a printer and networking peripherals.  The project
budget provides funds of \$5,000 per year for each year of the grant to
provide hardware and software related to this equipment.

\subsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
A summer salary of \$18,000 is budgeted for the principal investigator for
each of the three years.  This is less than two months of the principal
investigator's annual nine month salary (\$11,638/month), but allows for funding
for two graduate students during the grant period while staying under the 
CPATH Class I funding ceiling of \$300,000.

\paragraph*{Research Assistants.}  
This cost category provides support for two graduate research assistants
for each of the three years.  The salary for the research assistant is
based upon an 11 month RA-Step 5 salary (\$18,198) for each year.

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 3.5\% for the principal
investigator and 25\% for the graduate assistants.

\paragraph*{Overhead.}  
University of Hawaii overhead cost is calculated as 38.4\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).





