
\chapter {Questionnaire For Individuals (EIAM) --ICS313}

\begin{enumerate}

\item  My understanding of the source code before the review was: 
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.2)

\item  My understanding of the source code after the review was: 
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.7)

\item  My understanding of C++ programming language was improved after
this review.  
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.4)

\item  In general, I found it easy to understand the logic of
the code. 
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.2)

\item  For this review, I would have preferred working alone rather 
than working in a group (if I have had a choice).
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 2.6)

\item  For this review, I felt more confidence about the issues I raised when 
working alone rather than working in a group.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 2.3)

\item  The training session for individual review was sufficient.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.8)

\item   There was sufficient time to work on this review.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.9)

\item   I was motivated to do this review project.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.0)

\item  My overall confidence in the quality of my review was
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.1)

\item  I felt comfortable doing this review.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.5)

\item  It took me a good while to decide whether a particular program
segment contained valid issues.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.7)

\item  I often wanted to change or delete issues I just created.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 2.9)

\item  My overall satisfaction with individual review (EIAM) process as
outlined in the FTR guideline was:
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.7)

\item  I believe EIAM system made my review more productive (i.e., find
lots of issues) 
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.3)

\item  I believe EIAM system made my review more effective
(i.e., find lots of ``good'' issues in a relatively short time)
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 3.3)

\item  EIAM system is easy to use.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.2)

\item  EIAM system is useful. 
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.0)

\item  My overall satisfaction with EIAM system was:
\\
Very low \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very high\\
(Score: 3.9)

\item  Problems that I had with EIAM system (please explain): 
\begin{itemize}
\item It is very hard to decide the code is right or wrong.  Sometimes I
have hard time to find out the thing i want.
\item to choose proper subject
\item Figuring out which buttons to press for what.
\item I couldn't verify my doubts when I had questions on the  code.  
\item the system works okay and would prove very valuable to programmers
\item I was very distracted because another student in the lab broght to my
attention a misjudged grade from class.
\item None.
\item Looking at various windows simultaneously.  There isn't much room to
fit them all.  The node and issue windows could be a bit smaller.

\item none really, again interface was easy to use
\item I liked being able to talk the issues over with others.
\item Limited way of identifying issues.
\item insufficient training

\item some times I just wanted to change the whole algorithm of the program
code.
\item I often forget to close an issue before going to the next.  Also some
odd encounting of hight lighting
\end{itemize}


\item  Problems that I had with EIAM process (please explain): 
\begin{itemize}
\item ok
\item Understanding the logic of the code

\item None.
\item not much
\item None , but the distraction
\item I found that my confidence level in raising issues was lower than the
gropu session because in the group session I could raise something
that didn't seem right to me, but I wasn't sure about, and the others
would help decide if it was correct or not.  In the individual
process, I wasn't as secure in raising issues.
\item It's a bit too restrictive.  I wish that we could confer with each
other, at least for what a particular operating system would do for
a statement.
\item none

\item Lack of confidence on some complexities.
\item Requires a lot of concentration, and if there is excess noise in the
lab, it is hard to follow the code.
\item had to keep in mind the corrections that had been made previously. In
other words, the result of debugging is not visible through the entire process.
\item None
\end{itemize}


\item  Suggestions on how to improve EIAM system (please explain):
\begin{itemize}
\item good
\item I don't know if this is possible, but having a flow chart along with
all the functions would be a great help for me.

\item Automatically refresh summary buffer after each function is reviewed.
\item provide some type of reference for better quality analysis'.

\item maybe to improve the fonts to make the words bigger,
eventhough the programmer is just allowed to create issue
the programmer should be allowed to test the code
\item None.



\item Better layout of the screen.  Not so many clicks for a function.  The
voting issues in the issue screen could use additonal options, such as
"could be better" in the criticallity options.
\item none
\item Pair screens for related functions - object file paired with its functions.
\item Provide space to test.  Hard to follow code just by examining text.
\item Non;  not at the present moment.
\item It will be nice to have automatic closure of issue when a new one is created.
\end{itemize}

\item  Suggestions on how to improve EIAM process (please explain):
\begin{itemize}
\item if possible, offer correct solution after creating the proper issue.
\item combine both the individual and group EIAM process
\item None.
\item There should be a conferring part where we could verify that we are
thinking along the same lines.
\item none
\item Make sure lab is completely quiet when setting up the session.
\end{itemize}


\item  Other comments:
\begin{itemize}
\item give some of standard subjects to be choosen
\item In general, I think that this EIAM is very useful when it comes to
source code review. One of the attribute that I like most is that you
can jump between functions rather easily.

\item To me, the source code was kind of difficult, although we already did
this assignment.  I didn't review the material very much before the review.
\item I could have been much more effective if I were not distracted, once I
found out my grade I could think of nothing else!



\item I think this is a valuable debugging tool, but it is most valuable
when used in conjunction with the group session.  I think it would
probably be most efficient if the individual session was done first so
all the members would already be familiar with the code and could
contribute more to the sessions. (This is what I personally believe,
however, I can think of arguments for doing it the other way around as well)
\item This process is too impersonal.  I could imagine myself talking to no
one and getting frustrated while at it.  Also, during the test, I
found that I always suspected there was an error, even though I'm more
than 95% sure there wasn't.  Thus, I spent too much time looking for
something that wasn't there.
\item i think this is a good idea and it should be used however i believe
there should be a stepper/debugger to help debug the code as well
as using this method! i tend to actually step through the code on
paper while working (keeping track of the variables etc)
on this so the stepper would be useful.


\item None.
\item I enjoyed the reviewing thoroghly
\item Special Thanks to the TA.
\item I do think this process may not help me very much on improving my skill of
C++, although it does offer me an excellent medium to review my code
and debug my program.  I often do the some thing on paper.  There is
no boubt a need to do programe review.  This is a good system to go.
But I may not help me directly very much.  When you are in the wood,
all you see are just trees, not the forest!  You just have to have the
skills for you to see potential problems.  A good medium of reviewing
may not make you any smarter.

Good luck and Happy hacking...

\end{itemize}


\end{enumerate}

