\chapter{Conclusion}
\label{cha:conclusion}

This proposal laid out a research plan to investigate the information technology infrastructure that can support effective and efficient development of serious games for sustainability. The research includes the development of an innovative serious game framework for sustainability that combining education and behavior change, and an assessment method accessing the effectiveness and efficiency of the IT infrastructure for serious games for sustainability with regarding the most important stakeholder's perspective.

\section{Anticipated Contributions}

The anticipated contributions of this research are:

\begin{itemize}
	\item Makahiki: open source information technology for development of serious games for sustainability.
	\item SGSEAM: an assessment method for assessing serious game framework.
	\item Evidence regarding the effectiveness and efficiency of Makahiki as a framework for development of serious games for sustainability.
	\item Evidence regarding the effectiveness and efficiency of a second system (BuildingOS) as a framework for development of serious games for sustainability.
	\item Insights into the strengths and weaknesses of the assessment method.
\end{itemize}

\section{Future Directions}

There are a variety of directions that can be pursued once this research is complete, such as:

\begin{itemize}
	\item Evaluate the other stakeholders’ experiences

    \item Build a community to expand content and game library for Makahiki

    \item Scale and expand Makahiki to support other geographical and cultural different locations.

\end{itemize}

\section{Time line}

I started the research in September 2010 with the development of an initial version of Makahiki framework. Table \ref{table:completed} lists the completed tasks for this research:

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{\centering\tabhead{Time}} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{\centering\tabhead{Task}} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{September 2010} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{Makahiki v1 development} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{October 2011} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{First UH Kukui Cup} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{October 2011} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{Makahiki v2 development} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{September 2012} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{HPU, EWC, UH Kukui Cup challenge} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{March 2013} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{evaluation of Makahiki using ICS691} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{September 2013} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{preliminary data analysis} \\
    \hline
  \end{tabular}
  \caption{Completed Tasks}
  \label{table:completed}
\end{table}

Table \ref{table:pending} lists the planed tasks for the proposed continuing research:

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{\centering\tabhead{Time}} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{\centering\tabhead{Task}} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{November 2013} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{post-hoc system admin, game designer, game manager assessment using HPU 2013 instance} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{February 2014} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{player self-reported and engagement metrics assessment using UH 2014 instance} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{March 2014} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{post-hoc developer assessment using UH 2014 instance} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{April 2014} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{Lucid BuildingOS assessment using CCN 2014 instances} \\
    \hline
    \multicolumn{1}{|p{0.2\columnwidth}|}{August 2014} &
    \multicolumn{1}{|p{0.7\columnwidth}|}{dissertation defense} \\
    \hline
  \end{tabular}
  \caption{Planed Tasks}
  \label{table:pending}
\end{table}
