\chapter{In-Game Questionnaire}
\label{app:in-game-questionnaire}

This appendix details the contents of the questionnaire made available to participants via the in-game questionnaire during the challenge.

When participants filled out the questionnaire via the SurveyGizmo~\cite{surveygizmo} website, the questions were broken into pages. Each page provided participants the ability to move forward to the next page in the questionnaire, but not back to previous pages. 

\section{Gamification}

This section primarily covers questions about the game aspects of the challenge.

\vspace{5 mm}
\noindent
22. Have you made any commitments through the website during the game?

\begin{radiobutton}
	\item Yes
	\item No
\end{radiobutton}

\noindent
[If Yes] Did you change your behavior during the competition based on the commitment(s) you made?

\begin{radiobutton}
	\item Yes
	\item No
	\item Not sure
\end{radiobutton}

\vspace{5 mm}
\noindent
23. Which of the followings Kukui Cup achievements would you want to share on Facebook? (choose all that apply)

\begin{checkbox}
	\item made a commitment
	\item participated in an activity
	\item attended an event or excursion
	\item earned a badge
	\item current leader in the scoreboard
	\item other
\end{checkbox}

\vspace{5 mm}
\noindent
24. How much time do you usually spend on the following activities?

[Options for each activity:]

\begin{enumerate}
	\item 3 or more hours a day
	\item about 1 hour a day
	\item about 1 hour a week
	\item 1 hour a month or less
	\item never
\end{enumerate}

\noindent
List of activities:

\begin{itemize}
	\item Playing games on a laptop computer
	\item Playing games on a game console (Xbox, PS3, Wii)
	\item Playing games on a handheld game device (DS3, PSP)
	\item Playing games on a mobile phone
	\item Checking Facebook
	\item Checking Twitter
\end{itemize}

\vspace{5 mm}
\noindent
25. How would you describe the Kukui Cup? (check all that apply)

\begin{checkbox}
	\item Fun
	\item Educational
	\item So-so
	\item Boring
	\item Not useful
	\item Difficult
	\item Addictive
	\item Other
\end{checkbox}

\vspace{5 mm}
\noindent
26. The Kukui Cup website shows energy data updated every 15 seconds. Did you find this helpful in conserving energy?

\begin{radiobutton}
	\item not really, updating the data daily would be enough
	\item not really, updating the data hourly would be enough
	\item not really, I only care about the final result of the competition
	\item yes, it is helpful to see the energy usage changing in real time
\end{radiobutton}

\vspace{5 mm}
\noindent
27. Which of the following do you wish there were more of in the game? (choose all that apply)

\begin{checkbox}
	\item events
	\item excursions
	\item commitments
	\item videos
	\item social activities
	\item physical activities
	\item online activities
\end{checkbox}

\vspace{5 mm}
\noindent
28. On average, how many minutes a day did you spend on the Kukui Cup website?

\vspace{5 mm}
\noindent
29. On average, how many hours a week did you spend at Kukui Cup events?


\section{General Feedback}

This section covers feedback on the challenge overall.

\vspace{5 mm}
\noindent
30. What can we do to improve participation in the Kukui Cup website?

\vspace{5 mm}
\noindent
31. What was the best thing you liked about the Kukui Cup so far?

\vspace{5 mm}
\noindent
32. What was the thing you liked the least about the Kukui Cup so far?

\vspace{5 mm}
\noindent
33. If you were able to play the Kukui Cup next year, would you?

\begin{radiobutton}
	\item Yes
	\item I enjoyed it, but I wouldn't play again
	\item I didn't enjoy it, and I wouldn't play again
	\item No, because: \underline{\hspace{5cm}}
\end{radiobutton}

\vspace{5 mm}
\noindent
34. How likely would you be to recommend playing the Kukui Cup to a first year student in Fall 2012?

\begin{radiobutton}
	\item Very Likely
	\item Likely
	\item Neutral
	\item Unlikely
	\item Very Unlikely
	\item Not Applicable
\end{radiobutton}

\vspace{5 mm}
\noindent
35. Is there anything else you would like to tell us about your experience playing the Kukui Cup that this survey didn't ask?
