\chapter{Conclusion}
\label{cha:conclusion}

This research investigates the information technology infrastructure that can support effective and efficient development of serious games for sustainability. The research includes the development of an innovative serious game framework for sustainability that combining education and behavior change, and an assessment method accessing the effectiveness and efficiency of the IT infrastructure for serious games for sustainability with regarding the most important stakeholder's perspective.

\section{Contributions}

The contributions of this research are:

\begin{itemize}
	\item Makahiki: open source information technology for development of serious games for sustainability.
	\item SGSEAM: an assessment method for assessing serious game framework.
	\item Evidence regarding the effectiveness and efficiency of Makahiki as a framework for development of serious games for sustainability.
	\item Evidence regarding the effectiveness and efficiency of a second system (BuildingOS) as a framework for development of serious games for sustainability.
	\item Insights into the strengths and weaknesses of the assessment method.
\end{itemize}

\section{Future Directions}

There are a variety of directions that can be pursued once this research is complete, such as:

\begin{itemize}
	\item Evaluate the other stakeholders’ experiences

    \item Build a community to expand content and game library for Makahiki

    \item Scale and expand Makahiki to support other geographical and cultural different locations.

\end{itemize}

