\chapter{Publication List}
\label{app:publication-list}

These are the publications that have come out of this research that I have authored or co-authored:

\section{Journal Paper}

\begin{itemize}

\item Robert S. Brewer, \textbf{Yongwen Xu}, George E. Lee, Michelle Katchuck, Carleton A. Moore, Philip M. Johnson. Three Principles for the Design of Energy Feedback Visualizations. In   
\emph{International Journal On Advances in Intelligent Systems}, Vol. 3 \& 4, No. 6. (2013), pp. 188-198

\end{itemize}

\section{Conference Papers}

\begin{itemize}

\item \textbf{Yongwen Xu}, Philip M. Johnson, George E. Lee, Carleton A. Moore, Robert S. Brewer. Makahiki: An open source serious game framework for sustainability education and conservation.   
In \emph{Proceedings of the 2014 International Conference on Sustainability, Technology, and Education (STE 2014)}, Taiwan, December 2014
 	
\item \textbf{Yongwen Xu}, Philip M. Johnson, Carleton A. Moore, Robert S. Brewer, Jordan Takayama. SGSEAM: Assessing serious game frameworks from a stakeholder experience perspective.   
In \emph{Proceedings of the First International Conference on Gameful Design, Research, and Applications (Gamification 2013)}, Ontario, Canada, October 2013

\item Robert S. Brewer, \textbf{Yongwen Xu}, George E. Lee, Michelle Katchuck, Carleton A. Moore, and Philip M. Johnson. Energy feedback for smart grid consumers: Lessons learned from the Kukui Cup. In \emph{Proceedings of the Third International Conference on Smart Grids, Green Communications and IT Energy-aware Technologies (ENERGY 2013)}, Lisbon, Portugal, 

March 2013.

\item Philip M. Johnson, \textbf{Yongwen Xu}, Robert S. Brewer, Carleton A. Moore, George E. Lee, and Andrea Connell. Makahiki+WattDepot: An open source software stack for next generation energy research and education. In \emph{Proceedings of the 2013 Conference on Information and Communication Technologies for Sustainability (ICT4S)}, Zurich, Switzerland, February 2013.

\item Philip M. Johnson, \textbf{Yongwen Xu}, Robert S. Brewer, George E. Lee, Michelle Katchuck, and Carleton A. Moore. Beyond kWh: Myths and fixes for energy competition game design. In \emph{Proceedings of Meaningful Play 2012}, October 2012.

\end{itemize}


\section{Workshop}

\begin{itemize}

\item Robert S. Brewer, George E. Lee, \textbf{Yongwen Xu}, Caterina Desiato, Michelle Katchuck, and Philip M. Johnson. Lights Off. Game On. The Kukui Cup: A dorm energy competition. In \emph{Proceedings of the CHI 2011 Workshop on Gamification}, Vancouver, Canada, May 2011.

\item \textbf{Yongwen Xu}. Designing a Serious Game Framework for Sustainability. In \emph{PhD Student  Workshop on ICT4S 2013}, Zurich, Switzerland, February 2013.

\end{itemize}

\section{Poster}

\begin{itemize}

\item  Robert S. Brewer, Philip M. Johnson, Michelle Katchuck, George E. Lee, \textbf{Yongwen Xu}. Lights Off. Game On. The 2011 Kukui Cup. In \emph{The Behavior, Energy and Climate Change Conference (BECC) 2011}, Washington DC, November 2011.

\end{itemize}

