%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% budget.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Jul 27 13:31:23 1994
%% Last Modified By: Philip M. Johnson
%% Last Modified On: Tue May 24 08:08:31 2005
%% Status          : Unknown
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\documentclass[11pt]{article} 
\usepackage{/export/home/csdl/tex/icse2003/latex8}
\usepackage{times}
\usepackage[final]{graphicx}
% uncomment the % away on next line to produce the final camera-ready version
% and uncomment the \thispagestyle{empty} following \maketitle
%\pagestyle{empty}



\begin{document}


\pagestyle{empty}
\section*{Budget Justification}

\subsection*{Salaries}

The project budget provides salary support for the principal investigator
(two summer months) and three graduate research
assistants (11 months) for each of the four years.  The principal
investigator will perform or supervise all major system design enhancements
and empirical studies, manage and train the graduate student
assistants, and supervise case studies. 

\subsection*{Travel}

The project budget provides funds for four trips per year from Hawaii to the
mainland.  Two trips per year will be to attend meetings with the other 
principle investigators.  The remaining trips will be to attend
relevant conferences (e.g., ICSE, FSE, ISERN, or STAR), where
he may report on the results of his own work and obtain first hand
information on related work. 

\subsection*{Computer Services}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise SUN server and
8 workstations, along with a printer and networking
peripherals.  The project budget provides funds of \$15,600 to purchase a
new enterprise level server (Dell 2850 with RAID, backup and associated
software and peripherals or equivalent) for Cedar in the first year of the
grant.  In addition, \$2,500 per year per developer is budgeted for
``supplies'', which include workstation purchases and upgrades, software,
and associated peripherals.


\subsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
Summer salary is calculated as two additional months of the principal
investigator's annual nine month salary.  Summer salary includes the pay
increases as negotiated through collective bargaining.  Pay increases are
as follows: 5\% in 2006 (year 1), 9\% in 2007 (year 2) and 11\% in 2008
(year 3).
 

\paragraph*{Research Assistants.}  
This cost category provides support for three graduate research assistants
for each of the four years.  The salary for each research assistant is
based upon RA-Step 3.

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 3.5\% for the principal
investigator and 25\% for the graduate assistants.

\paragraph*{Overhead.}  
University of Hawaii overhead cost is calculated as 36.3\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).





\end{document}
