\chapter {Error List}

\section{Driver Class}
\subsection{Driver.C}
\begin {enumerate}
\item Driver::Driver -$>$ Memory leak. Statement name="" overwrite the newly
allocated space.
\item Driver::change -$>$ Memory leak. Should call explicit strcpy (name, temp\_name)
\item Driver::~Driver -$>$ Destructor can't be empty. Should free or
delete name.
\item main -$>$ Wrong inequality in userinput\_shift $>$ 0 \&\&
userinput\_shift$<$4
\item main -$>$ delete [] userinput\_driver should be inside the while loop.
\item main -$>$ cin.get(tempchar) wrong place;should be outside the
switch statement. 
\item main -$>$ new char [MAXLENGTH] should have been [MAXLENGTH+1].
\end{enumerate}


\section{Yard Class}

\subsection{Driver.C}
\begin {enumerate}
\item driver::create -$>$ Should initialize years\_exp to -1 (negative
number). 
\item driver::create -$>$ Should not call cin.get(tempchar), first
character may be missing (flushing standard input should not be done
here). 
\item driver::create -$>$ Bogus test for strlen $>$ MAXLEN, since the
string is read up to MAXLENGTH.
\item driver::create -$>$ Should check if strlen $>$ MAXLEN after the
getline, if so flush standard input.
\item driver::create -$>$ Should have an operation to flush standard
input (after the last cin). 
\item driver::create -$>$ Memory leak: should deallocate userString.
\item driver::create -$>$ invoking strlen(userString) when userString
is still uninitialized.

\end{enumerate}

\subsection{Vehicle.C}
\begin {enumerate}
\item vehicle::create -$>$ Should check for valid value for seating
capacity.
\item vehicle::create -$>$ Should flush standard input after reading
integer. 
\end{enumerate}

\subsection{Bus.C}
\begin {enumerate}
\item bus::create -$>$ Should flush standard input after reading integer.
\item bus::print -$>$ Missing vehicle::print
\end{enumerate}

\subsection{Van.C}
\begin {enumerate}
\item van::create -$>$ Using strlen from newly allocated regionString is
incorrect
\item van::create -$>$ extra cin.get (tempchar): should be put in
vehicle::create to flush standard input.
\item van::create -$>$ Should deallocate old value of region before setting to
regionString.
\item van::create -$>$ Should not call cin.get(tempchar), first
character may be missing (flushing standard input should not be done
here). 
\item van::create -$>$ Bogus test for strlen $>$ MAXLEN, since the
string is read up to MAXLENGTH.
\item van::create -$>$ Should check if strlen $>$ MAXLEN after the
getline, if so flush standard input.
\item van::create -$>$ invoking strlen(regionString) when regionString
is still uninitialized.

\end{enumerate}

\subsection{Yard.C}
\begin{enumerate}
\item ~yard -$>$ Should loop until the number of vehicles instead of YARD\_SIZE
\item yard::add -$>$ Variable no\_of\_veh is not updated.
\item yard::remove -$>$ The inequality should be trash\_veh $<$= 0. 
\item yard::remove -$>$ The last yard/ parking [j] is not set to NULL.
\item main -$>$ Missing mybus = new bus in case 1
\item main -$>$ Missing myvan = new van in case 2
\item main -$>$ Missing mybus = new bus in case 5
\item main -$>$ Missing myvan = new van in case 6
\end{enumerate}


