%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% budget.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Jul 27 13:31:23 1994
%% Last Modified By: 
%% Last Modified On: Sat Sep 30 13:59:07 2006
%% Status          : Unknown
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\documentclass[11pt]{article} 
\usepackage{/export/home/csdl/tex/icse2003/latex8}
\usepackage{times}
\usepackage[final]{graphicx}
% uncomment the % away on next line to produce the final camera-ready version
% and uncomment the \thispagestyle{empty} following \maketitle
%\pagestyle{empty}


\begin{document}


\pagestyle{empty}
\section*{Budget Justification}

\subsection*{Salaries}

The project budget provides salary support for the principal investigator
(two summer months) and two graduate research
assistants (11 months) for each of the three years.  The principal
investigator will perform or supervise all major system design enhancements
and empirical studies, manage and train the graduate student
assistants, and supervise case studies. 

\subsection*{Travel}

The project budget provides funds for two trips per year from Hawaii to the
mainland.  These trips will be used to attend
relevant conferences (e.g., ICSE, FSE, ISERN, or STAR), where
he may report on the results of his own work and obtain first hand
information on related work. 

\subsection*{Computer Services}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise SUN server and
8 workstations, along with a printer and networking
peripherals.  The project budget provides funds of \$10,000/year to 
provide workstations and associated software for the principal investigator and 
the two graduate research assistants. In addition, these funds will maintain 
a server for the public Hackystat site and the developer services site. 

\subsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
Summer salary is calculated as two additional months of the principal
investigator's annual nine month salary.  Summer salary includes the pay
increases as negotiated through collective bargaining. 
 

\paragraph*{Research Assistants.}  
This cost category provides support for two graduate research assistants
for each of the four years.  The salary for each research assistant is
based upon RA-Step 3.

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 3.5\% for the principal
investigator and 25\% for the graduate assistants.

\paragraph*{Overhead.}  
University of Hawaii overhead cost is calculated as 38.4\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).





\end{document}
