%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% summary.tex -- 
%% RCS:            : $Id: nsf93-summary.tex,v 1.11 93/10/06 16:52:35 johnson Exp $
%% Author          : Philip Johnson
%% Created On      : Wed Aug 11 12:55:46 1993
%% Last Modified By: 
%% Last Modified On: Mon Oct 09 11:01:16 2006
%% Status          : Unknown
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1993 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%% History
%% 11-Aug-1993          Philip Johnson  
%%    

\documentclass[11pt]{article} 
\usepackage{/export/home/csdl/tex/icse2003/latex8}
\usepackage{times}

\pagestyle{empty}

\begin{document}
\section*{Project Summary}

This research proposal presents a new, evidence-based approach to the
discovery and evaluation of software engineering ``best'' practices.
Instead of looking outward into the community for best practices, and
attempting to adapt them to one's own environment, our research will
investigate how best practices can emerge organically from within one's
current organizational and project context.  Instead of relying on politics
or persuasiveness for adoption, our research approach involves
instrumentation that generates empirical data that can be used to either
argue for the benefits of adoption of the practice, or else provide
evidence that the practice is not effective in the current
context.  To accomplish this, we will synthesize and extend several streams
of research: (a) software project telemetry, which provides a mechanism for
in-process monitoring of software engineering data streams; (b) software
development stream analysis (SDSA), a mechanism for automated recognition
of certain best practices from low-level developer behaviors, and (c)
evidence-based software engineering, an approach to better structured and
more useful empirically-based research.  Our project has six objectives:

  
(1) Enhancement of the Software Development Stream Analysis mechanism to
  support validation and a variety of best practices, and determination of the
  kinds of abstractions, automation, and best practices that are amenable
  to recognition using SDSA.

(2) Development of integration mechanisms between SDSA and Software
  Project Telemetry in order to allow users to determine how practices
  recognized by SDSA relate to process and product outcomes. 
  
  
(3) Classroom-based evaluation of the proposed techniques. We will 
apply these techniques to gain evidence regarding programmer productivity and variability 
with respect to the Test Driven Design best practice. This activity will also 
refine the technology, develop curriculum materials, and ready the approach for 
industrial evaluation. 

(4) Industry-based evaluation of the proposed techniques. Following
classroom evaluation, we will carry out two industry-based case studies to
gather evidence regarding best practices related to high performance
computing and agile software development. This activity will help us
understand the strengths and limitations of this technique by applying it
to ``real world'' settings in two very different domains.
  
(5) Packaging of the system and methods for widespread dissemination.  We
will continue the process used by the open source Hackystat Project of making
our technology available to the software engineering community.  In addition, 
we will package and disseminate our experimental methods to support external
evidence-based software engineering efforts. 
  
(6) Development of curriculum materials regarding continuous,
  evidence-based discovery and assessment of software engineering best
  practices. As with the Hackystat Project, we will develop software engineering
  curriculum materials and assignments that enable the study and analysis
  of this approach in academic settings.

The intellectual merit of this research includes the application of novel
data gathering and analysis techniques for the discovery and evaluation of 
software engineering best practices, and the evaluation of this technique
through classroom and industrial case studies. 

The broader impact of this research includes the development of a
sophisticated, freely available, open source software system for use by
researchers and practitioners to study software engineering best practices,
the generation of new evidence-based research results, and curriculum
materials to support education and technology transfer.  As the University
of Hawaii is a university with 75\% minority students in an EPSCOR state,
this project will provide novel research opportunities to underrepresented
groups.

\end{document}

 







