%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% bio.johnson.tex -- 2011 Sustainable Energy Pathways Proposal
%% Author          : Philip Johnson
%% Created On      : Tue Nov  4 10:26:48 1997
%% Last Modified By: Philip Johnson
%% Last Modified On: Fri Jan 27 13:01:01 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
Philip Johnson \\
Information and Computer Sciences \hfill (808) 956-3489

University of Hawaii              \hfill fax: (808) 956-3548

1680 East-West Road               \hfill johnson@hawaii.edu

Honolulu, HI~~~96822              \hfill http://www.ics.hawaii.edu/$\sim$johnson/

\end{center}

\begin{Professional Preparation}
\item Ph.D.~in Computer Science, University of Massachusetts. 1990 
\item M.S.~in Computer Science, University of Massachusetts.  1985
\item B.S.~in Biology, University of Michigan. 1980
\item B.S.~in Computer Science, University of Michigan. 1980
\end{Professional Preparation}

\begin{Appointments}
\item {\em Professor},  Information and Computer Sciences, University of Hawaii.  2001---present
\item {\em Visiting Professor}, School of Engineering, Blekinge Institute of Technology, Sweden, 2006.
\item {\em Associate Professor},  Information and Computer Sciences, University of Hawaii.  1995---2001
\item {\em Senior Research Fellow},  Distributed Systems Technology Centre, University of Queensland, Australia, 1997.
\item {\em Assistant Professor},   Information and Computer Sciences, University of Hawaii.  1990---1995
\end{Appointments}


\begin{Publications: Closely Related}

\item Robert S. Brewer, Philip M. Johnson, 
{\em WattDepot: An open source software ecosystem for enterprise-scale
  energy data collection, storage, analysis, and visualization}, 
Proceedings of the First International Conference on Smart Grid
Communications, Gaithersburg, MD, 2010.

\item George E. Lee, Yongwen Xu, Robert S. Brewer, and Philip
  M. Johnson, {\em Makahiki: An Open Source Game Engine for Energy
    Education and Conservation},
  Submitted to Foundations of Digital Games 2012, Raleigh, North Carolina,
  2012.

\item Robert S. Brewer, George E. Lee, Yongwen Xu, Caterina
  Desiato, Michelle Katchuck, and Philip M. Johnson, 
{\em Lights Off. Game On. The Kukui Cup: A Dorm Energy Competition},
Proceedings of the CHI 2011 Workshop on Gamification, Vancouver, Canada,
2011.

\item Robert S. Brewer, George E. Lee, and Philip M. Johnson, 
{\em The Kukui Cup: a Dorm Energy Competition Focused on Sustainable
  Behavior Change and Energy Literacy}, 
Proceedings of the 44th Hawaii International Conference on System Sciences, 2011.


\end{Publications: Closely Related}


\begin{Publications: Other Significant}

\item H. Kou and P. Johnson and H. Erdogmus, {\em Operational Definition and 
Automated Inference of Test-Driven Development with Zorro}, 
Journal of Automated Software Engineering, December, 2009.

\item P. Johnson and S. Zhang, {\em We need more coverage, stat!  Experiences with the Software ICU}, 
Proceedings of  Empirical Software Engineering and Measurement, October, 2009.
  
\item V. Basili and M. Zelkowitz and D. Sjoberg and P. Johnson and T. Cowling,
{\em Protocols in the use of Empirical Software Engineering Artifacts}, 
Empirical Software Engineering, Volume 12, February, 2007.

\item P. Johnson, {\em Requirement and Design Trade-offs in Hackystat: An
in-process software engineering measurement and analysis system},
Proceedings of the 2007 International Symposium on Empirical Software
Engineering and Measurement, Madrid, Spain, September, 2007.

\item L. Hochstein and T. Nakamura and V. Basili and S. Asgari and 
M. Zelkowitz and J. Hollingsworth and F. Shull and J. Carver and 
M. Voelp and N. Zazworka and P. Johnson, {\em Experiments to 
understand HPC time to development}, CTWatch Quarterly, 
November, 2006.

\end{Publications: Other Significant}


\begin{Synergistic Activities}
\item {\em Project Director}, WattDepot Framework, http://wattdepot.googlecode.com/
\item {\em Project Director}, Kukui Cup, http://www.kukuicup.manoa.hawaii.edu/
\item {\em Co-founder and interim Chief Technology Officer}, HotU, Inc.  2000.
\item {\em Editorial Board}, IEEE Transactions on Software Engineering, 2000-2004.
\item {\em Program Chair}, International Software Engineering Research Network, Annual Meeting, 2000.
\item {\em Chair}, Hawaii Java Users Group, 1996-2005

\end{Synergistic Activities}


\begin{Collaborators and Co-Editors}
\item Victor Basili (University of Maryland),
Jeff Carver (University of Alabama),
Hackan Erdogmus (National Research Council, Canada), 
Stuart Faulk (University of Oregon),
Brian Pentland (Michigan State University),
Dan Port (University of Hawaii),
Forrest Shull (Fraunhofer-Maryland),
Larry Votta (Sun Microsystems),
Claes Wohlin (Blekinge Institute of Technology, Sweden),
Marvin Zelkowitz (University of Maryland)
\end{Collaborators and Co-Editors}


\begin{Graduate Advisors and Postdoctoral Sponsors}
\item Robert Brewer, Yongwen Xu, Michelle Katchuck, George Lee,
Monir Hodges,
Aaron Kagawa, 
Hongbing Kou,
Christoph Lofi,
Michael Paulding,
Carleton Moore,
Mette Moffett,
Alexey Olkov,
Pavel Senin, 
Danu Tjahjono,
Dadong Wan,
Takuya Yamashita,
Cedric Zhang, 
Shoaxuan Zhang
\end{Graduate Advisors and Postdoctoral Sponsors}

\begin{Thesis Advisor and Postgraduate-Scholar Sponsor}
\item Jack Wileden, 
Victor Lesser, 
Wendy Lehnert, 
David Stemple (deceased), 
Daniel Corkill (University of Massachusetts)
\end{Thesis Advisor and Postgraduate-Scholar Sponsor}

\end{document}












