%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 10-07-future.tex --  HICSS 44 Kukui Cup paper
%% Author          : Philip Johnson
%% Created On      : Mon Sep 23 11:52:28 2002
%% Last Modified By: Philip Johnson
%% Last Modified On: Mon Jun 14 12:55:47 2010
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2009 Philip Johnson
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{Future Directions}
\label{sec:future-directions}

As noted above, we are currently planning for the inaugural Kukui Cup
competition to take place in October, 2010.  In addition, we are planning
several other extensions to this research.

First, we are interested in adapting this technology and research approach
to the issue of residential energy consumption.  Residential energy
consumption differs in many ways from dorm energy consumption. Most
significantly, residential energy users are billed directly for the energy
use, so there is an important financial incentive for conservation.
However, the research indicates that tools such as goals and public
commitments remain important in the residential energy application domain.

Second, we would like to support the use of Makahiki as a framework for
other universities who desire to implement a dorm energy competition.  We
believe that Makahiki can significantly lower the ``barrier to entry'' for
universities, and allows them to get started with a simple site if required. 

Third, we are excited by the possibility of integrating grid-level
information into WattDepot for use by Makahiki and other high level
applications.  For example, we are working with our local utility to
provide WattDepot with information on the current mix of generation sources
(i.e. coal, oil, nuclear, solar, wind, etc.) and their aggregate output
(i.e. grid load).  Given this information, it would be possible for
WattDepot to provide the carbon intensity associated with the energy
consumed by end-users at any point in time.  This would make it possible
for the Kukui Cup to incentivize behaviors such as ``doing laundry when the
carbon intensity of the grid is low.''





