%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% bio.tex -- 
%% Author          : Philip Johnson
%% Created On      : Tue Nov  4 10:26:48 1997
%% Last Modified By: Philip Johnson
%% Last Modified On: Thu Dec 10 14:58:10 2009
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1997 Philip Johnson
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

%\section*{Biographical Sketch}
%\pagenumbering{arabic}
%\renewcommand{\thepage} {E--\arabic{page}}


%%      Command to define new categories:
\newcommand{\newcategory}[1]{\newenvironment{#1}
 {\sectionheading{#1}\begin{list}{}{\setlength{\labelwidth}{0cm} \setlength{\labelsep}{0cm} \setlength{\itemsep}{0ex plus0.2ex} \setlength{\itemindent}{-1cm} \setlength{\leftmargin}{1cm} \setlength{\parsep}{0ex plus0.2ex}}}{\end{list}\par}}

\newcommand{\sectionheading}[1]{\medskip\pagebreak[2]\par\noindent
 {\small\bf #1}\nopagebreak}

\begin{center}
Philip M. Johnson\\
Information and Computer Sciences \hfill (808) 956-3489

University of Hawaii              \hfill fax: (808) 956-3548

1680 East-West Road               \hfill johnson@hawaii.edu

Honolulu, HI~~~96822              \hfill http://philipmjohnson.org

\end{center}

\newcategory{Professional Preparation}
\begin{Professional Preparation}
\item Ph.D.~in Computer Science, University of Massachusetts. 1990 
\item M.S.~in Computer Science, University of Massachusetts.  1985
\item B.S.~in Biology, University of Michigan. 1980
\item B.S.~in Computer Science, University of Michigan. 1980
\end{Professional Preparation}

\newcategory{Appointments}
\begin{Appointments}
\item {\em Professor},  Information and Computer
  Sciences, University of Hawaii.  2001---present

\item {\em Visiting Professor}, School of Engineering, Blekinge Institute of Technology, Sweden, 2006.

\item {\em Associate Professor},  Information and Computer
  Sciences, University of Hawaii.  1995---2001

\item {\em Senior Research Fellow},  Distributed Systems Technology Centre,
University of Queensland, Australia, 1997.

\item {\em Assistant Professor},   Information and Computer
  Sciences, University of Hawaii.  1990---1995

\end{Appointments}




\newcategory{Publications: Closely Related}
\begin{Publications: Closely Related}

\item R.Brewer, Y. Xu, G. Lee, M. Katchuck, C. Moore, P. Johnson, {\em Three principals for the design of energy feedback visualizations}, International Journal on Advances in Intelligent Systems,Volume 6, No. 3, December 2013.


\item R. Brewer, Y. Xu, G. Lee, M. Katchuck, C. Moore, P. Johnson, {\em Energy feedback for smart grid consumers: Lessons learned from the Kukui Cup}, Proceedings of Energy 2013, Lisbon, Portugal, March 2013.


\item P. Johnson, Y. Xu, R. Brewer,  C. Moore,  G. Lee, A. Connell, {\em Makahiki+WattDepot: An open source software stack for next generation energy research and education}, Proceedings of the 2012 Conference on Information and Communication Technologies for Sustainability, Zurich, Switzerland, February 2013.

\item P. Johnson, Y. Xu, R. Brewer, G. Lee, M. Katchuck, C. Moore, {\em Beyond kWh: Myths and fixes for energy competition game design}, Proceedings of Meaningful Play 2012, Lansing, Michigan, October 2012.

\item R. Brewer and P. Johnson, {\em WattDepot: An open source software ecosystem for enterprise-scale energy data collection, storage, analysis, and visualization}, Proceedings of the First IEEE International Conference on Smart Grid Communications, Gaithersburg, Maryland, October 2010.


\end{Publications: Closely Related}


\newcategory{Publications: Other Significant}
\begin{Publications: Other Significant}

\item H. Kou and P. Johnson and H. Erdogmus, {\em Operational Definition and 
Automated Inference of Test-Driven Development with Zorro}, 
Journal of Automated Software Engineering, December, 2009.

\item P. Johnson and S. Zhang, {\em We need more coverage, stat!  Experiences with the Software ICU}, 
Proceedings of  Empirical Software Engineering and Measurement, October, 2009.
  
\item V. Basili and M. Zelkowitz and D. Sjoberg and P. Johnson and T. Cowling,
{\em Protocols in the use of Empirical Software Engineering Artifacts}, 
Empirical Software Engineering, Volume 12, February, 2007.

\item P. Johnson, {\em Requirement and Design Trade-offs in Hackystat: An
in-process software engineering measurement and analysis system},
Proceedings of the 2007 International Symposium on Empirical Software
Engineering and Measurement, Madrid, Spain, September, 2007.

\item L. Hochstein and T. Nakamura and V. Basili and S. Asgari and 
M. Zelkowitz and J. Hollingsworth and F. Shull and J. Carver and 
M. Voelp and N. Zazworka and P. Johnson, {\em Experiments to 
understand HPC time to development}, CTWatch Quarterly, 
November, 2006.

\end{Publications: Other Significant}


\newcategory{Synergistic Activities}
\begin{Synergistic Activities}

\item {Project Leader}, Hackystat Framework, http://www.hackystat.org/.

\item {Co-founder and interim Chief Technology Officer}, HotU, Inc.  2000.

\item {\em Editorial Board}, IEEE Transactions on Software Engineering,
2000-2004.

\item {\em Program Chair}, International Software Engineering Research Network
 Annual Meeting, 2000.

\item {\em Chair}, Hawaii Java Users Group, 1996-present

\end{Synergistic Activities}

\renewcommand{\newcategory}[1]{\newenvironment{#1}
 {\sectionheading{#1}\begin{list}{}{\setlength{\labelwidth}{0cm} \setlength{\labelsep}{0cm} \setlength{\itemsep}{0ex plus0.2ex} \setlength{\itemindent}{0cm} \setlength{\leftmargin}{0cm} \setlength{\parsep}{0ex plus0.2ex}}}{\end{list}\par}}


\newcategory{Collaborators and Co-Editors}
\begin{Collaborators and Co-Editors}
\item Victor Basili (University of Maryland),
Jeff Carver (University of Alabama),
Hackan Erdogmus (National Research Council, Canada), 
Stuart Faulk (University of Oregon),
Brian Pentland (Michigan State University),
Dan Port (University of Hawaii),
Forrest Shull (Fraunhofer-Maryland),
Larry Votta (Sun Microsystems),
Claes Wohlin (Blekinge Institute of Technology, Sweden),
Marvin Zelkowitz (University of Maryland)
\end{Collaborators and Co-Editors}


\newcategory{Graduate Advisors and Postdoctoral Sponsors}
\begin{Graduate Advisors and Postdoctoral Sponsors}
\item Robert Brewer,
Yongwen Xu,
Monir Hodges,
Aaron Kagawa, 
Hongbing Kou,
Christoph Lofi,
Michael Paulding,
Carleton Moore,
Mette Moffett,
Alexey Olkov,
Pavel Senin, 
Danu Tjahjono,
Dadong Wan,
Takuya Yamashita,
Cedric Zhang, 
Shoaxuan Zhang
\end{Graduate Advisors and Postdoctoral Sponsors}

\newcategory{Thesis Advisor and Postgraduate-Scholar Sponsor}
\begin{Thesis Advisor and Postgraduate-Scholar Sponsor}
\item Jack Wileden, 
Victor Lesser, 
Wendy Lehnert, 
David Stemple (deceased), 
Daniel Corkill (University of Massachusetts)
\end{Thesis Advisor and Postgraduate-Scholar Sponsor}












