%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% budget.tex -- 
%% Author          : Philip Johnson
%% Created On      : Wed Jul 27 13:31:23 1994
%% Last Modified By: Philip Johnson
%% Last Modified On: Thu Dec 10 14:00:04 2009
%% Status          : Unknown
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 


\section*{Budget Justification}

%\pagenumbering{arabic}
%\renewcommand{\thepage} {F--\arabic{page}}

\subsection*{Salaries}

The project budget provides two summer months salary support for the principal investigator
and one summer month salary support for the two co-principal investigators for both years. 
The project budget supports three graduate students (GA 11 Step 3) for both years. 

The three principal investigators will perform or supervise all major system design enhancements and empirical studies, and manage and train the graduate student assistant.

\subsection*{Travel}

The project budget provides funds for three trips per year from Hawaii to the
mainland at \$3,000 per trip.  These funds will be used by the principal investigators or
graduate assistants to attend relevant conferences (BECC, ICSE) or CyberSEES project meetings.

\subsection*{Materials and Supplies}

The principal investigator directs the Collaborative Software Development
Laboratory, which is currently equipped with one enterprise server and several
workstations, along with a printer and networking peripherals.  

The project budget provides funds of \$15,000 in year 1 and \$10,000 in year two to provide materials and supplies related to this research, including: (a) computers and storage for research data; (b) software for data analysis; and (c) cloud-based services; (d) design and manufacturing of power quality meters.

\subsection*{Consultant}

The project budget provides funds of \$10,000 in year 1 to support independent review and testing of the power quality meters for safety and accuracy.

\subsection*{Cost breakdown}
\label{cost-breakdown}

All calculations are rounded to the nearest dollar.

\paragraph{Summer Salary.}  
The budget provides for two months of summer salary (\$26,222) for the principal investigator (Johnson),
one month summer salary for co-principal investigator Fripp (\$10,556), and one month
summer salary for co-principal investigator Spirandelli (\$8,222). 
A 2\% increase is projected for year 2.

\paragraph*{Research Assistants.}  
The budget provides for three graduate research assistantships
for each of the two years.  The salary for each research assistant is
based upon an 11 month RA-Step 3 salary (\$20,472), with a 
2\% increase in year 2.

\paragraph*{Fringe benefits.} 
Fringe benefits for salaries are calculated at 2.26\% for the principal
investigators and 9.34\% for the graduate assistants.

\paragraph*{Overhead.}  
University of Hawaii overhead cost is calculated as 41.5\% of Modified Total
Direct Costs (MTDC), i.e. salaries, travel, and supplies (equipment is
excluded).





