
\chapter {Final Questionnaire (ICS313) }

\begin{enumerate}

\item  I would rather review source code manually using pencil and paper than
using CSRS (on-line Collaborative Software Review System).
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 2.7)

\item  In general, CSRS is useful.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5 \hfill Very true\\
(Score: 4.2)

\item  Overall, I enjoyed using CSRS.
\\
Not at all true \hfill 1 \dotfill  2 \dotfill 3 \dotfill 4 \dotfill 5
\hfill Very true\\
(Score: 4.0)

\item I would:
    \begin{enumerate}
   \item[(1)]Strongly prefer using EGSM
   \item[(2)]Somewhat prefer using EGSM
   \item[(3)]Equally prefer using EGSM or EIAM
   \item[(4)]Somewhat prefer using EIAM
   \item[(5)]Strongly prefer using EIAM
   \end{enumerate}
(Score: 2.2)

\item I am:
    \begin{enumerate}
     \item[(1)] Much more productive using EGSM
     \item[(2)] Somewhat more productive using EGSM
     \item[(3)] Equally productive with EGSM or EIAM
     \item[(4)] Somewhat more productive using EIAM
     \item[(5)] Much more productive using EIAM
     \end{enumerate}   
(Score: 2.0)

\item  I prefer working:
   \begin{enumerate}
     \item[(1)] Alone 
     \item[(2)] Group
   \end{enumerate} 
(Score: 1.9)

\item  Reasons for preferring working in a group:
\begin{itemize}
\item 
Working in group provides various ideas from group members.  That
helps the rest of the group members to understand the material and
innovate new ideas.  Consequently, result in better software development.
\item I have learned that it is important to have someone to look over the
mistakes that I might make during the debugging procress.
\item  
we can discuss the ? each other
\item I could find more errors when I worked with my group although we had
some disagreements on certain topics each other.
\item it provides analysis from different angles. Others may see problems
that are obvious to them but not to the programmer.
\item if no clue, group ok, if need quiet thinking, prefer
individual. Depends the situations.
\item I always prfered to work in a group, If one mistake is missed by one
it may be found by another!!!!!!!

\item In a group setting, each member can contribute their knowledge to the
checking process, and also verify the concepts or issues brought up by
other members.  There is a greater confidence that an issue is valid.
\item By working in a group, others can detect your errors, which you can
learn from since you might not have seen yourself.  Although not everyone will agree with you, you learn to
work together and share ideas peacefully.  Working alone sometimes is
hard when you don't have any idea of what's going on.  There is a
better chance that someone in your group will know the answer if you
don't.

\item my understanding of C++ is not thorough hence working in group I can
learn about the aspects of the language about which I am not confident.
\item Questionnaire

I am not very familiar with C++, so it helps to discuss in a group.
\item During EIAM, I had doubts whether an error has occurred or not.  I
could not refer to anyone in the group, so I had to determine it
myself.  That was pretty difficult.  With EGSM, I could refer to
anyone when there is something questionnable.  I enjoy the collective
reasoning we had during the session, and I can't say that I haven't
learned a thing or two from it.  In fact, I saw errors that I would
never have caught in the same amount of time.
\item I usually work alone, but I find it very helpful to work in groups.
Usually
bugs and logic errors come into focus much more quickly when in a
group.  I think that the type of group members matters a great deal,
and when the group memebers are poor, I'd rather work alone.  However,
such was not the case with this group.
\item I like writing code alone - but review without running is where I
would like some discussion.
\item I felt more confident in raising issues in the group session in
general. But, as I mentioned before, I think this is most valuable
when using both methods of evaluation.
\item group helps finds errors (or correct each other on finding errors)
faster and better.

\item In a group, we come up a lot of suggests and questions which is
helpful and useful.  I can learn things from others, and some of the
errors which I can't find out are found out by them.


\item I don't know my c++ well enough to contribute to discussions, so
working in a group is better for me to learn more.
\item Both styles have thier strong points.  However combining the 2 is better.
\item In this case, it was easier to have more people looking at the same
code.  Rather than just one knowledge base, there were three.  If one
person didn't understand, there were others around to help.  Working
alone, there was no one to ask, I have to rely on my own knowledge,
which is rather limited.

\item Still not very familiar with using this program.  Furthermore, not
good at C++ programming make the process harder as it takes more time
to understand the code first and remember the name of the valuable in
order to follow the logic of the program.  Working as a group is
better becasue the others know more about c++ and can identify more
errors than working by myself who is not familiar working with c++
yet.  

\end{itemize}

\item  Reasons for preferring working alone:
\begin{itemize}
\item spend less time argueing about whether the isssue is valid.
\item I trust my own judgement generally and I can follow my own line of
reasoning rather than relying on others to give me ideas.
\item I could get a lot more done in a given amount of time if working by
myself, but may not see the errors that I can in a group session.

\end{itemize}

\end{enumerate}

