%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Plain-Tex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgments.tex -- 
%% Author          : Dadong Wan
%% Created On      : Thu Apr  7 01:04:21 1994
%% Last Modified By: Dadong Wan
%% Last Modified On: Thu Apr 28 15:51:45 1994
%% Status          : Unknown
%% RCS: $Id: acknowledgments.tex,v 1.5 94/04/28 15:52:00 dxw Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{

Dissertation research is a solitary journey. As a researcher in
collaborative work, I also view it as a joint effort. The work reported
here, at least, would not have reached the present point without support,
discussions, and other forms of {\it collaboration\/} with many people, a
few of whom are mentioned below.

Philip Johnson, my advisor, provided me the opportunity and the environment
for the past three years. I thank him for taking a chance on me, for his
intellectual, financial, and emotional support, for the time he spent
reviewing my code and commenting drafts of my dissertation chapters, and
for his vision of making his group world-class in research and development.

I am grateful to other members of my committee for their guidance and
input. Daniel Watanabe raised a number of insightful questions during both
my proposal and final defenses. Martha Crosby increased my awareness of
cognitive and behavioral issues related to CLARE. Bill Remus made several
suggestions on improving the CLARE experiment. Steve Goldberg was a source
of immense encouragement. I thank Steve for his enthusiasm for what I was
doing, and for his perspective on how CLARE might be used in ways that had
not yet occurred to me.

I would like to express my special thanks to Jintae Lee for several
valuable discussions we had about CLARE and for his insights on such issues
as marketability and parsimonious representation. I also thank him for
encouraging his students to use CLARE.

I feel fortunate for being part of a supportive group while working on this
project. I thank both current and former members of the Collaborative
Software Development Laboratory (CSDL): Danu Tjahjono, Robert Brewer, Cam
Moore, Rosemary Andrada, Carol Izaki, Bruce Campbell, and Ram
Hariharan. Their contribution ranges from helping refine my research ideas,
reviewing my design and code, to willingly serving as guinea pigs at
various stages of the CLARE prototype.

I am grateful to a number of friends who have strengthened and challenged
me at difficult moments of my life: Andres Edwards, Yong Ma, Kexian Hu,
Dave Kim, Spike Yamamoto, and Scotty Blaisdell. Special thanks go to Brock
and Anne Roby, whose convictions on what they believe in often reminded me
what ultimately matters in life.

Finally, I thank my parents for their understanding of me not writing to
them for months and not visiting them for years. I wish I had devoted
myself to them as much as I did to this work. On the other hand, I know
deeply in my heart that, even if I had, I could still not, nor will I ever
be able to, pay back for what they have given to me. For that reason, I
dedicate this dissertation to them.

}



