%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% abstract.tex -- 
%% Author          : Carleton Moore
%% Created On      : Wed Dec 30 09:46:12 1998
%% Last Modified By: Carleton Moore
%% Last Modified On: Thu Feb 25 10:39:10 1999
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Carleton Moore
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%\documentclass[12pt]{article}
\documentstyle[twocolumn,icse99,times]{article}
%\RequirePackage{times}

%\usepackage{url}

%\oddsidemargin 0in   %   Note that \oddsidemargin = \evensidemargin
%\evensidemargin 0in
%\marginparwidth 0pt
%\marginparsep 10pt        % Horizontal space between outer margin and 
%\textheight = 8.45in
%\textwidth 6.5truein     % Width of text line.
%\topmargin 0.0in        %    Nominal distance from top of page to top of
%                        %    box containing running head.
%\headheight 0pt         %    Height of box containing running head.
%\headsep 0pt            %    Space between running head and text.
%% for some reason setting topskip to 0 creates a extra blank page in front!
%% So we use 1 pt. RSB 9/28/98
%\topskip = 1pt          %    '\baselineskip' for first line of page.

\begin{document}
\title{Project LEAP: Personal Process Improvement for the Differently Disciplined}
\author{
        \hspace*{-2ex}
        \parbox{2.3in} {\begin{center}
        {\authornamefont Carleton A. Moore}\\ 
        CSDL\\
        ICS Department\\
        University of Hawaii, Manoa\\
        Honolulu, Hawaii 96822  USA \\
        (808) 956-6920\\
        cmoore@hawaii.edu
        \end{center} }}
\maketitle
\copyrightspace

% Page number on the initial page can be omitted in both the review and
% final submission (and should be removed in the final submission).  The
% line below does that.

\thispagestyle{empty}  % suppresses page number on first page

% In the review submission, page numbers should appear (they can be omitted
%  from the first page).  The pagestyle command below puts them in.
% In the final submission of accepted papers, page numbers should be
%  omitted; remove or comment out the pagestyle line below to omit them. 

% \pagestyle{plain}

% Use \section* instead of \section to suppress numbering for
% the abstract, acknowledgements, and references.

\section*{ABSTRACT}

\section*{Research Area}
Process Improvement, Measurement, Personal Software Process
\section*{Problem}

Software developers and managers have faced the problem of producing
quality software since the beginning of the computer age.  Many people have
studied the software quality problem and have proposed many solutions.  We
can categorize these different solutions into two groups: (1) Top Down
solutions, that focus on software development as a group effort and (2)
Bottom up solutions, that focus on the individual software developer.  Some
of the many Top Down solutions include: the Capability Maturity Model,
Clean Room development, software quality assurance groups, and Formal
Technical Review.  These top down methods help improve the quality of the
software, however they may not be enough.

In the past four years, there has arisen a new focus on the individual
software developer.  One such effort is Watts Humphrey's Personal Software
Process (PSP)\cite{Humphrey95}.  In PSP, software engineers record the time
they spend programming, the defects they find in their software and the
size of the software.  Based upon these measurements, the engineer can
track their productivity, make better predictions for future projects, gain
insight to what types of errors they make, and learn how to remove defects
earlier in their development process.  The PSP, as described by Humphrey,
is a completely manual process.

After using the PSP for two years, we noticed three general problems.
First, we started to question the quality of the data recorded.  We noticed
that we did not record all of our defects, in part because the overhead of
recording each defect is too expensive.  Anne Disney and Philip Johnson
conducted a study to look at the data quality of PSP data.  They found that
there are significant data quality issues with manual PSP.\cite{Disney98,
  Disney98a}

Second, our experiences with industrial partners and management practices
and Robert Austin's book ``Measuring and Managing Performance in
Organizations''\cite{Austin96} made us think about the issues of
measurement dysfunction in PSP and review data.  An organization may
pressure their members to produce ``good'' results.  There are many ways
that the members can manipulate the personal data collected in the PSP to
get the ``right'' results.

Third, after four years, the results with long term adoption of PSP are
mixed.  Pat Ferguson and others report excellent results with PSP adoption
at Advanced Information Services, Motorola and Union Switch and
Signal\cite{Ferguson97}.  However, Barry Shostak and others report poor
adoption of PSP in industry\cite{Shostak96,Emam96}.

%Fourth, we had significant problems with PSP's strict water fall process
%for software development and data collection.  We wanted the flexibility to
%define our own development processes and explore the possibilities of
%collecting different types of data on our processes.

These issues started us thinking about designing an automated, empirically
based, personal process improvement tool.  Our goal is to reduce the
collection and analysis overhead for the engineer, to reduce the
measurement dysfunction of the collection process and allow the engineer to
use their own development style.  We are also incorporating collaborative
review support into our personal process improvement tool.  Adding review
support allows the developer to gain insight from other developers.  This
group input is an important feature lacking in the PSP.  These features
should improve the benefits to the engineer and the long term adoption of
empirically based process improvement.  To pursue this work, we initiated
Project LEAP, {\bf http://csdl.ics.hawaii.edu/Research/LEAP/LEAP.html}, and
began developing the Leap Tool Set.

\section*{Research Questions}

We intend to deploy the Leap Tool Set in both academic and industry
settings in order to investigate the following research questions:
\begin{itemize}
  
\item{What are the strengths and weaknesses of the Leap Tool Set and
    empirically based process improvement?}

\item{What are the barriers to adoption of the Leap Tool Set}
  
%\item{What are the benefits to the users of Leap?}

\item{How do users improve after using Leap and what are the kinds of
    improvements we can make using Leap beyond improved estimation?}

%\item{Is the integration of collaborative review and personal data
%    collection appropriate?}
  
\item{Is Leap an appropriate form of automated support for personal process
    improvement?}

\end{itemize}

Based upon these research questions we have developed the following 
testable hypotheses:
\begin{itemize}
\item{Automating data collection and analysis will lead to improved
    adoption of personal software process improvement.}

\item {Reducing the constraints on developers imposed by PSP will lead to
    improved adoption.}
  
\end{itemize}

\section*{Evaluation}

To evaluate these hypotheses we plan on conducting atleast two case
studies.  We will conduct the first pilot case study on senior level
undergraduates in Spring, 1999.  The undergraduates will learn about
empirically based process improvement and software engineering.  We will
train the students how to use the Leap Tool Set and they will use it to
record their development of several projects.  The students will submit
reports on their progress and findings.  We will interview the students to
determine their feelings and attitudes toward process improvement and the
Leap Tool Set.  We will also conduct a survey of their perceptions of the
Leap Tool Set and their environment.  The interviews and the surveys will
allow us to predict the students' adoption of the Leap Tool Set and
empirically base process improvement practices.  A few months after they
finish the class we will contact the students and ask them to fill out a
survey.  This survey will determine if they are still using Leap, and/or
any of the empirically base process improvement concepts in their work.
This survey will allow us to determine the level of adoption of Leap and
empirically based process improvement, and any barriers to adoption of
Leap.

We plan on conducting the second case study in an industrial research group
during Summer, 1999.  We will introduce the Leap Tool Set as automated
support for collaborative review and personal process improvement.  We will
train the members on the use of Leap and help them conduct reviews and
analyses.  During the training we will conduct a survey and interviews to
determine their perceptions of Leap and personal process improvement.
Again the surveys and interviews will allow us to predict the adoption of
Leap.  A few months after the training we will contact the members and
conduct a survey and interviews.  This survey will determine the level of
adoption of Leap, any barriers to adoption, and the member's attitude
toward personal process improvement.


\section*{Contributions}

We expect the following contributions from this research:
\begin{itemize}
  
\item{The Leap Tool Set provides a novel form of automated support for
    empirically based process improvement, including time, defect, size,
    and pattern recording and analysis. It is implemented in Java and runs
    on Windows, Unix, and Macintosh. You may down load it at
    {\bf http://csdl.ics.hawaii.edu/Tools/LEAP/LEAP.html}. }
  
\item{Insight into adoption of personal process improvement.  The results
    of this research will provide new insight into barriers to adoption of
    personal process improvement.  We designed the Leap Tool Set to
    overcome some of the known barriers to adoption.  If Leap is not
    adopted, then this suggests that these barriers are not key barriers.}
  
\item{Insight into empirical process improvement.  The use of Leap helps us
    learn what improvements we can make using empirical measurement.  The
    case studies may provide insights into the limits of empirically based
    approaches to process improvement.}
  
\item{Insight into process improvement issues.  The PSP uses the classic
    waterfall software development model, fixed forms, and a single size
    definition. Leap relaxes many of these constraints.  This research
    should help us learn if these constraints are required. }


\end{itemize}

%%% Input file for bibliography
\bibliography{/group/csdl/bib/psp}
%% Use this for an alphabetically organized bibliography
\bibliographystyle{plain}

\end{document}
