%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% proposal-abstract.tex -- 
%% Author          : Jitender Miglani
%% Created On      : Thu Aug 24 20:23:23 2000
%% Last Modified By: Jitender Miglani
%% Last Modified On: Mon Aug 28 13:12:00 2000
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2000 Jitender Miglani
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\uppercase{\abstractname}{

Effective software project estimation is one of the most challenging and 
important activities in software development. Usually, Project estimations are
done in large organizations to determine the cost and time required to complete
the project. This requires collection and analysis of large amounts of data and
above all the correct interpretation of data.

With the introduction of PSP in 1995 and rapid evolvement of a large number of
small Internet organizations, focus of software process improvement started 
shifting from large Internet organizations to individual developer. In PSP, 
developers gather their own measurements related to their work products and 
use these measures to drive changes in their development behavior.

Even though PSP is very useful for improving the productivity of a developer, 
it still requires collection of large amounts of data by the developer. That's
why the tools like LEAP were developed to help the developer collect and 
analyze the project data. However, that is still viewed by developers as a 
distraction from the task at hand. So, The need for a total automation of data
collection and analysis process were felt.

This thesis investigates whether it is possible to develop a tool which 
automatically collects the project size and time data without giving any extra
overhead to the user and still produces useful project estimation results. How
can we achieve the total automation in the process of data collection ? Can 
the tool work standalone, like LEAP, or a deep integration with some kind 
of development enviornment is required to have the total automation of the data
collection process.

}



