%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 06-09-abstract.tex -- 
%% Author          : Takuya Yamashita
%% Created On      : Mon Aug 28 16:08:49 2006
%% Last Modified By: Takuya Yamashita
%% Last Modified On: Mon Aug 28 16:08:49 2006
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 2006 Takuya Yamashita
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

Software engineers generally agree that code reviews reduce development costs and improve software quality by finding defects in the early stages of software development. In addition, code review software tools help the code review process by providing a more efficient means of collecting and analyzing code review data. On the other hand, software organizations that conduct code reviews often do not utilize these review tools. Instead, most organizations simply use paper or text editors to support their code review processes. Using paper or a text editor is potentially less useful than using a review tool for collecting and analyzing code review data.

In this research, I attempted to address the problems of previous code review tools by creating a lightweight and flexible review tool. This review tool that I have developed, called ``Jupiter'', is an Eclipse IDE Plug-In. I believe the Jupiter Code Review Tool is more efficient at collecting and analyzing code review data than the text-based approaches.

To investigate this claim, I have constructed a method to compare the Jupiter Review Tool to the text-based review approaches. I carried out a study using both approaches in a software engineering course with 19 students.

The results provide some supporting evidence that Jupiter is more useful and more usable than the text-based code review, requires less overhead than the text-based review, and appears to support long-term adoption.

The major contributions of this research are the Jupiter design philosophy, the Jupiter Code Review Tool, and the insights from the case study comparing the text-based review to the Jupiter-based review.

\end{abstract}
