%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Plain-Tex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%
%% thesis-acknowledgment.tex -- 
%% Author          : Dadong Wan
%% Created On      : Thu Apr  7 01:04:21 1994
%% Last Modified By: Anne Disney
%% Last Modified On: Sun Aug  2 13:05:20 1998
%% Status          : Unknown
%% RCS: $Id: acknowledgments.tex,v 1.5 94/04/28 15:52:00 dxw Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1994 University of Hawaii
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgements{


\indent To Jeanette Teehee, Kim Carr, Cyndee Grady, Dad, and Mom: Your kind
words of encouragement have made the last two years much easier.

To Jim Wuerstlin, Bryan Sakka, Dad, Gayle Johnson, and Clare Joslin: You
actually wanted to know what my research was about!

To Charles: You dropped me off at the airport and picked me up again every
week over three semesters, and didn't make me hitchhike even once!

To my friends at CSDL: Your support has been invaluable.  You all have
given me useful feedback as you critiqued my presentations and reviewed my
writing. Robert, your advice has been helpful on many subjects.  Cam, I've
asked you many, many questions on a wide variety of topics; you've almost
always known the answer and have never made me feel like an interruption or
annoyance. Jen, need I say more than ``ICS 613'' or ``Probability,
Statistics, and Queuing Theory''? Yes, actually, I do: Thank you for
allowing me to use your PSP data in this research and for being a steady,
hardworking partner through all those late nights and early mornings.

To everyone at Infoworld Management Systems; particularly Jim Sullivan,
Linda Augustine, and Laura Croce: You have never complained about my
unpredictable schedule and curtailed working hours.  You never asked, ``Are
you ever going to get done?'' or urged me to finish faster. I appreciate
your tremendous flexibility while I have pursued this goal.

To Cyndee and Denny Grady: Without your friendship and your generous offer
of a place to stay, I could not have continued after moving from O'ahu.
There are really no words for friends like you. And to Nicole: you're too
young to understand this yet, but your joyful cries of ``Auntie {\it Anne}
is home!'' cheered me up many times after a tiring day at school.

Finally, to Philip Johnson: Your enthusiasm gave me a vision of what I
could accomplish.  Your expertise and world-class teaching showed me how to
do it.  Your confidence in me has made everything possible.  Thank you!}











