%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% essay.tex -- 
%% Author          : Robert Brewer
%% Created On      : Wed Jan 27 17:17:21 1999
%% Last Modified By: Carleton Moore
%% Last Modified On: Mon Feb  1 13:01:01 1999
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1999 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\section{Essay Question}
%No more than 700 words

\begin{quote}
  {\em As an Aspect Technology Fund grant recipient, how would you contribute
    to the field of technology and promote the spirit of entrepreneurship?}
\end{quote}

%Personal qualifications
%Information technology as a new base for Hawaii's economy
%Consulting (a service business) as opposed to hard-core software development
%  as a means of generating revenue.

As an Aspect Technology Fund grant recipient,  I will contribute to the
field of technology and promote the spirit of entrepreneurship in many
ways.  

This project will change the technology of process improvement.  The basis
for this project, Leap, is personal process improvement tool.  We are
expanding Leap to group process improvement by using the same analyses but
looking at the data from many different users.  By combining the data from
many users we should find general or common trends that will teach us about
the process of business plan development.  The combined data will lead to
checklists and patterns that developers can use to avoid or solve the
problems with business plan development.  The combining of data and
producing solutions is a bottom up process that is very different from
traditional top down process improvement.  Most process improvement imposes
process on developers, not develops the process from the developers.


This project will use current information technology to expand business
plan review.  We are going to expand business plan review to the Web.
Reviewers will not have to be in the same place to conduct a review of a
business plan.  We want to focus on local Hawaii issues yet have access to
outside points of view.  Web based review will allow us to use the
expertise of people from around the world to look at local issues.

Another information technology we are going leverage is web based data
repositories.  We will combine web based data repositories with process
information.  The data repository will allows to search for interesting
process patterns and produce checklists to avoid process pit falls.  By
posting the data repository on the web we hope to share our insights with
entrepreneurs. 

To go along with the theme of sharing, we are going to use the Open Source
licensing model to distribute the tools.  Others can use and modify our
technology to help improve their organizations.  If successful our
technology will enable experts, including ourselves, to consult with many
different firms and organizations to help improve their processes.  

By focusing on business plan development and improving business plans I
hope to help entrepreneurs get their new businesses off the ground easier.
We can easily share the hard learned lessons of other entrepreneurs.  Also
by focusing on local issues, we will help improve the entrepreneurial
climate in Hawaii.  We can learn a great deal about Hawaii's economic
environment from the local repository of checklist and patterns for
successful business plans.  These lessons should help guide our leaders in
improving our economic environment.  Even if it does not, entrepreneurs
will have valuable insight into the specific issues of doing business in
Hawaii.  Hawaii can become a paradise for of high quality business ideas
and plans.

As more and more entrepreneurs learn how do develop high quality business
plans in Hawaii they can expand their scope.  They can start consulting
with other businesses in Hawaii and around the world.  An entrepreneur who
is able to produce several successful business plans can teach others how
to do the same.  I hope to see many Leap Business Plan Web sites helping
communities around the world.

If this project is successful we will develop a valuable database that will
be a valuable resource for research into business development in Hawaii.
The repository of successful patterns will help entrepreneurs develop good
ideas for new businesses.  More new and exciting businesses opportunities
will be explored after this project is finished.

Not only will this repository help entrepreneurs, but it will help
researchers.  Researchers can learn about the common issues with developing
a business plan in Hawaii.  If we develop another database for another area
of the country researcher could do some comparisons between the areas and
gain valuable insights into the economic forces in Hawaii and the world.


