%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% bio.tex -- 
%% Author          : Philip Johnson
%% Created On      : Tue Nov  4 10:26:48 1997
%% Last Modified By: Philip M. Johnson
%% Last Modified On: Mon May 23 10:49:09 2005
%% RCS: $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1997 Philip Johnson
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
\documentclass[11pt]{article} 
\input{/export/home/csdl/tex/psfig/psfig}
\usepackage{/export/home/csdl/tex/icse2003/latex8}
\usepackage{times}
\usepackage[final]{graphicx}
% uncomment the % away on next line to produce the final camera-ready version
% and uncomment the \thispagestyle{empty} following \maketitle
%\pagestyle{empty}



\begin{document}

%\section*{Biographical Sketch}

%%      Command to define new categories:
\newcommand{\newcategory}[1]{\newenvironment{#1}
 {\sectionheading{#1}\begin{list}{}{\setlength{\labelwidth}{0cm} \setlength{\labelsep}{0cm} \setlength{\itemsep}{0ex plus0.2ex} \setlength{\itemindent}{-1cm} \setlength{\leftmargin}{1cm} \setlength{\parsep}{0ex plus0.2ex}}}{\end{list}\par}}
\newcommand{\sectionheading}[1]{\medskip\pagebreak[2]\par\noindent
 {\small\bf #1}\nopagebreak}

\begin{center}
Philip M. Johnson\\
Information and Computer Sciences \hfill (808) 956-3489

University of Hawaii              \hfill fax: (808) 956-3548

1680 East-West Road               \hfill johnson@hawaii.edu

Honolulu, HI~~~96822              \hfill http://www.ics.hawaii.edu/$\sim$johnson/

\end{center}

\newcategory{Professional Preparation}
\begin{Professional Preparation}
\item Ph.D.~in Computer Science, University of Massachusetts. 1990 
\item M.S.~in Computer Science, University of Massachusetts.  1985
\item B.S.~in Biology, University of Michigan. 1980
\item B.S.~in Computer Science, University of Michigan. 1980
\end{Professional Preparation}

\newcategory{Appointments}
\begin{Appointments}
\item {\em Professor},  Information and Computer
  Sciences, University of Hawaii.  2001---present

\item {\em Associate Professor},  Information and Computer
  Sciences, University of Hawaii.  1995---2001

\item {\em Senior Research Fellow},  Distributed Systems Technology Centre,
University of Queensland, Australia, 1997.

\item {\em Assistant Professor},   Information and Computer
  Sciences, University of Hawaii.  1990---1995

\end{Appointments}




\newcategory{Publications: Closely Related}
\begin{Publications: Closely Related}

\item P.~M.~Johnson and H.~Kou and M.~Paulding and Q.~Zhang and A.~Kagawa
and T.~Yamashita, {\em Improving Software Development Management through
Software Project Telemetry}, IEEE Software, August, 2005 (to appear).

\item S.~Faulk and P.~M.~Johnson and J.~Gustafson and A.~Porter and W.~Tichy
and L.~Votta, {\em Measuring HPC Productivity}, International Journal of
High Performance Computing Applications, December, 2004.


\item P.~M.~Johnson and H.~Kou and J.~Augustin and C.~Chan and C.~Moore and
J.~Miglani and S.~Zhen and W.~Doane, {\em Beyond the Personal Software
Process: Metrics collection and analysis for the differently disciplined},
Proceedings of the 2003 International Conference on Software Engineering,
Portland, Oregon, May, 2003.
  
\item P.~M.~Johnson and C.~A.~Moore and J.~A.~Dane and R.~S.~Brewer, {\em
    Empirically Guided Software Effort Guesstimation}.  IEEE Software, Vol.
  17, No. 6, December 2000.
  
\item P.~M.~Johnson and A.~M.~Disney, {\em A Critical Analysis of PSP Data
    Quality: Results from a Case Study}.  Journal of Empirical Software
  Engineering, Volume 4, December, 1999.

\end{Publications: Closely Related}


\newcategory{Publications: Other Significant}
\begin{Publications: Other Significant}

\item P.~M.~Johnson and M.~Paulding, {\em Understanding HPCS development
through automated process and product measurement with Hackystat}, Second
Workshop on Productivity and Performance in High-End Computing (P-PHEC),
February, 2005.

\item P.~M.~Johnson and H.~Kou and J.~Augustin and Q.~Zhang and A.~Kagawa
and T.~Yamashita, {\em Practical automated process and product metric
collection and analysis in a classroom setting: Lessons learned from
Hackystat-UH}, Proceedings of the 2004 International Symposium on Empirical
Software Engineering, Los Angeles, California, August, 2004.

\item P.~M.~Johnson, M.~L.~Moffett, and B.~T.~Pentland, {\em Lessons
Learned from VCommerce: A virtual environment for interdisciplinary
learning about software entrepreneurship}, Communications of
the ACM, December, 2003.

\item P.~M.~Johnson and A.~M.~Disney, {\em The Personal Software Process: A
    Cautionary Case Study}.  In IEEE Software, Volume 15, No. 6, November,
  1998.

\item P.~M.~Johnson, {\em Design for Instrumentation: High Quality
    Measurement of Formal Technical Review}.  Software Quality Journal,
  Volume 5, March, 1996.

\end{Publications: Other Significant}


\newcategory{Synergistic Activities}
\begin{Synergistic Activities}
\item {\em Editorial Boards}:\\ 
Journal of Empirical Software Engineering, 2004-present.\\
IEEE Transactions on Software Engineering, 2000-2004.

\item {\em Program Chair}: \\
Workshop on Software Engineering for High Performance Computing, 2004, 2005.\\
International Software Engineering Research Network Annual Meeting, 2000.

\item {\em Member, Board of Directors}:\\
Hawaii Strategic Development Corporation, 1999-present.\\
LavaNet, Inc., 2002-present.

\item {\em Advisory Board Member:}\\ 
  International Journal of Computer Supported Cooperative Work

\item {\em Program Committees}: \\
  XP/Agile Universe, 2004-present.\\
  PROFES 2004-present\\
  International Software Metrics Symposium, 2003-present.\\
  International Symposium on Empirical Software Engineering, 2002-present.\\
  European Conference on Computer Supported Cooperative Work, 1997,1999.\\
  Software Architectures for Cooperative Systems Workshop, 1994 \\
  CSCW Tools and Technologies Workshop, 1992,1993 \\

\end{Synergistic Activities}



\newcategory{Collaborators}
\begin{Collaborators}
\item J. Augustin, A, Kagawa, H.Kou, D. Port, M. Paulding, T. Yamashita, Q. Zhang, University of Hawaii.
\item S. Faulk, University of Oregon
\item A. Porter, University of Maryland
\item L. Votta, J. Gustafson, Sun Microsystems
\item W. Tichy, University of Karlsruhe
\item D. Reed, University of North Carolina
\end{Collaborators}


\newcategory{Thesis Advisor and Postgraduate-Scholar Sponsor}
\begin{Thesis Advisor and Postgraduate-Scholar Sponsor}
\item Qin Zhang, University of Hawaii
\item Carleton Moore, Orincon, Inc.
\item Danu Tjahjono, Aloha Networks, Inc.
\item Dadong Wan, Accenture, Inc.
\item Robert Brewer, LavaNet, Inc.
\item Mette Moffett, hotU, Inc.
\item Monir Hodges, Honolulu Community College.
% CIS Ph.D. students (D. Wan, D. Tjahjono, R. Tokuyama), 
% MS students (J. Geis, A. Disney, R. Brewer, C. Moore, R. Andrada, K. Ram, J. Johnson, F. Palit)
% EE Masters student (W. Kwan)

\item {\em Total graduate students advised:} 27

\end{Thesis Advisor and Postgraduate-Scholar Sponsor}

\newcategory{Graduate and Postdoctoral Advisors}
\begin{Graduate and Postdoctoral Advisors}
\item Jack Wileden, University of Massachusetts
\item Victor Lesser, University of Massachusetts
\item Wendy Lehnert, University of Massachusetts
\item David Stemple, University of Massachusetts
\item Daniel Corkill, Blackboard Technology Group
\end{Graduate and Postdoctoral Advisors}











\end{document}
