\chapter{Anticipated contribution} \label{contribution}
As the main contribution of my research to the software engineering and especially software process community I am envisioning development of a previously unexplored approach of discovering recurrent behaviors in software process through data mining of low-level process and product artifacts.

The results of the experimental work, and evaluation of the approach taken, will be another contribution of my research. By performing these, I am expecting to discover unknown patterns in the software process, improving our knowledge about it.

The software framework and data mining toolkit which I will develop for my research will be integrated with existing family of Hackystat tools contributing software-engineering community.

\section{Estimated Timeline}
By the beginning of the Fall'09 semester, I plan to design a taxonomy and implement a low-level software process events abstraction into symbolic representation. The database schema will be extended for storing these data and indexes. Immediately after that, I will proceed with the implementation of fundamental algorithms for associated and sequential patterns mining. This new data model and KDD toolkit will allow me to discover more low-level behavior patterns some of which might be found useful though the empirical evaluation during the period of data collection from the Fall'09 Software engineering class. I will be working with those students who will demonstrate a relevant recurrent behavior in software process for clarification of observed phenomena.

By the end of the Spring'10 semester I am planning to finish all of my experiments and software system implementation, and will move on to a peer-reviewed publication and the dissertation thesis defense.
