\chapter{Methods} \label{methods}
My research in the software process discovery mainly rests on a mining of recurrent behavioral patterns from a representation of software processes as a temporal sequences of events performed by individual developers or automation with or without concurrency. As we seen in the related work section, it is possible not only to infer the known high-level processes from observing such a low-level artifacts, but also to discover a novel processes through the use of a unsupervised process mining techniques.

In the context of my research, the collection of development artifacts is performed by Hackystat, a framework for automated software process and product metrics collection and analysis. The event streams provided by Hackystat are very rich in terms of enclosed information. It provides temporal data about atomic events, such as invoking of a build tool or performing a test, along with a great variety of process and product metrics such as cyclomatic complexity of the code, or amount of effort applied to the software development and many other. It is possible to retrieve other very low-level artifacts such as buffer transfers within the IDE editor or background compilation activities. All of these artifacts characterize the dynamic behavior of a software process in the very detail.

In order to perform analyses in my system, I am extracting the necessary process data with a desired granularity from the Hackystat and converting it into the symbolic representation by performing a direct mapping based on the taxonomy of events or by approximating telemtery streams (time-series) with Piecewise Aggregate Approximation (section \ref{paa}) and Symbolic Aggregate approXimation (section \ref{sax}). This symbolic representations of the observed software processes are used in the Trajectory analyses which built upon discussed in this chapter Symbolic Temporal Data Models (section \ref{tconcepts_models}), Temporal Concepts (section \ref{tconcepts}) and Temporal Operators (section \ref{tconcepts_operators}). 

The Temporal patterns and indexing (\ref{tpatterns}) section defines \textit{temporal patterns} of \textit{motif} and \textit{surprise} along with discussing relevant pattern search algorithms and data structures used for the indexing. The last section presents AprioriAll algorithm for unsupervised pattern mining.