\chapter{Research Questions}
\label{ch:researchquestions}
The long-term goal of my research is to understand how to characterize
and improve low-level software development behaviors.  As a step in
that direction, I am focusing for my Ph.D. research on a specific kind
of low-level software development behavior: Test-Driven
Development. The Zorro system, which attempts to infer TDD low-level
development behaviors, provides a way to partially evaluate the
overall approach and begin to understand its strengths and
limitations.

Zorro infers developer's TDD development behaviors using SDSA. It is
easy for software developers to collect in-process development
activities using Hackystat sensors, and it is also easy for them to
evaluate their TDD development behaviors using Zorro. If Zorro's TDD
inference is correct, then we can use it to assess TDD process
conformance during the daily practice of TDD as well as during empirical
studies of TDD. However, does Zorro infer developers' TDD development
behaviors correctly? Will it falsely categorize some non-TDD
development behaviors as TDD? Or, will it misinterpret some TDD
development behaviors as non-TDD? To answer these questions, we need 
to conduct validation studies of Zorro. Some of the most important research
questions are: 
\begin{itemize}
\item Q1: Can Zorro automate the recognition of Test-Driven Development
using automatically collected low-level software development activities?  
\item Q2: Can Zorro help to improve the practice of TDD? 

This is a hard question, but we can divide it into three small questions
with regard to user's roles.
   \begin{itemize}
   \item For beginners, can Zorro help them improve the compliance to TDD?
   \item For experienced TDD practitioners, will Zorro help them
   improve their TDD practice by analyzing their TDD development
   behaviors?
   \item For researchers, can Zorro help them reach legitimate
   research conclusions on TDD experiments by providing the TDD
   process conformance information.
   \end{itemize}
\end{itemize}

Answering these questions requires a ``mixed methods'' research strategy
\cite{Creswell:03}. Questions Q1 can be investigated by evaluating 
Zorro's data collection and TDD inference capability using field observation
research method. Investigating question Q2 requires research methods such  
as collecting users' feedback or interviewing them. In my research, 
I designed a series of case studies using these research methods to 
investigate the research questions I presented above.