\chapter{Implementation}
\label{chap:Implementation}

%%\begin{comment}
\section{Graphic View of Software Testing Process}
Software process is constructed by a series of analysis, design,
development, testing and debugging activities. It starts from requirement
analysis and ends after software products are delivered. Rational Unified
Process (RUP), Personal Software Process (PSP), Team Software Process (TSP)
and Extreme Programming (XP) are some well-defined software processes.
These processes were defined by process pioneers from their best practice
and critical thinkings on their development activities. All processes are
constructed by a set of rules and advices from requirement analyses to
testing. They exist in many kinds of software development organizations and
the rules are enforced by development team leaders or managers. Software
development process is thought as intellectual, non-repeatable and invisible.

In my thesis work I will focus on studying tests in software development,
especially Test-Driven Development to see how tests are being created and
exercised by developers incrementally. A test process view is going to be
implemented to display and analyze how developers create and execute tests
in their development. This tool is called TDPViewer, which stands for Test
Development Process Viewer. With this view support I will be able to
study development process to see whether developers follow a set of
demanded rules.
\section{Process Pattern and Quantification}
Speaking of unit testing execution in software development it could be
either test-first as specified by Test-Driven Development, test-last, or
hybid mode of test-first and test-last. In Hackystat we collect both the
development activities including implementation, compilation, unit testing
and debugging in Eclipse IDE so it is clearly feasible to study how unit
tests are implemented in the development process. Software process rules
can be used to generate development patterns to categorize how developers
do unit testing in the real implementation. One thought here is to design a
rule-based agent to study the development pattern [further research to be
conducted] to do the categorization.
%%\end{comment}




















