\chapter{Publication List}
\label{app:publication-list}

These are the publications that have come out of the Kukui Cup project that I have authored or co-authored:

\section{Conference Papers}

\begin{itemize}

\item \textbf{Robert S. Brewer}, Yongwen Xu, George E. Lee, Michelle Katchuck, Carleton A. Moore, and Philip M. Johnson. Energy feedback for smart grid consumers: Lessons learned from the Kukui Cup. In \emph{Proceedings of the Third International Conference on Smart Grids, Green Communications and IT Energy-aware Technologies (ENERGY 2013)}, Lisbon, Portugal, March 2013. To Appear.

\item Philip M. Johnson, Yongwen Xu, \textbf{Robert S. Brewer}, Carleton A. Moore, George E. Lee, and Andrea Connell. Makahiki+WattDepot: An open source software stack for next generation energy research and education. In \emph{Proceedings of the 2013 Conference on Information and Communication Technologies for Sustainability (ICT4S)}, February 2013.

\item Philip M. Johnson, Yongwen Xu, \textbf{Robert S. Brewer}, George E. Lee, Michelle Katchuck, and Carleton A. Moore. Beyond kWh: Myths and fixes for energy competition game design. In \emph{Proceedings of Meaningful Play 2012}, October 2012.

\item \textbf{Robert S. Brewer}, George E. Lee, and Philip M. Johnson. The Kukui Cup: a dorm energy competition focused on sustainable behavior change and energy literacy. In \emph{Proceedings of the 44th Hawaii International Conference on System Sciences}, January 2011.

\item \textbf{Robert S. Brewer} and Philip M. Johnson. WattDepot: An open source software ecosystem for enterprise-scale energy data collection, storage, analysis, and visualization. In \emph{Proceedings of the First International Conference on Smart Grid Communications}, Gaithersburg, MD, October 2010.

\end{itemize}


\section{Workshop Papers}

\begin{itemize}

\item \textbf{Robert S. Brewer}. The Kukui Cup: Shaping everyday energy use via a dorm energy competition. In \emph{Proceedings of the CHI 2011 Workshop on Everyday Practice and Sustainable HCI}, Vancouver, Canada, May 2011.

\item \textbf{Robert S. Brewer}, George E. Lee, Yongwen Xu, Caterina Desiato, Michelle Katchuck, and Philip M. Johnson. Lights Off. Game On. The Kukui Cup: A dorm energy competition. In \emph{Proceedings of the CHI 2011 Workshop on Gamification}, Vancouver, Canada, May 2011.

\end{itemize}
