%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% thesis-acknowledgments.tex -- 
%% Author          : Robert Brewer
%% Created On      : Fri Sep 25 14:35:00 1998
%% Last Modified By: 
%% Last Modified On: Mon Mar 31 11:24:03 2003
%% RCS: $Id: thesis-acknowledgments.tex,v 1.3 2000/03/17 21:26:58 rbrewer Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\begin{acknowledgments}
This research would not be possible without the following people who have
provided me with guidance, support, and encouragement along the way.

First I would like to thank Dr.~Philip Johnson for his patience, endless
words of encouragement, and letting me use his ICS 414 class.  No matter
how worried or frustrated I became, you always reminded me that things were
never as bad as I thought they were.  Thank you for bringing out the best
in me and taking the chance two years ago on a student who had no clue
about what potential she possessed.

Next, I would like to thank the following past and present members of CSDL:
Mike Staver, William Albritton, Jitender Miglani, Hongbing Kou, Aaron
Kagawa, Chris Chan, and Takuya Yamashita.  You have each in your own way
kept me sane throughout the ending half of my college career.  Mike,
William, Jitender, and Hongbing -- we served the most time together.  I
thank you guys for always being there with a ready ear to listen to
anything and everything that ever came out of my mouth (both research and
non-research related) and always being ready talk story.

I would also like to thank my committee members, Dr.~Wes Peterson and
Dr.~Martha Crosby, for for their time and effort in evaluating my thesis as
well as all the help they have given me along the way.

To the Fall 2002, ICS 414 class, thanks for participating in the evaluation
of this research.  Without you guys, my life would have been that much more
difficult.  You also gave me my first taste of working with ``customers''.

Last, but not least, I would like to thank my family: mom, dad, and Robert.
Behind every great achievement is a loving and supportive family.  You have
all patiently listened to all of my doubts and insecurities and turned them
into words of encouragement and support no matter how early in the day or
late in the night the phone rang.  Thank you for always being there when I
needed you the most.
\end{acknowledgments}
% LocalWords:  tex Sep Exp Agustin
% LocalWords:  CSDL Philip Staver Albritton Jitender Miglani
% LocalWords:  Hongbing Kou Kagawa Chan Takuya Yamashita Peterson Crosby
