\begin{acknowledgments}
 
I would like to thank many people without whom this endeavor would not be possible. My adviser, Philip Johnson, was on my side throughout my graduate studies giving me guidance and encouragement that were essential for its completion. Prof. Jessica Lin helped me with the understanding of symbolic discretization. Sergey Malinchik helped me with SAX-VSM development.  

I would like to thank my committee members: Prof. Guylaine Poisson, Prof. Kyungim Baek, Prof. Henri Casanova, and Prof. Daniel Port for taking the time to read and evaluate this dissertation.

I thank the Hackystat hackers for the initiation into software process analysis, ideas, and discussions. I am grateful to have received support from the research laboratories where I worked throughout the years. Many people, including Maqsudul Alam, Thomas Brettin, Chris Detter, Patrick Chain, Jean Challacombe, Monica Misra, Jacques Nicolas, Mondher Bouzayen, Christophe Klopp, and Maria-Soledad Goni Urriza have taken an active interest in my research and provided me with much needed encouragement.

Jimmy Saw, Tracey Freitas, Johanne Aub\'{e}, Aaron Kagawa, and Robert Brewer deserve a special mention as the senior students I have collaborated with and who have been role models I looked up to.

Finally, I acknowledge the long-suffering and seemingly inexhaustible support of my family. I thank my wife Inna, who has put up with a great deal during the pursuit of my ``mad dream'' -- she deserves much of the credit for the dissertation's completion. And last, but certainly not least, a special thanks to my kids, Alexander and Daria, who put a smile on my face after long days and helped me to put things in perspective.

\end{acknowledgments}