\chapter*{Glossary}
\addcontentsline{toc}{chapter}{Glossary}  

\begin{list}{}{\setlength\labelwidth{0pt}%
           \setlength\itemindent{-\leftmargin}%
           \let\makelabel\descriptionlabel}

\item[1NN] -- \textbf{One} \textbf{N}earest \textbf{N}eighbor. A variant of kNN classification where we assign each entity to the class of its closest neighbor.            
           
\item[DIRECT] -- \textbf{DI}viding \textbf{RECT}angles algorithm. A general parameters optimization algorithm that is based on the iterative partitioning of the search space into hyper-rectangles \cite{citeulike:4210208} \cite{citeulike:12563460}. 

\item[FLOSS] -- \textbf{F}ree/\textbf{L}ibre/\textbf{O}pen \textbf{S}ource \textbf{S}oftware. A computer software that can be classified as both free software and open-source software, and is licensed for free use, copying, study, and change. This software model is the opposite to proprietary software, which is typically closed-source and distributed under a restrictive copyright.

\item[PAA] -- \textbf{P}iecewise \textbf{A}ggregate \textbf{A}pproximation. A technique for reducing the dimensionality of time series by aggregating its segment to their mean values \cite{citeulike:2946589}.

\item[SAX] -- \textbf{S}ymbolic \textbf{A}ggregate appro\textbf{X}imation. A technique for symbolic discretization of the continuous signal \cite{citeulike:2821475}.

\item[SCM system] -- \textbf{S}oftware \textbf{C}onfiguration \textbf{M}anagement system. A software system which enables tracking and controlling changes in the software \cite{scm_book}.

\item[SCRUM] -- An iterative and incremental agile software development framework for managing product development \cite{Cohn_SCRUM}.

\item[TF*IDF] -- \textbf{T}erm \textbf{F}requency -- \textbf{I}nverse \textbf{D}ocument \textbf{F}requency. A numerical statistic that is intended to reflect how important a word is to a document in a collection of documents \cite{salton-71}. 

\item[TDD] -- \textbf{T}est \textbf{D}riven \textbf{D}evelopment. A software development process that relies on the repetition of a very short development cycle: coding of an automated test, producing the minimum amount of code to pass that test, and refactoring the new code to acceptable standards \cite{Beck_TDD}.

\item[XP] -- E\textbf{x}treme \textbf{P}rogramming. One of several lightweight software development methodologies based on values of simplicity, communication, feedback, courage, and respect \cite{Beck_XP}.

\end{list}