%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% thesis-acknowledgments.tex -- 
%% Author          : Robert Brewer
%% Created On      : Fri Sep 25 14:35:00 1998
%% Last Modified By: Robert Brewer
%% Last Modified On: Thu Mar 16 12:03:50 2000
%% RCS: $Id: thesis-acknowledgments.tex,v 1.3 2000/03/17 21:26:58 rbrewer Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\begin{acknowledgments}

  Like any research project, this one could not have been completed without the
  help of a number of people. I'd like to thank my parents for putting me
  through college and encouraging my academic ambitions. Your lifelong
  scholarship set an example for me.
  
  My thanks go to Philip Johnson, for putting up with me and my smart-alecky
  ways. You have always supported my efforts, even when they were at best
  vaguely related to the group's research focus. You even supported my decision
  to temporarily leave the program (though I think you should have taken me up
  on my investment offer :)
  
  I'd also like to thank the past and present members of CSDL in chronological
  order: Dadong Wan, Danu Tjahjono, Rosemary Sumajit, Cam Moore, Anne Disney,
  Jennifer Geis, Russell Tokuyama, Joe Dane, Tie Fang, Monir Hodges, and Mette
  Moffett. Being able to bounce ideas off of all of you was incredibly valuable
  to me.
  
  Kirill Levchenko deserves mention as someone outside UHM who provided
  feedback about the research. It was appreciated (even if your first comment
  on it was ``weren't you going to do something cool?'' :)
  
  I would like to thank my committee members Wesley Peterson and Edoardo
  Biagioni for taking the time to read and evaluate this thesis.
  
  Tim Endres is the maintainer of the mailing list which was used in the case
  study of MCS. His willingness to let his list be used and his general support
  is appreciated.
  
  Last but not least, I thank Yuka Nagashima for all that she has done in
  support of me. You have put up with griping, provided encouragement when I
  was depressed, and read drafts at the drop of a hat. Thank you.

\end{acknowledgments}
% LocalWords:  tex Sep Exp alecky Dadong Danu Tjahjono Sumajit Geis Tokuyama
% LocalWords:  Kirill Levchenko Wes Edo Biagioni Monir Mette Edoardo
% LocalWords:  Endres Moffett
