\documentstyle [nftimes,12pt,/home/4/johnson/Tex/letteradj,
                     /home/4/johnson/Tex/definemargins] {letter}  


%%%%%%% check this
%\makelabels

\definemargins{1.0in}{1.0in}{1.0in}{1.0in}{0.3in}{0.3in}
\begin {document}
\name {Philip Johnson}
\signature {Philip Johnson \\Assistant Professor}

\address{548 Kaimake Loop\\
         Kailua, HI 96734}
\telephone{(808) 956-3489}  %%department phone
%\telephone{(808) 262-3845}  %%home phone
%\fax{(808) 956-3548}

\netaddress{johnson@hawaii.edu}
\indentclosing

\letterhead{coins}
\begin {letter}
  {Catherine Russell\\
   Publisher's Editor\\
   Chapman and Hall\\
   2-6 Boundary Row \\
   London SE1 8HN}

\opening {Dear Ms. Russell,}

Enclosed please find the corrections to the page proofs for my article
(reference number 930716).  I apologize for the lateness of these
corrections.  In the interests of time, I am itemizing the corrections to
the bibliography section in this letter, and faxing along with it the
pages of the manuscript where changes should be made.  I will return the
hardcopy by regular postal mail.

\begin{itemize}
\item Arthur, 1993.  Where published: New York, NY.
\item Brothers et al, 1990.  Where held: Los Angeles, CA.
\item Freedman, 1990. Publisher/where published: Little, Brown, Inc. New
York, NY.
\item Gintell et al, 1993. Pages: 72-90.
\item Hutchings, 1993.  Volume: 36(10), Pages: 104-113.
\item Johnson, 1994a.   Pages: 87-107.
\item Johnson, 1994b.  Where held: Sorrento, Italy.  Pages: 113-122.
\item Johnson, 1994c. Page numbers: No page numbers in this proceedings.
\item Mashayekhi, 1994. Pages: 66-75.
\end{itemize}

Also, I believe the last page of the paper was not supplied to me for
corrections (my copy ends on page 18).  In the event that the end of the
paper is actually missing, the final paragraph continues from the bottom
of page 18 as follows:

\begin{quotation}
\noindent CSRS.  Many review factors can and should be methodically tested
within the CSRS framework to discover their optimum values within
a particular organization and application context.  Some of these
factors include: artifact size and complexity, review team size
and composition, private review analysis technique, review checklist
composition (if checklists are used), public review scope and
duration, individual and team effort, and scope and duration of 
group review. 
\end{quotation}

Thanks very much.

\closing{Sincerely,}
\end {letter}
\end {document}
